Metabolic compensation via gluconeogenesis explains the non-essentiality of glycogen phosphorylase as an insecticidal target in Plutella xylostella

Version 1

Abstract

Benzoylphenylurea (BPU) insecticides disrupt insect chitin formation, yet their molecular target is debated. Although resistance-associated mutations map to the Chitin Synthase (CHS) gene, direct inhibition of CHS has not been demonstrated. Given the structural similarity of BPUs to mammalian glycogen phosphorylase (GP) inhibitors, GP has been proposed as a potential target controlling chitin precursor flux. We characterized Plutella xylostella GP (PxGP) and found that the human GP inhibitor GPI potently inhibits both recombinant PxGP (IC₅₀ = 2.96 nM) and native larval GP activity (57.5% reduction), while the BPU diflubenzuron showed no effect. Despite this biochemical inhibition, neither GPI treatment nor RNAi-mediated PxGP knockdown (87.6% suppression) caused mortality or developmental defects. Mechanistic analysis revealed coordinated downregulation of glycogenolysis genes (trehalase 69%, hexokinase 77%) and robust upregulation of gluconeogenolysis genes (PEPCK 3.95-fold, G-6-Pase 3.34-fold). Critically, protein catabolism (29.4% decline at 72 h) provided substrates for gluconeogenolysis by releasing amino acids. Metabolite profiling validated the full substrate-to-product pathway: transient glucose shortfall was followed by massive accumulation of trehalose (7.4-fold) and glucose-6-phosphate (6.5-fold), confirming a quantitatively sufficient metabolic rescue via de novo glucose synthesis. These findings demonstrate that BPUs do not target GP, and that GP inhibition alone is not lethal due to metabolic compensation. This highlights a fundamental metabolic plasticity that must be considered in future target-based insecticide design.

Introduction

The molecular target of benzoylphenylurea (BPU) insecticides—a multi-billion-dollar class of insect growth regulators widely used in agriculture—remains one of the most significant unresolved controversies in insecticide toxicology [1]. Since their introduction over over four decades ago, BPUs have been known to catastrophically disrupt chitin formation, causing molting failure and larval death [2]. Despite this clear phenotypic effect, their direct molecular target has not been definitively identified, creating a persistent gap between their in vivo activity and their presumed mechanism of action.

The prevailing hypothesis posits that BPUs directly inhibit chitin synthase (CHS), the enzyme catalyzing the final polymerization of UDP-N-acetylglucosamine (UDP-GlcNAc) into chitin. This hypothesis is supported by indirect evidence, including the accumulation of UDP-GlcNAc upon BPUs treatment [3] and the genetic co-localization of BPU resistance mutations with the CHS locus in multiple insect species, most notably Plutella xylostella [1, 4, 5]. However, direct biochemical studies, beginning with the seminal work by Deul et al. in 1978 [6], have consistently failed to demonstrate CHS inhibition by BPUs in cell-free assays, even at supralethal concentrations [7-9]. This discrepancy between genetic association and enzymatic dataology argues against it—represents an unresolved paradox that has significant implications for understanding BPUs resistance mechanisms and for rational insecticide design.

A largely overlooked clue may help resolve this contradiction: the core acylurea structure of BPUs (Figure 1) closely resembles that of potent inhibitors of mammalian Glycogen Phosphorylase (GP) developed as therapeutic agents for type 2 diabetes [10, 11]. GP catalyzes the rate-limiting step of glycogenolysis, the phosphorolytic cleavage of glycogen to glucose-1-phosphate, which serves as the metabolic gateway controlling flux into the chitin biosynthetic pathway [12]. In insects, this pathway is particularly critical: GP-derived glucose is converted to trehalose (the primary hemolymph sugar), which is subsequently channeled through a series of enzymatic steps to generate UDP-GlcNAc, the substrate monomer for chitin synthesis [13]. GP also plays key roles in carbohydrate homeostasis, osmotic regulation, stress adaptation [14-17]. This structural similarity between BPUs and GP inhibitors (GPIs), combined with GP’s upstream regulatory position in chitin precursor synthesis, raises a compelling alternative hypothesis: BPUs may exert their insecticidal effects not by inhibiting CHS directly, but by targeting GP and thereby depleting chitin biosynthesis of its essential precursors.

Structural comparison of benzoylphenylurea (BPU) insecticides and a mammalian glycogen phosphorylase inhibitor (GPI) reveals a shared acylurea scaffold.

Chemical structures of four BPU insecticides (Diflubenzuron, Teflubenzuron, Novaluron, and Flufenoxuron) and a human glycogen phosphorylase inhibitor (GPI) are shown. All share a central N,N’-diphenylurea (acylurea) core (dashed boxes). BPUs are characterized by halogenated (F, Cl) aromatic substituents and in some cases by additional ether or alkyl side chains. The GPI contains a methoxyphenyl group bearing an additional methylurea side chain. This structural similarity underpinned the hypothesis that insect GP might be a molecular target for BPUs.

Here, we systematically test this hypothesis and evaluate GP’s potential as an insect control target. We cloned, expressed, and biochemically characterized GP from the globally significant agricultural pest P. xylostella (PxGP), measured its inhibition by the BPU—diflubenzuron (DFB) and a specific mammalian GPI, and assessed the physiological consequences of GP loss-of-function in vivo. Our findings definitively exclude GP as the target of DFB. More significantly, we identify a robust metabolic compensatory response—the upregulation of gluconeogenesis—that completely rescues insects from GP inhibition. This discovery not only explains why GP alone is not a viable insecticidal target but also reveals a principle of metabolic plasticity that must be considered in the design of metabolism-targeting pesticides.

Results

Recombinant P. xylostella glycogen phosphorylase (PxGP) is a functional, activatable enzyme

To investigate its biochemical properties and evaluate its potential as an insecticidal target, the full-length PxGP cDNA (predicted molecular mass of 98.1 kDa) was cloned and expressed in Sf9 cells using the Bac-to-Bac baculovirus expression system. Following purification via immobilized metal-affinity chromatography (IMAC) of the N-terminal 6×His tagged protein [18], SDS-PAGE analysis and Western blot analysis confirmed a highly pure protein band migrating at approximately 100 kDa, consistent with its theoretical mass (Figure 2A). A total yield of 7.5 mg of purified protein was obtained from 8.25 g of cell pellet (specific yield=0.91 mg/g cells).

Recombinant PxGP is inhibited by a mammalian GP inhibitor (GPI) but not by the benzoylphenylure Diflubenzuron.

(A) Purification of 6×His-tagged PxGP from Sf9 cells. Left panel: Coomassie-stained SDS-PAGE.. Right: Western blot with anti-His antibody. The expected ∼100 kDa band is indicated (red box). M, marker; S, soluble fraction; P, pellet; F, flow-through; lanes 1–4, imidazole elution fractions. (B) Dose-response inhibition by GPI. Activated PxGP-a activity was measured after pre-incubation with GPI (0.16–500 nM). Data (mean ± SEM, n=3) were fit to a four-parameter logistic curve. The calculated IC₅₀ is 2.96 nM. Inset: GPI structure with its characteristic asymmetric side chains (orange). (C) Inhibition assay with DFB (0.54–1700 nM). DFB did not yield a complete sigmoidal inhibition curve, causing ≤62% inhibition at the highest concentration. Inset: DFB structure with its characteristic halogenated benzoyl groups (orange). This rules out PxGP as a direct target of DFB.

Glycogen phosphorylase activity is regulated by reversible phosphorylation, existing as a less-active dephosphorylated b form (PxGP-b) and a fully active phosphorylated a form (PxGP-a) [19]. Recombinant PxGP-b was activated by incubation with immobilized phosphorylase kinase (PhK) in the presence of ATP and Mg²⁺ for 2 h [20]. The activated enzyme showed robust catalytic activity in a coupled-enzyme spectrophotometric assay, with NADPH production (monitored by increase in A340) increasing linearly for the first 15 min and reaching a plateau by approximately 20 min, consistent with substrate depletion and/or product accumulation. The specific activity of activated PxGP-a was 294.25 ± 3.45 U/mg protein (Figure S1A), confirming that the recombinant enzyme is catalytically competent for inhibitor screening.

GPI potently inhibits PxGP in vitro and in vivo

Given the structural similarity between insecticidal BPUs and mammalian GPIs [11] (Figure 1), we tested whether PxGP could be a target of acylurea-scaffold compounds. GPI exhibited potent, dose-dependent inhibition of recombinant PxGP-a, with a sigmoidal dose-response curve typical of competitive or allosteric inhibition (Figure 2B). Non-linear regression analysis yielded an IC₅₀ value of 2.96 nM (95% confidence interval [CI]: 1.66–5.35 nM, Hill slope = 0.74, R² = 0.877, n = 3). Near-complete inhibition (>95%) was achieved at 500 nM GPI, indicating saturable, high-affinity binding. The potency of GPI against PxGP is comparable to its effect on mammalian orthologs (see below) [11].

In striking contrast, the BPU insecticide Diflubenzuron (DFB), which shares the acylurea core with GPI (Figure 1), did not effectively inhibit PxGP-a. No standard sigmoidal dose-response curve was observed over 0.54 nM to 1700 nM (Figure 2C). Maximum inhibition observed did not exceed 61.96 ± 2.54% (mean ± SEM) even at 100 µM, with high variability at lower concentrations. These data indicate that DFB, unlike GPI, is not an effective inhibitor of PxGP, ruling out GP as the direct molecular target of this BPU insecticide.

Consistent with the recombinant enzyme results, GPI effectively inhibited native GP activity in crude lysates from third-instar larvae in a dose- and time-dependent manner (Figure S1B). After 45 min of pre-incubation, 10 µM GPI reduced native GP activity by 57.52 ± 1.88% (n = 3). DFB, however, showed no significant inhibitory effect on native GP activity at any concentration tested (10⁻⁷, 10⁻⁶, or 10⁻⁵ M) or time point (5, 10, 20, or 30 min), with activity remaining at 97-108% of control (Figure S1C).

Collectively, these results establish PxGP as a high-affinity molecular target for GPI but definitively exclude it as a target for DFB. This resolves the primary hypothesis by demonstrating that BPUs do not inhibit insect glycogen phosphorylase.

Paradoxical finding: potent PxGP inhibition by GPI lacks in vivo toxicity

Because GPI potently inhibit PxGP, the rate-limiting enzyme in glycogenolysis, which supplies precursors for chitin synthesis [21], we expected it to exhibit significant insecticidal activity. In leaf-dip bioassays with third-instar P. xylostella larvae continuously exposed to GPI (250 or 500 mg/L) or DFB (125 or 250 mg/L), GPI caused no significant toxicity or developmental disruption across the entire time course (24 to 120 h post-treatment) (Figure 3A). At 120 h, mortality with 500 mg/L GPI was only 12.78 ± 2.42% (n = 3), which was not significantly different from the solvent control (CK: 5.00 ± 5.00%, P = 0.21) (Figure 3B). Larvae exposed to 250 mg/L GPI exhibited even lower mortality (2.50 ± 2.50%).

The GP inhibitor (GPI) is non-toxic to Plutella xylostella, unlike the benzoylphenylurea Diflubenzuron (DFB).

(A) Developmental stage distribution over time for control, GPI (250, 500 mg/L), and DFB (125, 250 mg/L) treated larvae. DFB causes developmental arrest. (B) Quantitative effects on mortality (top), pupation (middle), and adult emergence (bottom). GPI groups did not differ from control (P > 0.05), whereas DFB caused significant, dose-dependent mortality and severely impaired development (*P < 0.05). Data are mean ± SEM (n=3). Different letters denote statistical significance. (C) Larval phenotypes at 96 h. Control and GPI-treated larvae are normal. DFB-treated larvae show catastrophic molting failure and “double cuticle” (red arrows). Scale bar, 1 mm. (D) Pupal and adult phenotypes. GPI- treated individuals develop normally. DFB treatment results in deformed pupae and trapped pharate adults (red arrows). Scale bar, 1 mm.

In stark contrast, DFB induced significant, dose-dependent mortality. At 120 h, DFB treatment at 125 mg/L caused 41.60 ± 4.85% mortality (P < 0.01 vs. control) and at 250 mg/L caused 61.36 ± 10.02% mortality (P < 0.01 vs. control). Probit analysis of the DFB dose-response data yielded calculated LC₃₀ and LC₅₀ values of 141.900 mg/L (95% CI: 67.893–206.576 mg/L) and 229.537 mg/L (95% CI: 145.287-331.784 mg/L), respectively, with an acceptable goodness-of-fit (χ² = 11.946, P = 0.683) (Table S2).

Furthermore, GPI-treated larvae showed no developmental abnormalities. Pupation rates were 64.17 ± 4.79% for the 500 mg/L GPI group and 69.72 ± 6.81% for the 250 mg/L group, which were statistically indistinguishable from the control group (64.72 ± 9.33%, P > 0.05 for both comparisons) (Figure 3B, middle panel). Similarly, adult emergence rates were 56.11 ± 5.64% and 56.39 ± 6.24% for the 500 and 250 mg/L GPI groups, respectively, compared to 54.10 ± 11.91 % for the control (P > 0.05, Figure 3B). Phenotypically, GPI-treated larvae molted normally, fed actively, and displayed no cuticle defects (Figure 3C). DFB-treated larvae exhibited the characteristic phenotypes of BPU intoxication: catastrophic molting failure, larval-pupal intermediates, and complete failure of adult emergence (Figure 3C and 3D). This striking discrepancy between biochemical potency and in vivo inefficiency presents a profound paradox that demands mechanistic investigation.

GPI exposure triggers a compensatory upregulation of PxGP transcription

To explore this discrepancy, we measured PxGP gene expression after GPI exposure. RT-qPCR revealed that 500 mg/L GPI rapidly upregulated PxGP mRNA transcription. As shown in Figure 4, its relative expression increased 3.24-fold at 24 h, peaked at 3.48-fold at 48 h, and remained elevated at 96 h. This indicates that the insect responds to the pharmacological inhibition by transcriptionally increasing the de novo synthesis of the target enzyme.

Ingestion of GP inhibitor (GPI) induces compensatory upregulation of PxGP transcription.

Relative PxGP mRNA levels in larvae after continuous exposure to 500 mg/L GPI or control (CK). Expression was normalized to PxRPS13 (control set to 1.0 at each time point) using the 2-ΔΔCt method. Data are mean ± SEM (n=3). GPI triggered a rapid and sustained transcriptional increase, peaking at 48 h (3.48-fold). Significance vs. control: *P < 0.05, **P < 0.01, ***P < 0.001 (two-way ANOVA with Sidak’s test).

Sequence homology and rationale for target site selection

To achieve effective gene knockdown coupled with functional ablation, precise selection of the interference target site is critical. Sequence analysis of PxGP (837 residues) revealed that residues 88-680 form the active site pocket of the GT35_Glycogen_Phosphorylase family, and the C-terminal region contains a phosphatase-pyridoxal phosphate linkage site (residues 672-684). Eleven residues within positions 42-318 are conserved and correspond to the AMP-binding site in human GP (Figure 5A) [22].

PxGP sequence conservation and domain architecture.

(A) Schematic of PxGP functional domains: pyridoxal phosphate (PLP) binding site (blue boxes), AMP allosteric site (yellow boxes), and glycogen-binding/active site pocket (blue shading). (B) Multiple sequence alignment of GP from Plutella xylostella (PxGP), other lepidopterans (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP). Identical residues are shaded red. The gray box indicates the 606 bp region (aa 22–224) targeted for RNAi, which overlaps the conserved AMP-binding domain. PxGP shares 82.29% identity with human GP.

Multiple sequence alignment of PxGP with GP orthologs from other key lepidopteran pests (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP) demonstrated high conservation across species (Figure 5B). PxGP shares 82.29% sequence identity with human GP (Figure S2). Since GP activity is allosterically regulated by AMP binding and phosphorylation at Ser14, triggering a conformational shift from the inactive T-state GPb (unphosphorylated) to the active R-state GPa (phosphorylated) [19], we targeted the core AMP-binding site region (gray box in Figure 5B) for RNAi. Theoretically, interference in this region should prevent AMP-mediated activation of existing GPb while degrading the entire mRNA pool, drastically reducing total GP protein synthesis and ensuring complete functional ablation.

RNAi confirms PxGP is non-essential for larval development

The compensatory upregulation of PxGP mRNA expression in response to pharmacological inhibition (Section 3.4) indicated that the gene is under active transcriptional regulation and suggested that increased enzyme synthesis might partially mitigate the effects of GPI. However, this compensation alone could not fully explain the complete absence of phenotype. To determine whether a more profound loss-of-function—achieved by directly suppressing PxGP gene expression—is lethal, we performed RNAi to knock down PxGP at the transcriptional level.

Firstly, we quantified the expression profile of PxGP across developmental stages in P. xylostella, PxGP was highly expressed in fourth-instar larvae (L4) and adults (Figure 6), stages that correlate with the energy demands for metamorphosis and flight, respectively [12, 23]. Given the inherent time lag required for systemic dsRNA delivery to achieve effective post-transcriptional knockdown [24], dsRNA was administered during the mid-third-instar larval stage. Mid-third-instar larvae were microinjected with dsRNA targeting a conserved 606 bp region of PxGP (dsGP) or with a non-specific control dsGFP. Three concentrations of dsGP (6,000, 10,000, and 14,000 ng/µL) were tested, and knockdown efficiency was assessed by RT-qPCR at 24, 48, 72, and 96 h post-injection (Figure 7). The highest concentration (14,000 ng/µL) achieved maximal suppression at 48 h, reducing PxGP transcript levels to 12.47 ± 1.68% of the dsGFP control (87.59% knockdown, P < 0.001) (Figure 7B). Significant knockdown persisted at 72 h (69.86% suppression,P < 0.05), with partial recover by 96 h, likely due to dsRNA degradation and/or compensatory transcriptional upregulation. The 10,000 ng/µL dose also produced significant knockdown at 48 h (65.76% suppression, P < 0.01), whereas the 6,000 ng/µL dose showed more modest effects (Figure 7B).

Developmental expression profile of PxGP.

Relative PxGP mRNA levels across life stages (Egg, larval instars L1-L4, Pupa, Adult), normalized to PxRPS13 (L1 set to 1.0). Data are mean ± SEM (n=3). Expression peaks in adults. Different letters indicate significant differences (P < 0.05, one-way ANOVA with Tukey’s test).

RNAi-mediated knockdown of PxGP is dose- and time-dependent.

(A) Experimental workflow. Third-instar larvae were injected with dsRNA targeting PxGP (dsGP) at three concentrations or a control dsRNA (dsGFP), then sampled for analysis. (B) Time-course of PxGP mRNA levels post-injection, relative to dsGFP control (set to 1.0). The highest dsGP dose (14,000 ng/µL) caused maximal suppression (87.6%) at 48 h. Data are mean ± SEM (n=3). Significance vs. dsGFP at each time point: **P < 0.01, ***P < 0.001 (one-way ANOVA with Tukey’s test).

Despite this robust, transient silencing, no adverse phenotypic effects were observed. Developmental stage distribution and mortality rates in all dsGP-treated groups (6,000, 10,000, and 14,000 ng/µL) were statistically indistinguishable from the dsGFP control group across the entire 120 h observation period (P > 0.05 at all time point) (Figure 8A and 8B). At 120 h, cumulative mortality was 8.10 ± 4.23%, 9.82 ± 0.81%, and 3.33 ± 3.33% for the 6,000, 10,000, and 14,000 ng/µL dsGP groups, respectively, comparable to the dsGFP control (Figure 8B). Pupation and adult emergence also indistinguishable from controls (Figure 8B). Phenotypic examination revealed that dsGP-injected larvae developed normally, molted successfully, formed morphologically normal pupae, and emerged as viable adults without cuticle defects or developmental delays (Figure 8C and 8D). Thus, acute, profound suppression of PxGP expression is not lethal to P. xylostella during the critical larval-pupal transition, despite the enzyme’s established role in glycogenolysis and chitin precursor synthesis. This finding necessitates an investigation into the compensatory mechanisms that allow the insect to survive in the absence of functional GP.

PxGP knockdown does not impair development or survival.

(A) Developmental stage distribution post-injection shows similar progression across all groups. (B) Cumulative mortality (top), pupation (middle), and adult emergence (bottom) rates were not significantly different between dsGP and dsGFP control groups (P > 0.05). Data are mean ± SEM (n=3). (C) Representative larval phenotypes at 48 h and 96 h post-injection are normal in both dsGFP and high-dose dsGP groups. Scale bar, 1 mm. (D) Pupae and adults from both groups are morphologically normal. Scale bar, 1 mm.

PxGP knockdown coordinately suppresses the glycogenolysis-to-chitin pathway

To elucidate how larvae survive GP loss, we examined downstream genes linking glycogenolysis to chitin synthesis. Specifically, we measured the expression of trehalase (PxTre), which hydrolyzes circulating trehalose to glucose [12], and hexokinase (PxHex), which phosphorylates glucose to glucose-6-phosphate—a key branch point leading to UDP-GlcNAc synthesis for chitin production [25, 26].

RT-qPCR analysis revealed that suppression of PxGP triggered a coordinated transcriptional downregulation of both downstream genes (Figure 9A and 9B). At 48 h post-injection, coinciding with the peak of PxGP knockdown (87.59% suppression), PxTre expression was reduced to 30.46 ± 1.25% of the dsGFP control (69.54% suppression, P < 0.001,), and PxHex to 22.82 ± 3.39% of control (77.18% suppression, P < 0.001). This suppression persisted at 72 h (PxTre: 61.27 ± 2.21% suppression, P < 0.001; PxHex: 41.96 ±4.02% suppression, P < 0.001) and returned to near baseline levels by 96 h, mirroring the recovery kinetics of PxGP itself.

PxGP knockdown triggers biphasic metabolic reprogramming, culminating in gluconeogenic compensation.

(A, B) Concurrent suppression of downstream genes Trehalase (PxTre) and Hexokinase (PxHex) post-knockdown. (C) Upregulation of gluconeogenic genes PEPCK and G6Pase at 96 h. Data in A-C are mean ± SEM (n=3); *P < 0.05, **P < 0.01, ***P < 0.001 vs. dsGFP control. (D-H) Metabolite levels post-knockdown. (D) Total protein decreased at 72 h (P < 0.01). (E) Glycogen remained stable. (F) Trehalose surged 7.4-fold at 96 h (P < 0.001). (G) Glucose-6-phosphate (G6P) increased 6.5-fold at 96 h (P < 0.001). (H) Free glucose showed a transient dip at 72 h (P < 0.05) before recovery. Data are mean ± SEM (n=3); *P < 0.05, ***P < 0.001 vs. control at each time point.

This coordinated downregulation indicates that the insect responds to the loss of GP not by futilely attempting to maintain flux through a pathway that lacks its key upstream enzyme, but rather by implementing a broader transcriptional shutdown of the glycogenolysis-to-glucose-to-chitin pathway. This suggests a coherent metabolic response program, raising the question: if glycogenolysis is blocked, how does the insect obtain the glucose required for survival and chitin synthesis?

Biphasic metabolic response: transient depletion followed by gluconeogenic rescue

The absence of a lethal or developmental phenotype despite glycogenolysis pathway suppression suggested activation of alternative glucose synthesis pathway to compensate for GP loss. We therefore measured expression of key gluconeogenic enzymes and metabolite levels across the glycogeno-glucose axis.

Gene expression analysis reveals gluconeogenic activation

We assessed transcript levels of two rate-limiting gluconeogenic enzymes: phosphoenolpyruvate carboxykinase (PEPCK, encoded by PxPEPCK), which catalyzes the conversion of oxaloacetate to phosphoenolpyruvate; glucose-6-phosphatase (G-6-Pase, encoded by PxG-6-Pase), which catalyzes the dephosphorylation of glucose-6-phosphate to free glucose. These enzymes are reciprocally regulated relative to glycolysis/glycogenolysis and serve as molecular markers of gluconeogenic flux [27, 28].

At 96 h post-injection―a critical pre-pupal stage with high carbohydrate demand for pupal cuticle synthesis [29, 30]―both PxPEPCK and PxG-6-Pase were significantly upregulated in dsGP-treated larvae (3.95 ± 0.44-fold and 3.34 ± 0.20-fold versus dsGFP control, respectively; P < 0.001 for each) (Figure 9C and 9D). This indicates robust transcriptional activation of the gluconeogenesis pathway in response to GP loss.

Protein catabolism provides gluconeogenic substrate mobilization

To identify the carbon source for gluconeogenesis, we quantified total protein content in dsGP-treated larvae across the time course. Strikingly, a significant 29.39% reduction in protein was observed at 72 h post-injection (P < 0.01), coinciding precisely with the nadir of glucose insufficiency (Figure 9D).

This protein decline provides direct biochemical evidence for substrate mobilization: the degradation of cellular proteins liberates amino acids, which serve as the primary carbon source for gluconeogenesis in insects [27, 31-33]. The magnitude of protein loss (∼30%) is sufficient to provide substantial amino acid flux into the gluconeogenic pathway.

Metabolite profiling reveals a biphasic adaptation

Quantification of four key metabolites (glycogen, trehalose, glucose-6-phosphat [G6P], and free glucose) along the glycogen-glucose axis at 48, 72, and 96 h post-injection uncovered a dynamic, biphasic response (Figure 9E-9H).

Glycogen Content: Contrary to the expectation of accumulation upon GP blockade, glycogen levels remained unchanged relative to controls throughout the time course (P > 0.05) (Figure 9E). This paradox indicates that insects activate alternative, GP-independent pathways for glycogen catabolism―most likely via α-amylase or glycogen debranching enzymes [34-36]―to provide emergency glucose mobilization during the period of metabolic stress, suggesting that these compensatory pathways are sufficient for short-term energy needs but are eventually superseded by gluconeogenic glucose production.

Trehalose: Trehalose levels remained stable at 48-72 h but surged dramatically by 96 h (7.44-fold elevation, P < 0.001) (Figure 9F and 9G). This massive accumulation, coinciding precisely with the peak gluconeogenic gene (e.g., PEPCK and G-6-Pase) expression, indicates that newly synthesized glucose derived from gluconeogenesis is being actively channeled into trehalose synthesis. The magnitude of this accumulation indicates that gluconeogenic glucose production exceeds the immediate metabolic demands of the larva, creating a metabolic “overshoot” that ensures adequate carbohydrate reserves for the upcoming pupation [37].

G6P: G6P, which sits at the intersection of glycolysis, gluconeogenesis, the pentose phosphate pathway, and glycogen synthesis [38], exhibited a pattern mirroring that of trehalose (Figure 9G). G6P content declined gradually from 48 to 72 h, consistent with reduced glucose mobilization from glycogen, but then increased sharply at 96 h (6.49-fold elevation, P < 0.001). This surge directly reflects the massive influx of glucose generated via the gluconeogenic pathway, which enters the metabolic network at the G6P node. The coordination between G6P accumulation and trehalose accumulation confirms that gluconeogenesis-derived glucose is being efficiently converted into storage and transport carbohydrates.

Free Glucose: Free glucose levels showed the most subtle but perhaps most physiologically informative pattern (Figure 9H). At 48 h―coinciding with peak PxGP suppression―glucose content was statistically indistinguishable from controls (P > 0.05), indicating that initial homeostatic mechanisms (including GP-independent glycogen breakdown and dietary carbohydrate absorption) successfully buffer against acute GP loss. However, at 72 h, the nadir of the metabolic stress response, free glucose levels declined significantly, albeit modestly (18.15% reduction, P < 0.05). This transient dip reveals a critical window of metabolic insufficiency during which the insect’s compensatory mechanisms are insufficient to fully maintain glucose homeostasis. Importantly, by 96 h, glucose levels fully recovered to control values (P > 0.05), demonstrating that the gluconeogenic pathway successfully restores glucose homeostasis before critical developmental transitions occur.

Integrated interpretation

Collectively, these metabolite and gene expression data reveal a sophisticated, biphasic metabolic adaptation strategy that fully explains the paradox of why profound GP loss (87.59% knockdown) causes no mortality or developmental defects:

Phase I (emergency compensation, 48-72 h): Metabolic stress and emergency compensation during the initial response to PxGP suppression, the insect experiences genuine metabolic stress, evidenced by the glycogen levels neither accumulated nor were depleted and the transient reduction in free glucose at 72 h. The static glycogen content indicates activation of GP-independent catabolism pathways, which provide emergency glucose but are insufficient for complete homeostatic maintenance.The modest glucose dip at 72 h demonstrates that this stress is real but tolerable, likely due to several factors: (i) the magnitude is modest (only 18.15% reduction), (ii) the duration is brief (∼24 h), (iii) the timing coincides with the mid-L4 larval stage when energy demands are moderate and active feeding allows dietary carbohydrate supplementation, and (iv) trehalose stores (which remain stable during this phase) can be mobilized to buffer hemolymph glucose.

Phase II (gluconeogenic rescue, 96 h): Robust transcriptional upregulation of PEPCK and G-6-Pase (3-4-fold increases) drives massive de novo glucose synthesis from non-carbohydrate precursors such as amino acids, lactate, and glycerol. The surge of gluconeogenesis-derived glucose is rapidly converted into trehalose and G6P, creating a metabolic overshoot that restores basal glucose levels. This overcompensation ensures adequate carbohydrate supply for the energetically demanding process of pupal cuticle synthesis, which begins shortly after 96 h.

This substrate-to-product chain for gluconeogenesis-mediated metabolic rescue—protein degradation → amino acid release → gluconeogenic conversion → glucose synthesis → trehalose storage—constitutes a complete compensatory mechanism that sustains development despite profound GP ablation.

Discussion

Resolving a long-standing controversy: BPUs do not target GP

A central challenge in mechanism-based insecticide development is distinguishing potent in vitro biochemical inhibitors from physiologically effective in vivo toxicants. This study addresses this challenge by investigating the debated molecular target of BPU insecticides. For over 40 years, the field has contented with conflicting evidence: while genetic studies implicated chitin synthase [5], direct biochemical assays consistently failed to validate it as a direct BPU target [6, 39, 40]. The structural similarity between BPUs and mammalian GPIs proposed an alternative hypothesis that BPUs might inhibit insect GP, thereby limiting glucose-derived precursors for chitin synthesis. Our systematic investigation demonstrates that while a human GPI potently inhibits PxGP, the BPU insecticide diflubenzuron does not. This discrepancy is attributable to key substitutions in its aromatic halogen substituents, which preclude effective binding to GP’s active or allosteric sites. This provides the first direct biochemical evidence excluding GP as a candidate molecular target for BPUs. Consequently, BPU lethality must arise from a different mechanism, likely involving interference with chitin assembly, cuticle organization, or CHS trafficking rather than direct enzymatic inhibition, a mechanism that remains to be conclusively identified [21, 39, 40].

Metabolic plasticity and gluconeogenic compensation underlie GP non-essentiality

The resolution of the BPU target question raised a more profound and unexpected puzzle: despite being a potent GP inhibitor in vitro, the GPI exhibited no insecticidal activity in vivo. This paradox—wherein a nanomolar-potency inhibitor of a putatively essential metabolic enzyme causes no phenotype—demanded mechanistic investigation.

Our investigation shows that this tolerance stems from a robust, developmentally-coordinated metabolic compensation mechanism. Even acute RNAi-mediated suppression of PxGP expression (87.59% knockdown) induced no discernible phenotype, as insects activated a compensatory gluconeogenic pathway. This was evidenced by the concerted suppression of downstream glycogenolytic enzymes and, crucially, the significant upregulation of the key gluconeogenic enzymes PEPCK (3.95-fold) and G-6-Pase (3.34-fold) at 96 h post-knockdown. This response provides a metabolic “escape route,” synthesizing glucose de novo from non-carbohydrate precursors to bypass the blocked glycogenolytic pathway and maintain homeostasis, particularly during the high carbohydrate demand preceding metamorphosis. Future work utilizing transcriptomic and proteomic approaches will be essential to delineate the signaling pathways (potentially involving nutrient sensors like AMPK and transcription factors such as FOXO) that link GP suppression to this precise reprogramming of metabolic gene expression [41, 42].

Direct metabolite quantification validated this compensation and revealed its sophisticated temporal dynamics. The most striking finding was the lack of glycogen accumulation upon GP knockdown, a paradox explained by the activation of GP-independent glycogen catabolism via alternative enzymes like α-amylase (which hydrolyzes internal α-1,4-bonds) and glycogen debranching enzymes (which hydrolyze α-1,6-branch points) [34-36]. Definitive metabolic rescue emerged at 96 h, characterized by a dramatic surge in trehalose and G6P, directly driven by the 3-4-fold upregulation of PEPCK and G-6-Pase, providing definitive biochemical proof that de novo glucose synthesis is the primary rescue mechanism. This biphasic response—initial metabolic stress followed by delayed (48-72 h), robust compensation—creates a temporal buffer (96 h), explaining the absence of phenotypes despite severe GP knockdown. This temporal coordination suggests a programmatically timed, possibly hormonally regulated response that anticipates developmental energy needs. Critically, it reveals a metabolic vulnerability window (∼72 h); strategies that disrupt both glycogenolysis and gluconeogenesis simultaneously could overcome this compensation.

Implications for insecticide target selection: metabolic redundancy and the dual-target strategy

Our findings offer direct and actionable insights for the rational design of metabolism-targeting insecticides. A key lesson is that the viability of a metabolic enzyme as an insecticidal target is determined not solely by its biochemical function or pathway position but by the presence of compensatory metabolic routes. GP fails as a standalone target due to compensation via gluconeogenesis. This contrasts sharply with other carbohydrate-metabolizing enzymes that have proven to be highly effective insecticidal targets. For example, trehalase inhibitors such as Validamycin A are potent insecticides [43, 44], suggests that inhibition of trehalase does not cause other pathways to initiate glucose compensation. This distinction informs a refined strategy: dual-target inhibition. Simultaneously blocking both glycogenolysis (via GP) and gluconeogenesis (e.g., via PEPCK or G-6-Pase) would create irreversible glucose starvation, a “metabolic pincer” attack that could overcome adaptive compensation—a strategy with precedent in cancer therapy targeting both glycolysis and oxidative phosphorylation [45, 46]. Ultimately, our work highlights the value of systems-level metabolic modeling for target prioritization. Enzymes at metabolic branch points with redundant outputs are less vulnerable than those in non-bypassable pathways (e.g., chitin synthase, trehalase). Prospective application of computational flux balance analysis and constraint-based modeling, as supported by existing literature, can systematically identify such non-compensatable nodes for effective target selection [47].

Study limitations and future directions

We acknowledge several limitations that point to productive future investigations. First, the transient nature of RNAi-mediated gene silencing in Lepidoptera precluded assessment of long-term or stage-specific consequences of GP loss; its role in adult flight or reproduction warrants investigation via CRISPR/Cas9 knockout or advanced dsRNA delivery systems [48-50]. Second, the generalizability of the this gluconeogenic compensation mechanism across insect orders (e.g., Coleoptera, Diptera) remains to be tested, as metabolic regulation can vary with ecology and life history. Third, while our multi-layered evidence—from transcriptional upregulation to phenotypic rescue—provides robust qualitative support for the activation of gluconeogenesis, future studies employing ¹³C-isotope tracer analysis would be valuable to quantify the precise metabolic rates identify primary carbon precursors, and further refine potential metabolic targets for insecticide development [51].

Concluding remarks

In conclusion, our study definitively rules out GP as the target of BPU insecticides, resolving a long-standing controversy. More importantly, it uncovers a fundamental principle of insect metabolic biology: robust compensatory pathways, specifically gluconeogenesis, can buffer against the loss of key catabolic enzymes. This metabolic plasticity renders GP non-viable as a standalone insecticidal target but illuminates a strategic path forward: the development of dual-target inhibitors that simultaneously block glycogenolysis and gluconeogenesis. Our findings underscore the necessity of adopting a systems-level perspective in target-based insecticide discovery. Biochemical potency in vitro is necessary but not sufficient for in vivo efficacy; the latter requires consideration of metabolic network architecture, pathway redundancy, and compensatory mechanisms. As the pest management community confronts the twin challenges of insecticide resistance and environmental sustainability, this mechanistic understanding of metabolic robustness will be essential for designing the next generation of selective, effective and durable insect control agents.

Materials and Methods

Insects and rearing conditions

A susceptible laboratory strain of P. xylostella (L.), maintained for multiple generations at the College of Plant Protection, Northwest A&F University, without prior insecticide exposure, was used for all experiments. Larvae were reared on radish (Raphanus sativus) seedlings grown in vermiculite. Adults were provisioned with a 10% (w/v) honey solution. All insect stages were maintained in a controlled environment at 25 ± 1°C, 60 ± 5% relative humidity (RH), and under a 16:8 h (light:dark) photoperiod.

Cell culture and maintenance

Spodoptera frugiperda (Sf9) cells (Gibco, CA, USA) were cultured in Sf-900™ III SFM medium (Gibco) supplemented with 10% fetal bovine serum (FBS; Gibco) and 1% (v/v) Antibiotic-Antimycotic solution (Gibco, Cat# 15240062). Adherent monolayer cultures were maintained in T-flasks at 27°C. For large-scale protein expression, cells were adapted to suspension culture in baffled flasks, agitated at 120 rpm at 27°C. Cells were harvested by centrifugation (4,000 rpm, 10 min, 4°C) when the density reached 5.0 × 10⁶ cells/mL. Cell pellets were flash-frozen in liquid nitrogen and stored at -80°C until use.

Chemicals and reagents

Diflubenzuron (DFB, 1-(4-chlorophenyl)-3-(2,6-difluorobenzoyl) urea, ≥98% purity) was purchased from Shanghai Topscience Co., Ltd. (Shanghai, China, Cat# T0109).

The Glycogen phosphorylase-IN-1 (GPI), 1-(3-(3-(2-Chloro-4,5-difluorobenzoyl)ureido)-4-methoxyphenyl)-3-methylurea (also known as compound 42 in ref [11]), was purchased from TargetMol (Shanghai, China, Cat# T37577, ≥98% purity).

Glycogen (from bovine liver, Type III), β-glycerophosphate, HEPES, imidazole, EDTA, DTT, ATP, MgCl₂, protease inhibitor cocktail components (AEBSF, Pepstatin A, Leupeptin), and all other biochemical reagents were purchased from Sigma-Aldrich (Shanghai, China) or Sangon Biotech (Shanghai, China) and were of molecular biology grade (≥95% purity) unless otherwise specified.

Recombinant PxGP expression and purification

The full-length coding sequence of PxGP (2,511 bp, encoding 837 amino acids; GenBank Accession No. XM_038119125.2 was amplified from P. xylostella third-instar larval cDNA by PCR using gene-specific primers (Table S1), and cloned into the pFastBac1 vector (Invitrogen, Carlsbad, CA, USA) using homology-based cloning (ClonExpress II One Step Cloning Kit, Vazyme, Nanjing, China). The construct was engineered to express PxGP with an N-terminal 6×His tag (HHHHHH-) immediately following the initiating methionine, under the control of the polyhedrin (PH) promoter. The sequence fidelity of the construct was verified by Sanger sequencing (Tsingke Biotechnology, Beijing, China).

The sequence-verified pFastBac1-PxGP construct was transformed into DH10Bac™ competent cells (Invitrogen) to generate recombinant bacmid DNA via Tn7-mediated transposition, following the manufacturer’s Bac-to-Bac® protocol. Recombinant bacmid DNA was isolated, verified by PCR using M13 forward and reverse primers (Table S1), and transfected into Sf9 cells using Cellfectin® II Reagent (Invitrogen) to generate P0 viral stock. The viral titer was amplified through two additional passages (P1, P2) to obtain high-titer P3 viral stock.

For large-scale protein production, Sf9 cells in suspension culture (2.0 × 10⁶ cells/mL, 500 mL culture volume in a 2-L baffled Erlenmeyer flask) were infected with P3 viral stock at a Multiplicity of Infection (MOI) of 5. Infected cultures were maintained at 27°C with agitation at 120 rpm. Cells were harvested 72 h post-infection by centrifugation (4,000 rpm, 10 min, 4°C), yielding approximately 8.25 g of cell pellet. Cell pellets were flash-frozen in liquid nitrogen and stored at -80°C until use.

For protein purification, frozen cell pellets were thawed on ice and resuspended in ice-cold Lysis Buffer (50 mM β-glycerophosphate [pH 7.5], 150 mM NaCl, 0.2 mM DTT, 0.5 mM EDTA, and 1× protease inhibitor cocktail [1 mM AEBSF, 15 µM Pepstatin A, 20 µM Leupeptin]) at a ratio of 4 mL buffer per gram of cell pellet. Cells were lysed by sonication on ice using a Scientz-IID sonicator (Ningbo Scientz Biotechnology, China) with the following parameters: 6 cycles of 30 s ON / 60 s OFF at 40% amplitude. The lysate was clarified by centrifugation (16,000 × g, 30 min, 4°C).

The clarified supernatant was loaded onto a 5-mL HisTrap™ HP column (GE Healthcare, Uppsala, Sweden) pre-equilibrated with Lysis Buffer containing 10 mM imidazole. The column was first washed with 10 column volumes (CV) of Equilibration Buffer (Lysis Buffer + 10 mM imidazole), followed by 5 CV of Wash Buffer (Lysis Buffer + 30 mM imidazole). Bound 6×His-tagged PxGP protein was eluted with a 15 CV linear gradient from 30 to 200 mM imidazole in Lysis Buffer. Fractions (2 mL each) were collected and analyzed by 13% SDS-PAGE with Coomassie Brilliant Blue R-250 staining. Fractions containing PxGP, identified as a prominent band at ∼100 kDa, were pooled and dialyzed overnight at 4°C against Storage Buffer (50 mM HEPES [pH 7.4], 150 mM NaCl, 1 mM DTT, 10% [v/v] glycerol) using a 10 kDa molecular weight cut-off dialysis membrane (Spectrum Labs, Rancho Dominguez, CA, USA), with three buffer changes. Protein concentration was determined using the Pierce™ BCA Protein Assay Kit (Thermo Fisher Scientific, Waltham, MA, USA) with bovine serum albumin (BSA) as the standard. Aliquots of purified PxGP were flash-frozen in liquid nitrogen and stored at -80°C. The typical yield was 7.5 mg of purified protein from 8.25 g of cell pellet.

In vitro activation of recombinant PxGP

Purified recombinant PxGP exists predominantly in the inactive dephosphorylated b form (PxGP-b) and requires phosphorylation at a conserved serine residue (corresponding to Ser14 in mammalian GP) for conversion to the active a form (PxGP-a) [19]. Activation was performed in vitro using immobilized Phosphorylase Kinase (PhK) (Cat# HY-P2757, MedChemExpress, Monmouth Junction, NJ, USA).

Immobilized PhK beads (100 µL slurry) were washed three times with 1 mL of Kinase Buffer (25 mM β-glycerophosphate, pH 7.5, 0.5 mM EDTA, 0.25 mM DTT, 1× protease inhibitor cocktail) and resuspended in 500 µL of dialyzed PxGP-b solution (final protein concentration: 1 mg/mL). The activation reaction was initiated by adding ATP (sodium salt) to a final concentration of 7 mM and MgCl₂ to 22 mM. The suspension was incubated for 2 h at room temperature (22-25°C) with gentle end-over-end rotation.

Following incubation, the activated PxGP-a (present in the supernatant) was separated from the immobilized PhK beads using a magnetic separation stand. The supernatant was passed through a 0.22 µm syringe filter (Millipore) to remove any residual beads and was used immediately for enzymatic assays. The extent of activation was assessed by measuring specific activity (see Section 2.7). Activated PxGP-a was used fresh and not stored, as phosphorylated GP is subject to dephosphorylation by endogenous phosphatases over time.

Western blot analysis

Protein samples were separated by 13% SDS-PAGE (120 V, 90 min) and electrophoretically transferred to a nitrocellulose (NC) membrane (0.45 µm pore size, Millipore, Burlington, MA, USA) at 300 mA for 90 min at 4°C using a Bio-Rad Mini Trans-Blot® system. The membrane was blocked for 1 h at room temperature in 5% (w/v) non-fat dry milk in TBST (10 mM Tris-HCl, pH 7.4, 150 mM NaCl, 0.1% [v/v] Tween-20).

The membrane was incubated overnight at 4°C with an anti-His mouse monoclonal antibody (1:1500 dilution in blocking buffer; Cat# ABT2050, Abbkine, Wuhan, China). After three washes with TBST (10 min each), the membrane was incubated for 1 h at room temperature with an alkaline phosphatase (AP)-conjugated goat anti-mouse IgG secondary antibody (1:4000 dilution in blocking buffer; Cat# A0208, Beyotime Biotechnology, Shanghai, China). Following three additional TBST washes, immunoreactive bands were visualized using BM Purple AP Substrate (Roche, Basel, Switzerland) according to the manufacturer’s instructions. Images were captured using a ChemiDoc™ XRS+ imaging system (Bio-Rad, Hercules, CA, USA).

Glycogen phosphorylase activity assay

GP enzymatic activity was measured using a coupled-enzyme spectrophotometric assay kit (Cat# BC3345, Solarbio Life Science, Beijing, China). This assay quantifies GP activity by measuring the GP-catalyzed phosphorolysis of glycogen to glucose-1-phosphate (G-1-P). The product G-1-P is enzymatically converted to glucose-6-phosphate (G-6-P) by phosphoglucomutase (PGM), and G-6-P is subsequently oxidized by glucose-6-phosphate dehydrogenase (G6PDH), which stoichiometrically reduces NADP⁺ to NADPH. The rate of NADPH formation, which is directly proportional to GP activity, was monitored by measuring the increase in absorbance at 340 nm (A₃₄₀) using a Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland).

The reaction mixture (200 µL total volume per well in a 96-well UV-transparent microplate) contained the following components (as supplied in the kit): 160 µL of Reaction Buffer (containing glycogen [2 mg/mL final], inorganic phosphate [10 mM], AMP [1 mM], NADP⁺ [0.4 mM], PGM, G6PDH, MgCl₂, and appropriate buffer), 30 µL of deionized water, and 10 µL of enzyme sample (purified PxGP-a or crude larval lysate). The reaction was initiated by adding the enzyme sample, and A₃₄₀ was recorded continuously at 15 s intervals for 30 min at 25°C.

Specific activity was calculated using the following formula: Specific Activity (U/mg)=(ΔA340/min)/(ε×1×[ Protein ])×Vt/Vs×1000. where:

ΔA₃₄₀ represents the change in absorbance at 340 nm,

ε = molar extinction coefficient of NADPH at 340 nm (6.22 mM⁻¹ · cm⁻¹),

l = light path length (0.6 cm for 200 µL in a 96-well microplate),

[Protein] = protein concentration in the assay (mg/mL),

Vt = total reaction volume,

Vs = sample volume added.

One Unit (U) of GP activity is defined as the amount of enzyme that catalyzes the formation of 1 µmol of NADPH per minute under the assay conditions.

Inhibitor screening and IC₅₀ determination

For in vitro enzyme inhibition assays, purified activated PxGP-a was diluted to 10 µg/mL in Assay Dilution Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl₂. Test compounds (GPI, DFB) were prepared as 100× stock solutions in DMSO and serially diluted in Assay Dilution Buffer to achieve final assay concentrations ranging from 0.16 nM to 500 nM (for GPI) and 0.54 nM to 1700 nM (for DFB), respectively (final DMSO concentration ≤ 1% [v/v] in all wells, including vehicle controls).

The enzyme (10 µL of 10 µg/mL PxGP-a, final concentration 10 µg/mL in the assay) was pre-incubated with 10 µL of inhibitor solution (or vehicle control) for 15 min at room temperature in a 96-well microplate. The reaction was then initiated by adding 180 µL of the Reaction Buffer (as described in Section 2.7), and A₃₄₀ was monitored continuously for 30 min at 25°C.

Residual enzyme activity at each inhibitor concentration was calculated as a percentage of the vehicle control (1% DMSO alone). Dose-response curves were generated by plotting percent residual activity versus log[inhibitor concentration]. IC₅₀ values (the concentration of inhibitor required to reduce enzyme activity by 50%), Hill slopes, and 95% confidence intervals (CI) were calculated by non-linear regression analysis using a four-parameter logistic equation in GraphPad Prism 9.0 (GraphPad Software Inc., San Diego, CA, USA): Y= Bottom +( Top − Bottom )/(1+10(LogIC 50−X)× Hill slope ). where Y is percent residual activity, X is log[inhibitor concentration], Top and Bottom are the upper and lower plateaus of the curve (constrained to 100% and 0%, respectively), and Hill slope describes the steepness of the curve.

For each compound, at least three independent experiments were performed, each with triplicate technical replicates.

Preparation of larval crude lysate for in vivo activity

To assess the inhibitory effects of GPI and DFB on native GP activity in vivo, crude enzyme extracts were prepared from P. xylostella larvae. Healthy third-instar (L3) larvae (n = 30 per biological replicate, three replicates total) were collected, briefly washed in ice-cold phosphate-buffered saline (PBS, pH 7.4) to remove surface contaminants, and gently blotted dry on filter paper. Larvae were then homogenized on ice in 3 volumes (w/v) of ice-cold Homogenization Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl₂, 1× protease inhibitor cocktail).

The homogenate was centrifuged at 15,000 × g for 20 min at 4°C, and the supernatant (crude enzyme extract) was carefully collected, avoiding the lipid layer and pellet. Protein concentration was determined using the BCA assay, and extracts were either used immediately for activity assays or aliquoted and stored at -80°C (activity remained >90% for up to 2 weeks).

GP activity in crude lysates was measured using the same coupled-enzyme assay described in Section 2.7, with the following modifications: crude lysate (diluted to 2 mg/mL total protein) was used in place of purified enzyme, and pre-incubation with inhibitors was performed for 5 to 45 min to assess time-dependent inhibition. Final inhibitor concentrations tested were 10⁻⁵, 10⁻⁶ or 10⁻⁷ M.

Sequence alignment and bioinformatic analysis

The amino acid sequence of PxGP (XP_037975053.2) was obtained from the NCBI database. Multiple sequence alignment with orthologs from Homo sapiens (NP_002854.3), Oryctolagus cuniculus (NP_001075653), Helicoverpa armigera (XP_021190210.1), Spodoptera exigua (ACN78408.1), and Ostrinia furnacalis (AFO54708.2) was performed using ClustalX. Conserved domains and motifs were analyzed using the ScanProsite and NCBI CD-Search web servers. Protein pairwise similarity matrix was visualized as a heatmap using the “HeatMap” module in TBtools-Ⅱ (v2.390) [52].

Larval bioassays

Larval bioassays were conducted using a standard leaf-dip method adapted from established protocols [53]. Stock solutions of GPI and DFB were prepared in HPLC-grade acetone at 10,000 mg/L. Serial dilutions were prepared in deionized water containing 0.1% (v/v) Triton X-100 as a wetting agent to achieve final test concentrations of 250 and 500 mg/L for GPI, and 125 and 250 mg/L for DFB. The final acetone concentration in all treatment solutions was 1% (v/v). A solvent control group (CK) consisted of 1% acetone and 0.1% Triton X-100 in water without any test compound.

Fresh cabbage (Brassica oleracea var. capitata) leaf discs (5 cm diameter) were cut using a cork borer, dipped in the respective treatment solutions for 6-8 s with gentle agitation, and then air-dried for 30 min at room temperature in a fume hood. Once dry, each leaf disc was placed in a sterile 90 mm petri dish lined with moistened filter paper.

Ten synchronized third-instar larvae of P. xylostella (24-36 h post-molt to L3) were transferred to each treated leaf disc using a soft brush. Each treatment consisted of three biological replicates (i.e., three Petri dishes, 30 larvae total per treatment). Petri dishes were sealed with Parafilm® and maintained under standard rearing conditions (25 ± 1°C, 60 ± 5% RH, 16:8 h L:D photoperiod). Leaf discs were replaced with freshly treated discs every 12 h to ensure continuous exposure.

Mortality was recorded every 24 h for 120 h (5 days). Larvae that failed to respond to gentle prodding or showed moribund behavior (inability to right themselves or move coordinately) were considered dead. Corrected mortality was calculated using Abbott’s formula [54]: Corrected Mortality (%)=[( Treatment Mortality − Control Mortality )/(100− Control Mortality )]×100 Pupation rates (number of pupae formed / initial number of larvae × 100) and adult emergence rates (number of emerged adults / number of pupae formed × 100) were recorded at 8 and 10 days post-treatment, respectively.

Lethal concentration values (LC₃₀ and LC₅₀) and 95% confidence intervals for DFB were calculated using probit analysis in POLO-Plus software (LeOra Software, Berkeley, CA, USA).

GPI exposure for gene expression analysis

For time-course gene expression analysis, synchronized third-instar larvae were exposed to 500 mg/L GPI (the highest concentration tested in bioassays) or solvent control (1% acetone + 0.1% Triton X-100) using the leaf-dip method described in Section 2.11.

Surviving larvae were collected at 24, 48, 72, and 96 h post-treatment. At each time point, three biological replicates (10 pooled larvae each) were flash-frozen in liquid nitrogen and stored at -80°C until RNA extraction. This sampling scheme ensured sufficient RNA yield while minimizing inter-individual variability.

Quantitative real-time PCR (RT-qPCR) analysis

Total RNA was extracted from frozen samples using RNAiso Plus reagent (TaKaRa, Kusatsu, Japan). RNA concentration and purity were assessed by NanoDrop 2000c spectrophotometer (Thermo Fisher Scientific). Only samples with A₂₆₀/A₂₈₀ ratios between 1.9 and 2.1 and A₂₆₀/A₂₃₀ ratios >1.8 were used for downstream analysis. RNA integrity was confirmed by agarose gel electrophoresis (sharp 28S/18S rRNA bands, ratio ≈ 2:1). cDNA was synthesized from 1 µg RNA using PrimeScript™ RT Reagent Kit with gDNA Eraser (TaKaRa).

Gene-specific primers for PxGP, trehalase (PxTre), hexokinase (PxHex), phosphoenolpyruvate carboxykinase (PxPEPCK), PxG-6-Pase, and the reference gene PxRPS13 (ribosomal protein S13) were designed using Primer-BLAST (NCBI) and are listed in Table S1. PxRPS13 was selected based on its stable expression across developmental stages and treatment conditions (geNorm M value < 0.5) [55]. All primers were validated for specificity (single peak in melt-curve), efficiency (90-110%, R² > 0.98), and absence of primer-dimers (no-template controls).

RT-qPCR reactions were performed on a QuantStudio™ 5 Real-Time PCR System (Thermo Fisher Scientific) using TB Green® Premix Ex Taq™ II (Tli RNaseH Plus) (2×) (TaKaRa) in 20 µL reactions. The thermal cycling conditions were as follows: initial denaturation at 95°C for 30 s; 40 cycles of 95°C for 15 s (denaturation) and 60°C for 30 s (annealing/extension); followed by melt-curve analysis. Each sample was analyzed in technical triplicate, and each biological replicate consisted of three independent samples. Relative gene expression was calculated using the 2-ΔΔCt method [56].

RNA interference (RNAi)

A conserved 606 bp region of PxGP (nucleotides 66-671, part of the AMP-binding domain) was selected for RNAi. The fragment was amplified from larval cDNA using primers flanked by T7 promoters (Table S2). Double-stranded RNA (dsRNA) was synthesized with the T7 RiboMax™ Express RNAi System (Promega, Madison, WI, USA). The dsRNA of green fluorescent protein (dsGFP, 541 bp) was synthesized using a pGFP plasmid template as a non-specific negative control. Following synthesis, dsRNA was treated with RNase-free DNase I (30 min at 37°C), purified by isopropanol precipitation, and integrity verified by 1.5% agarose gel electrophoresis. Concentration and purity (A₂₆₀/A₂₈₀=1.9-2.1) was measured using a NanoDrop spectrophotometer. Purified dsRNA was resuspended in nuclease-free water at concentrations of 6,000, 10,000, or 14,000 ng/µL and stored in single-use aliquots at -80°C.

For microinjection, synchronized early third-instar larvae (6-12 h post-molt) were briefly chilled on ice for 2-3 min and injected laterally using a DMP-300 Injector (Micrology Precision Instruments, Ltd, Wuhan, China). After recovery on fresh cabbage leaves for 30 min at room temperature, larvae were maintained under standard rearing conditions. Phenotypic parameters (mortality, molting success, pupation rate, adult emergence) were monitored daily for 10 days. Larvae were collected at 24, 48, 72, and 96 h post-injection for RT-qPCR analysis to confirm target gene knockdown (n = 3 biological replicates, each containing 10 pooled larvae per time point).

Metabolite quantification

To directly assess metabolic flux alterations following PxGP knockdown, we quantified four key metabolites in the glycogen-glucose metabolic axis using commercial enzymatic assay kits: glycogen and glucose (Glycogen Assay Kit, Grace Biotechnology, Suzhou, China, Cat# G0590W96), trehalose (Trehalose Assay Kit, Grace Biotechnology, Cat# G0553W96), and glucose-6-phosphate (G6P Assay Kit, Beyotime Biotechnology, Shanghai, China, Cat# S0185). All kits utilize colorimetric or fluorometric enzymatic reactions and were performed according to the manufacturers’ protocols with minor modifications as described below.

Third-instar larvae injected with dsGP (14,000 ng/µL, the highest dose achieving maximal knockdown) or dsGFP control were collected at 48, 72, and 96 h post-injection. At each time point, three biological replicates were collected, each consisting of 10 pooled larvae. Samples were immediately flash-frozen in liquid nitrogen and stored at -80°C until analysis.

For metabolite extraction, frozen larvae were homogenized on ice in ice-cold extraction buffer specific to each metabolite assay (typically 100 µL buffer per 10 mg tissue). For glycogen extraction, samples were homogenized in deionized water and immediately boiled for 3 min to inactivate endogenous glycogenolytic enzymes. For trehalose, glucose, and G6P extraction, samples were homogenized in the respective assay buffers provided in the kits. Homogenates were centrifuged at 12,000 × g for 10 min at 4°C, and supernatants were immediately used for metabolite quantification or stored at -80°C for no more than 24 h.

Metabolite quantification was performed in 96-well microplates using Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland) at the wavelengths specified in each kit’s protocol. Sample metabolite concentrations were calculated by comparison with the absorbance of the kit’s standard. Metabolite concentrations were normalized to tissue protein content to account for variation in tissue input. Protein concentration in the crude lysates was determined using the Protein Quantification Kit (BCA Assay) (Abbkine Scientific) with bovine serum albumin as the standard. All samples were analyzed in technical triplicate. Final metabolite levels were expressed as: Glycogen: mg/mg protein; Glucose: µmol/mg protein; Trehalose: µmol/mg protein; G6P: µmo/mg protein.

Data are presented as mean ± standard error of the mean (SEM). Statistical significance between dsGP and dsGFP groups at different time point was assessed using two-way ANOVA (GraphPad Prism 8.0), with Sidak’s multiple comparisons test [57]. Statistical significance was set at P < 0.05 (*), P < 0.01 (**) or P < 0.001 (***).

eLife Assessment

This study addresses the mechanism of action of benzoylurea insecticides and explores the metabolic consequences of inhibiting glycogen breakdown in insects. Both reviewers identify major flaws with the premise of the work. The strength of the provided evidence is inadequate as the data do not, or poorly, support several central claims. The significance of the findings is considered marginal.

Reviewer #1 (Public review):

Summary:

In this study, the authors investigate whether glycogen phosphorylase is a potential molecular target of benzoylphenylurea insecticides and examine the physiological consequences of inhibiting glycogen breakdown in the diamondback moth Plutella xylostella. The authors express and characterize recombinant glycogen phosphorylase, test its inhibition by a mammalian glycogen phosphorylase inhibitor and by the insecticide diflubenzuron, and assess the physiological effects of glycogen phosphorylase inhibition through chemical exposure and RNA interference. Based on these experiments, the authors conclude that benzoylphenylurea insecticides do not target glycogen phosphorylase and propose that insects compensate for glycogen phosphorylase inhibition through activation of gluconeogenesis, allowing them to maintain glucose homeostasis and complete development despite strong suppression of the enzyme.

Strengths:

The study addresses an interesting and long-standing question in insect toxicology regarding the mechanism of action of benzoylphenylurea insecticides. The authors combine several complementary approaches, including recombinant enzyme characterization, inhibitor assays, RNA interference, gene expression analyses, and metabolite measurements. The biochemical characterization of the recombinant glycogen phosphorylase and the demonstration that the tested glycogen phosphorylase inhibitor can strongly inhibit enzyme activity represent important technical strengths. In addition, the study integrates biochemical and physiological observations to explore how insects might compensate for disruptions in central carbohydrate metabolism.

Weaknesses:

Several aspects of the central conclusions rely on indirect evidence and would benefit from additional validation. The proposed compensatory mechanism (gluconeogenesis supported by amino acid mobilization) is inferred primarily from transcriptional changes in gluconeogenic genes, reduced protein levels, and changes in metabolite concentrations. While these observations are consistent with increased gluconeogenic activity, they do not directly demonstrate metabolic flux through this pathway. Direct measurements of gluconeogenic flux would be required to confirm that carbon derived from non-carbohydrate substrates contributes to glucose production.

Some interpretations are also speculative. For example, the lack of glycogen accumulation following glycogen phosphorylase knockdown is attributed to alternative glycogen degradation pathways, such as α-amylase or glycogen debranching enzymes, but these possibilities are not experimentally examined. Measuring the expression or activity of these enzymes would help evaluate whether such pathways contribute to the observed metabolic response.

The physiological consequences of the proposed metabolic compensation are also not fully explored. If proteins are mobilized to support gluconeogenesis, this shift might be expected to affect organismal traits such as adult body size, flight capacity, or reproductive performance. Assessing these traits could provide valuable insight into whether the proposed compensatory metabolism carries fitness costs.

Finally, some conclusions extend beyond the direct evidence presented. The study shows that diflubenzuron does not inhibit glycogen phosphorylase in vitro, but broader conclusions regarding the mechanism of action of benzoylphenylurea insecticides as a class may require additional evidence. In addition, some biochemical and cell-based observations would benefit from confirmation in whole insects, given that metabolic regulation can differ substantially between isolated enzyme or cell-based systems and intact larvae, where hormonal signaling, tissue interactions, and nutrient availability influence metabolic responses.

Reviewer #2 (Public review):

(1) Significance of the findings and strength of the evidence

This manuscript evaluates the hypothesis that benzoylurea (BPU) insecticides exert their effects through inhibition of glycogen phosphorylase rather than chitin synthase (CHS). The central premise-that structural similarity among acylurea compounds implies shared molecular targets-is not supported by existing evidence.

Extensive genetic and biochemical studies, including Reference 5, demonstrate that chitin synthase is the primary insecticidal target of BPUs. In particular, amino acid substitutions at a single site in CHS confer high levels of resistance to diflubenzuron and related compounds, with causality established through CRISPR/Cas9 editing in Drosophila melanogaster. This body of evidence substantially weakens the rationale for proposing glycogen phosphorylase as an alternative primary target.

The manuscript reports that an acylurea compound previously identified as an inhibitor of mammalian glycogen phosphorylase also inhibits glycogen phosphorylase from Plutella xylostella, while diflubenzuron does not. This observation is consistent with prior work showing that glycogen phosphorylase inhibition among acylureas depends on specific side chain substitutions rather than the shared acylurea core. Consequently, the finding does not support the broader inference that acylurea structure predicts common biological function.

The manuscript further argues that inhibition of glycogen phosphorylase is not insecticidal and attributes this to metabolic compensation through alternative glucose producing pathways. While it is well established that eukaryotic cells possess multiple mechanisms for maintaining glucose availability, the evidence provided here does not fully support the broader claim that this mechanism explains the lack of insecticidal activity. In particular, the conclusion that the study "resolves" the primary hypothesis is not justified by the data presented.

Overall, while some experimental observations are sound in isolation, the overarching conclusions are not supported by the strength of the evidence. The significance of the findings is therefore limited.

(2) Interpretation in the context of existing literature

The introduction states that the molecular target of BPU insecticides remains a major unresolved controversy. However, multiple prior studies, including References 1, 4, and 5, provide strong genetic evidence that CHS is the primary and essential target of BPUs. These results demonstrate causality rather than simple correlation, particularly through targeted gene editing approaches.

The manuscript further claims that biochemical studies have failed to demonstrate CHS inhibition by BPUs in cell free assays. However, the cited references (6-9) did not express CHS in such assays and therefore do not directly address this question. As a result, the suggested discrepancy between genetic and enzymatic evidence is not well founded.

Structural analysis of acylurea compounds indicates that biological activity depends on side chain composition rather than the conserved acylurea core. Prior screening studies (Reference 11) show substantial variability in glycogen phosphorylase inhibition among acylureas despite a shared core structure. This undermines the proposal that the acylurea moiety itself constitutes a meaningful clue to a shared molecular mechanism.

Regarding implications for pesticide design, targeting chitin synthesis remains an attractive strategy because chitin is essential for arthropods and absent in mammals, providing both efficacy and specificity. By contrast, metabolic enzymes such as glycogen phosphorylase are widely conserved, making them less suitable targets from a toxicological and safety perspective.

(3) Specific technical comments

The manuscript uses the term "dataology," which is neither defined nor contextualized within the text. As currently used, the term appears unrelated to the subject matter and may be confusing to readers. Clarification or removal would improve clarity.

Author response:

The following is the authors’ response to the original reviews.

eLife Assessment

This study addresses the mechanism of action of benzoylurea insecticides and explores the metabolic consequences of inhibiting glycogen breakdown in insects. Both reviewers identify major flaws with the premise of the work. The strength of the provided evidence is inadequate as the data do not, or poorly, support several central claims. The significance of the findings is considered marginal.

The Assessment stated that “both reviewers identify major flaws with the premise of the work” and that “the strength of the provided evidence is inadequate.” We have addressed both dimensions:

(1) Premise: The Introduction has been substantially restructured to explicitly acknowledge the compelling CRISPR/Cas9 evidence establishing CHS as the primary site of BPU resistance (Reference 1). The study is now reframed as a systematic evaluation of GP as an independent insecticidal target and an investigation of metabolic compensation mechanisms — questions with scientific value independent of the BPU mechanism debate (see details in lines 47-54 of the revised manuscript).

(2) Evidence: Four new sets of experiments directly address the specific evidence gaps identified by the reviewers: (i) GP enzyme activity measurements in RNAi-treated larvae; (ii) expression analysis of alternative glycogen catabolic enzymes; (iii) molecular docking and MM/GBSA binding free energy analysis; (iv) comprehensive fitness cost assessment including feeding rate, larval weight, pupal weight, adult wing area, and female fecundity.

Public Reviews:

Reviewer #1 (Public review):

Summary:

In this study, the authors investigate whether glycogen phosphorylase is a potential molecular target of benzoylphenylurea insecticides and examine the physiological consequences of inhibiting glycogen breakdown in the diamondback moth Plutella xylostella. The authors express and characterize recombinant glycogen phosphorylase, test its inhibition by a mammalian glycogen phosphorylase inhibitor and by the insecticide diflubenzuron, and assess the physiological effects of glycogen phosphorylase inhibition through chemical exposure and RNA interference. Based on these experiments, the authors conclude that benzoylphenylurea insecticides do not target glycogen phosphorylase and propose that insects compensate for glycogen phosphorylase inhibition through activation of gluconeogenesis, allowing them to maintain glucose homeostasis and complete development despite strong suppression of the enzyme.

Strengths:

The study addresses an interesting and long-standing question in insect toxicology regarding the mechanism of action of benzoylphenylurea insecticides. The authors combine several complementary approaches, including recombinant enzyme characterization, inhibitor assays, RNA interference, gene expression analyses, and metabolite measurements. The biochemical characterization of the recombinant glycogen phosphorylase and the demonstration that the tested glycogen phosphorylase inhibitor can strongly inhibit enzyme activity represent important technical strengths. In addition, the study integrates biochemical and physiological observations to explore how insects might compensate for disruptions in central carbohydrate metabolism.

We are grateful that Reviewer 1 recognized the study's strengths, including the complementary multi-approach strategy, the biochemical characterization of recombinant PxGP, and the integration of biochemical and physiological observations.

Weaknesses:

(1) The proposed compensatory mechanism relies on indirect evidence; direct measurements of gluconeogenic flux are lacking.

We agree that isotopic tracer experiments would provide the most direct evidence for gluconeogenic flux. Such experiments are beyond the scope of the current revision, and we now explicitly acknowledge this as a key limitation and an important direction for future research (revised Discussion: Study limitations and future directions).

However, we note that the convergent evidence from multiple independent lines collectively supports gluconeogenic activation: (i) transcriptional upregulation of PEPCK and G-6-Pase; (ii) declining protein levels now independently confirmed by our new enzyme activity data showing a 30.78% decrease in total protein concentration at 24 h post-RNAi (new Figure 10A); (iii) altered amino acid profiles; and (iv) maintained trehalose levels. The revised manuscript presents this evidence more cautiously, framing it as “consistent with gluconeogenic compensation” rather than establishing metabolic flux.

Additionally, we now provide new data on GP enzyme activity (new Figure 10A, B; see response to Recommendation 4 below) and alternative glycogen catabolic enzyme expression (new Figure 10C, D; see response to Recommendation 2 below) that further strengthen the evidence chain.

(2) Alternative glycogen degradation pathways are proposed but not experimentally examined.

We have now directly addressed this concern. RT-qPCR analysis of glycogen branching enzyme (GBE) and α-amylase following PxGP knockdown reveals a striking and informative differential response (new Figure 10C, D):

GBE was significantly upregulated at 24 h (+29.24%, P < 0.05), 48 h (+16.78%, P < 0.05), and 96 h (+44.46%, P < 0.001), indicating transcriptional activation of an alternative glycogen-remodeling enzyme in response to GP suppression.

α-Amylase showed no significant change at any time point, demonstrating that the compensatory response is pathway-specific rather than a generalized upregulation of all glycogen-degrading enzymes.

This differential pattern — GBE up, α-Amylase unchanged — provides the first evidence that P. xylostella selectively activates specific glycogen remodeling pathways when GP function is compromised. Upregulation of GBE, which increases glycogen branching and solubility, may facilitate glycogen mobilization through alternative routes even when GP-mediated phosphorolysis is impaired. These data are incorporated as new Figure 10C, D and discussed in the revised Results and Discussion.

(3) Physiological consequences (fitness costs) are not explored.

We have now conducted a comprehensive fitness cost assessment (new Figure 11). The results reveal a transient but significant fitness cost confined to the larval stage:

Feeding rate: no significant difference between dsGP and dsGFP groups at any time point (24–120 h; Figure 11B), confirming that the observed metabolic changes are not attributable to reduced food intake.

Larval weight: significantly reduced at 24 h (−29.10%, P < 0.05) and 48 h (−25.38%, P < 0.05; Figure 11C), demonstrating a measurable short-term cost of metabolic compensation.

Pupal weight: no significant difference (Figure 11D), indicating full recovery before the pupal transition.

Adult wing area: no significant difference (Figure 11E, Figures S5–S6), suggesting no impairment of flight capacity.

Female fecundity (3-day egg production): no significant difference (Figure 11F), demonstrating no reduction in reproductive output.

This pattern — transient larval weight loss with complete recovery of pupal weight, wing morphology, and reproductive performance — is consistent with our proposed model: GP suppression triggers protein catabolism to fuel gluconeogenesis (explaining the short-term weight loss), but the compensatory mechanism is sufficiently effective to restore metabolic homeostasis before pupation. These data strengthen the conclusion that GP is functionally non-essential for completing development and reproduction.

(4) Broader conclusions regarding BPU class may require testing additional compounds.

We agree. The revised manuscript now explicitly limits the biochemical conclusion to diflubenzuron: “DFB does not inhibit PxGP” rather than making broader claims about the BPU class as a whole. We discuss this as a limitation and note that testing additional BPU compounds would be needed before generalizing.

(5) Some biochemical and cell-based observations would benefit from confirmation in whole insects.

We have now provided whole-insect confirmation through: (i) GP enzyme activity measurements in RNAi-treated larvae (new Figure 10A, B); (ii) in vivo fitness assessment showing measurable physiological consequences of GP suppression (new Figure 11); and (iii) expression analysis of compensatory enzymes in intact larvae (new Figure 10C, D). These data bridge the gap between our cell-free biochemical observations and whole-organism biology.

Reviewer #2 (Public review):

(1) The central premise — that structural similarity among acylurea compounds implies shared targets — is not supported.

We agree that the original manuscript overstated the significance of the shared acylurea core as a predictor of common biological activity. The Introduction has been substantially restructured to:

Explicitly acknowledge the compelling genetic evidence from CRISPR/Cas9 experiments (Reference 5) establishing CHS as the primary site conferring BPU resistance.

Reframe the study's objective: rather than proposing to “resolve” the BPU target controversy, the revised manuscript focuses on the systematic evaluation of GP as an independent insecticidal target and the discovery of a gluconeogenic compensation mechanism — questions with scientific value independent of the BPU mechanism debate.

Remove the claim that the study “resolves the primary hypothesis.” The conclusion now states that our biochemical data demonstrate DFB does not inhibit PxGP, adding enzyme-level evidence to the existing genetic framework.

(2) Target selectivity is determined by side-chain composition, not the shared acylurea core.

We fully agree, and our new structural data now provide a molecular explanation for this principle at the atomic level. Molecular docking and MM/GBSA analysis (new Figure 12, new Table 1) reveal that both GPI and DFB anchor to PxGP through their common acylurea carbonyl groups (Arg193), but diverge dramatically in side-chain engagement:

GPI's methoxyphenyl-methylurea moiety establishes extensive contacts with seven residues across both subunits (Asn44 and Val45 from chain A; Trp67, Gln71, Tyr75, Arg193, and Asp227 from chain B), binding at the allosteric site at the dimer interface — consistent with the experimentally determined binding mode of acylurea inhibitors in mammalian GP (PDB: 2ATI).

DFB contacts six residues primarily from subunit B, and its difluorobenzoyl moiety remains entirely solvent-exposed without productive protein contacts.

MM/GBSA analysis confirms GPI binds with substantially higher affinity (ΔG = −34.63 vs. −29.29 kcal/mol; ΔΔG = −5.34 kcal/mol), with van der Waals interactions as the dominant driver (ΔVDW = −11.49 kcal/mol), reflecting superior shape complementarity of GPI.

These structural data directly support Reviewer 2's important point and are now presented as new Figure 12 and Table 1.

(3) References 6–9 characterization.

We have replaced the original citations (former References 6–9) in the Introduction with references that directly demonstrate the absence of CHS inhibition by BPUs in cell-free preparations: Cohen & Casida (1980) showed DFB did not inhibit Tribolium gut chitin synthetase; Mayer et al. (1981) and Cohen (1985) systematically confirmed that BPU-type insect growth regulators do not inhibit chitin synthase in cell-free assays; and Zhang & Zhu (2013) reported only slight in vitro CHS inhibition by DFB in Anopheles gambiae with no in vivo effect. We also cite authoritative reviews by Matsumura (2010) and Merzendorfer (2013) that contextualize this evidence gap. We thank Reviewer 2 for identifying this important citation issue, which has led to a substantially more accurate and well-supported presentation of the literature.

(4) The term “dataology” is non-standard.

This term has been removed and replaced with “data.” In accordance with eLife's policy on AI tools and technology, we have added a statement in the Materials and Methods section declaring that AI-based language editing tools were used for English grammar and style refinement. All scientific content was generated entirely by the authors.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

(1) Direct assessment of gluconeogenic flux (e.g., metabolic tracer experiments).

As discussed above, isotopic tracer experiments are beyond the current scope. We acknowledge this as a key limitation in the revised Discussion (see details in lines 560-565 of the revised manuscript). However, we now provide additional supporting evidence: the 30.78% decline in total protein at 24 h post-RNAi (from our new enzyme activity data, Figure S3) provides independent biochemical confirmation of protein catabolism, consistent with amino acid mobilization for gluconeogenesis (see details in lines 357-359 of the revised manuscript). We have also added discussion of potential future approaches, including measurements of key enzymes in amino acid catabolism (e.g., aspartate aminotransferase, glutamate aminotransferase) and lipid content dynamics (see details in lines 565-569 of the revised manuscript).

(2) Expression or activity of alternative glycogen degradation enzymes (α-amylase, glycogen debranching enzymes).

We have measured the expression of GBE (glycogen branching enzyme) and α-amylase by RT-qPCR in RNAi-treated insects. We also attempted to measure glycogen debranching enzyme (GDE), but multiple primer pairs failed to yield amplification products, likely due to sequence annotation issues; this is noted as a limitation.

Results (new Figure 10C, D): GBE was significantly upregulated at 24 h (+29.24%), 48 h (+16.78%), and 96 h (+44.46%). α-Amylase was unchanged at all time points (see details in lines 377-381 of the revised manuscript). The selective upregulation of GBE but not α-amylase suggests a targeted compensatory response within the glycogen remodeling pathway.

The absence of glycogen accumulation following GP knockdown may reflect reduced flux into glycogen synthesis (potentially through feedback inhibition of glycogen synthase) rather than activation of alternative degradative routes. This possibility is discussed in the revised manuscript, with glycogen synthase expression identified as a key target for future investigation (see details in lines 501-505 of the revised manuscript).

(3) Fitness cost assessment (body size, flight capacity, reproductive performance).

Complete data are now provided (new Figure 11A–F, Figures S5–S6):

The transient larval weight reduction (24–48 h) with complete pupal and adult recovery demonstrates that metabolic compensation carries a short-term physiological cost but is ultimately effective in maintaining developmental trajectory and reproductive fitness (see details in lines 388-408 of the revised manuscript).

(4) Enzyme activity measurements in RNAi-treated insects.

GP enzyme activity (GP-a) was measured in crude extracts from RNAi-treated larvae using a coupled-enzyme spectrophotometric assay kit (Solarbio BC3345) at 24, 48, 72, and 96 h post-injection (new Figure 10A, B).

Two normalization approaches were used: - Per-protein activity: significant reduction only at 48 h (−10.35%, P < 0.05; Figure 10A). The modest per-protein reduction reflects a concurrent 30.78% decline in total protein at 24 h, which inflates per-protein specific activity when the protein pool shrinks. - Per-larva activity: significant reduction at 24 h (−27.57%, P < 0.05) and 48 h (−29.28%, P < 0.01; Figure 10B), confirming that RNAi-mediated transcript suppression translates to reduced enzyme function in vivo.

The 30.78% decline in total protein at 24 h provides independent biochemical confirmation of protein catabolism — consistent with amino acid mobilization for gluconeogenesis (see details in lines 349-370 of the revised manuscript).

(5) Scope of BPU conclusion — clarify whether additional compounds should be tested.

The revised manuscript explicitly states that the biochemical conclusion applies to diflubenzuron specifically. We have added a Discussion paragraph noting that extending this conclusion to additional BPU compounds would require systematic testing, and that our structural analysis (Table 1) provides a framework for predicting which acyl urea side-chain architectures are compatible with GP binding (see details in lines 455-466 of the revised manuscript).

(6) RNAi transcript recovery at 96 h and implications for non-essentiality.

Our new enzyme activity data directly address this concern. GP activity (per-larva) showed partial recovery at 72 h and 96 h (Figure 10B), mirroring the transcript recovery pattern. However, the critical observation is that even during the period of maximum suppression (24–48 h), when per-larva GP activity was reduced by ~27–30%, larvae maintained glucose homeostasis and completed development. This confirms that GP is non-essential even during the period of strongest suppression. The revised Discussion addresses this point explicitly (see details in lines 479-488 of the revised manuscript).

(7) GPI concentrations in larval exposure experiments and pharmacokinetic considerations.

We have added a dedicated Discussion paragraph addressing this concern in detail. The GPI concentrations used (250–500 mg/L in diet) encompass a wide dose range; even at the highest concentration (500 mg/L, approximately 409,000-fold above the in vitro IC50 of 2.96 nM), no toxicity was observed. We discuss several pharmacokinetic factors that may contribute to this apparent discrepancy, including limited oral bioavailability, metabolic inactivation by detoxification enzymes, and sequestration by hemolymph binding proteins. However, we note that the metabolic phenotype observed (elevated trehalose, reduced protein, upregulated gluconeogenic enzymes) provides indirect evidence that GPI does reach its target. The most parsimonious interpretation is that GPI achieves sufficient target engagement to partially suppress GP activity in vivo, but metabolic compensation renders this suppression non-lethal — an interpretation reinforced by our RNAi data, in which direct genetic suppression of GP (bypassing all pharmacokinetic barriers) similarly fails to cause mortality (see details in lines 536-552 of the revised manuscript).

(8) Structural evidence that GPI binds PxGP comparably to its mammalian target.

This has been comprehensively addressed through molecular docking and MM/GBSA analysis (new Figure 12, new Table 1). The PxGP homodimer structure was modeled using SWISS-MODEL with the human liver GP–acyl urea co-crystal structure (PDB: 2ATI) as the template. Docking and binding free energy calculations were performed in Cresset Flare V11.

Key findings: GPI binds at the allosteric site at the dimer interface with ΔG = −34.63 kcal/mol, engaging seven residues across both subunits — a binding mode consistent with the experimentally determined site in mammalian GP. DFB binds with lower affinity (ΔG = −29.29 kcal/mol) and its difluorobenzoyl moiety is entirely solvent-exposed. Van der Waals interactions are the dominant driver of selectivity (ΔVDW = −11.49 kcal/mol). See new Figure 12 and Table 1 for complete data.

(9) Dietary carbohydrate compensation and feeding behavior.

Our new data directly address this concern. Feeding rate measurements show no significant difference between dsGP and dsGFP groups at any time point (24–120 h; Figure 11B), confirming that the metabolic changes are not attributable to altered food intake. A Discussion paragraph has been added acknowledging the potential contribution of dietary carbohydrates to glucose homeostasis and noting that starvation-challenge experiments would provide additional insight (see details in lines 570-581 of the revised manuscript).

Minor comments and suggestions

(1) Terminology.

“Gluconeogenolysis” has been replaced with “gluconeogenesis” throughout the manuscript.

(2) Typographical errors.

A thorough language revision has been performed. “Over over four decades” and other errors have been corrected. The term “dataology” has been removed.

(3) Metabolite normalization.

We now present GP enzyme activity using two normalization approaches (per-protein and per-larva; Figure 10A, B) and discuss the implications of protein level changes on per-protein normalization in the Results section. For metabolite data, we have added a note in the Methods explaining our normalization approach and discussing how declining protein levels may influence interpretation (see details in lines 908-920 of the revised manuscript).

(4) Clarity of pathway descriptions.

The description of metabolic pathways has been simplified and a revised schematic figure has been included (Figure 13).

(5) Figure clarity.

Figures have been added with clearer labeling and simplified schematics. To improve clarity, we used red arrows to show the blocked metabolic flow when GP is inhibited, and green arrows to depict the activated gluconeogenic pathway (Figure 13).

Cohen E, Casida JE. Inhibition of Tribolium gut chitin synthetase. Pestic Biochem Physiol. 1980;13(2):129-36. doi: 10.1016/0048-3575(80)90064-4.

Mayer RT, Chen AC, DeLoach JR. Chitin synthesis inhibiting insect growth regulators do not inhibit chitin synthase. Experientia. 1981;37(4):337-8. doi: 10.1007/BF01959848.

Cohen E. Chitin synthetase activity and inhibition in different insect microsomal preparations. Experientia. 1985;41(4):470-2. doi: 10.1007/BF01966152.

Zhang X, Yan Zhu K. Biochemical characterization of chitin synthase activity and inhibition in the African malaria mosquito, Anopheles gambiae. Insect Sci. 2013;20(2):158-66. doi: 10.1111/j.1744-7917.2012.01568.x.

Matsumura F. Studies on the action mechanism of benzoylurea insecticides to inhibit the process of chitin synthesis in insects: A review on the status of research activities in the past, the present and the future prospects. Pestic Biochem Physiol. 2010;97(2):133-9. doi: 10.1016/j.pestbp.2009.10.001.

Merzendorfer H. Chitin synthesis inhibitors: old molecules and new developments. Insect Sci. 2013;20(2):121-38. doi: 10.1111/j.1744-7917.2012.01535.x

Preiss J. Bacterial glycogen synthesis and its regulation. Annual review of microbiology. 1984;38:419-58. doi: 10.1146/annurev.mi.38.100184.002223.

Janeček Š, Svensson B, MacGregor EA. α-Amylase: an enzyme specificity found in various families of glycoside hydrolases. Cell Mol Life Sci. 2014;71(7):1149-70. doi: 10.1007/s00018-013-1388-z.

Waterhouse A, Bertoni M, Bienert S, Studer G, Tauriello G, Gumienny R, et al. SWISS-MODEL: homology modelling of protein structures and complexes. Nucleic Acids Res. 2018;46(W1):W296-W303. doi: 10.1093/nar/gky427.

İnak E, De Rouck S, Van Leeuwen T. Molecular mechanisms of pesticide selectivity: Insights from acaricide toxicology. Pestic Biochem Physiol. 2025;213:106537. doi: 10.1016/j.pestbp.2025.106537.

David MD. Insecticide ADME for support of early-phase discovery: combining classical and modern techniques. Pest Manage Sci. 2017;73(4):692-9. doi: 10.1002/ps.4345.

Haunerland NH, Bowers WS. Binding of insecticides to lipophorin and arylphorin, two hemolymph proteins of Heliothis zea. Arch Insect Biochem Physiol. 1986;3(1):87-96. doi: 10.1002/arch.940030110.

Other revisions

Correction of primer sequences. Upon re‑checking the primer sequences during revision, we noticed that the originally reported primers for dsRNA synthesis of the GP gene (Table S1) were inadvertently copied incorrectly. The correct sequences have now been substituted in the revised manuscript (dsPxGP-F, dsPxGP-R). This correction does not affect any of the experimental data, results, or conclusions of the study. We apologize for the oversight.

Version 2

Abstract

Glycogen phosphorylase (GP) catalyzes the rate-limiting step of glycogenolysis and occupies a central position in insect carbohydrate metabolism, supplying precursors for chitin biosynthesis. Acyl urea compounds structurally related to benzoylphenylurea (BPU) insecticides are potent inhibitors of mammalian GP, raising the question of whether insect GP could serve as an independent insecticidal target. Here, we systematically evaluate this possibility in the diamondback moth Plutella xylostella. We show that a mammalian GP inhibitor (GPI) potently inhibits recombinant PxGP (IC50 = 2.96 nM), while the BPU insecticide diflubenzuron (DFB) does not. Molecular docking and MM/GBSA analysis reveal that this selectivity reflects differential side-chain engagement: GPI binds the allosteric site at the dimer interface via seven cross-subunit contacts (ΔG = −34.63 kcal/mol), whereas DFB’s difluorobenzoyl moiety fails to establish productive protein contacts (ΔG = −29.29 kcal/mol). Despite potent in vitro inhibition, GPI exhibits no insecticidal activity, and RNAi-mediated knockdown of PxGP (confirmed by enzyme activity measurements showing ∼27-30% reduction in per-larva GP-a activity) does not impair development or survival. We demonstrate that insects compensate through a multi-layered metabolic response: upregulation of gluconeogenic enzymes (PEPCK, G-6-Pase), selective activation of glycogen branching enzyme (GBE) but not α-amylase, and protein catabolism providing gluconeogenic substrates. Fitness assessment reveals transient larval weight loss (24–48 h) with complete recovery of pupal weight, adult wing morphology, and female fecundity, confirming that metabolic compensation is ultimately effective. These findings establish GP as functionally non-essential due to gluconeogenic rescue and highlight the importance of considering metabolic network redundancy in target-based insecticide design.

Introduction

Benzoylphenylurea (BPU) insecticides, including diflubenzuron (DFB), have been widely used for pest control for over four decades. The molecular basis of BPU-mediated insecticidal activity has been the subject of extensive investigation. Genetic studies, most notably the demonstration through CRISPR/Cas9 gene editing that amino acid substitutions at a single site in chitin synthase (CHS) confer high-level resistance to DFB in Drosophila melanogaster [1], provide compelling evidence that CHS is the primary target gene associated with BPU resistance. These findings establish a clear causal genetic link between CHS and BPU activity.

Despite this genetic evidence, the biochemical mechanism by which BPUs inhibit CHS remains incompletely characterized. Multiple in vitro studies of chitin synthase, including those using cell-free chitin synthase preparations, have failed to demonstrate direct enzymatic inhibition by BPU compounds: DFB did not inhibit Tribolium gut chitin synthetase [2], and a systematic evaluation concluded that chitin synthesis-inhibiting insect growth regulators do not inhibit chitin synthase in cell-free assays [3, 4]. More recently, Zhang and Zhu (2013) reported only slight in vitro inhibition of Anopheles gambiae CHS activity by DFB at the highest concentration tested, with no in vivo inhibition detected [5]. These findings led to the recognition that a key obstacle in identifying the action site of BPU insecticides has been the lack of in vitro demonstration of their ability to inhibit insect chitin synthesis under cell-free conditions. This gap between genetic and biochemical evidence does not diminish the importance of the genetic findings, but it does highlight the need for continued investigation at the enzyme level [6, 7].

Glycogen Phosphorylase (GP) catalyzes the rate-limiting step in glycogenolysis and plays a central role in insect carbohydrate metabolism, providing glucose-1-phosphate that feeds into the hexosamine pathway and ultimately supplies UDP-N-acetylglucosamine for chitin biosynthesis [8, 9]. In insects, GP-derived glucose is converted to trehalose (the primary hemolymph sugar), which is subsequently channeled through a series of enzymatic steps to generate UDP-GlcNAc, the substrate monomer for chitin synthesis. GP also plays key roles in carbohydrate homeostasis, osmotic regulation, stress adaptation [10–13]. Acyl urea compounds structurally related to BPUs have been identified as potent inhibitors of mammalian GP (Figure 1) [14, 15], raising the question of whether insect GP could itself represent a viable insecticidal target — independent of the BPU mechanism-of-action question. However, the feasibility of targeting insect GP for pest control has not been systematically evaluated.

Structural comparison of benzoylphenylurea (BPU) insecticides and a mammalian glycogen phosphorylase inhibitor (GPI) reveals a shared acylurea scaffold.

Chemical structures of four BPU insecticides (Diflubenzuron, Teflubenzuron, Novaluron, and Flufenoxuron) and a human glycogen phosphorylase inhibitor (GPI) are shown. All share a central N,N’-diphenylurea (acylurea) core (dashed boxes). BPUs are characterized by halogenated (F, Cl) aromatic substituents and in some cases by additional ether or alkyl side chains. The GPI contains a methoxyphenyl group bearing an additional methylurea side chain.

In this study, we address two distinct questions. First, we test whether DFB directly inhibits Plutella xylostella GP (PxGP) at the enzyme level, adding biochemical data to the existing genetic framework. Second, and more importantly, we systematically evaluate GP as an independent insecticidal target by characterizing the metabolic consequences of GP inhibition and knockdown in P. xylostella. Our findings reveal that PxGP is functionally non-essential for larval development due to a previously uncharacterized gluconeogenic compensation mechanism, providing new insight into the metabolic plasticity of insects and its implications for target-based insecticide design.

Results

Recombinant P. xylostella glycogen phosphorylase (PxGP) is a functional, activatable enzyme

To investigate its biochemical properties and evaluate its potential as an insecticidal target, the full-length PxGP cDNA (predicted molecular mass of 98.1 kDa) was cloned and expressed in Sf9 cells using the Bac-to-Bac baculovirus expression system. Following purification via immobilized metal-affinity chromatography (IMAC) of the N-terminal 6×His tagged protein [16], SDS-PAGE analysis and Western blot analysis confirmed a highly pure protein band migrating at approximately 100 kDa, consistent with its theoretical mass (Figure 2A). A total yield of 7.5 mg of purified protein was obtained from 8.25 g of cell pellet (specific yield=0.91 mg/g cells).

Recombinant PxGP is inhibited by a mammalian GP inhibitor (GPI) but not by the benzoylphenylure Diflubenzuron.

(A) Purification of 6×His-tagged PxGP from Sf9 cells. Left panel: Coomassie-stained SDS-PAGE.. Right: Western blot with anti-His antibody. The expected ∼100 kDa band is indicated (red box). M, marker; S, soluble fraction; P, pellet; F, flow-through; lanes 1–4, imidazole elution fractions. (B) Dose-response inhibition by GPI. Activated PxGP-a activity was measured after pre-incubation with GPI (0.16–500 nM). Data (mean ± SEM, n=3) were fit to a four-parameter logistic curve. The calculated IC50 is 2.96 nM. Inset: GPI structure with its characteristic asymmetric side chains (orange). (C) Inhibition assay with DFB (0.54–1700 nM). DFB did not yield a complete sigmoidal inhibition curve, causing ≤62% inhibition at the highest concentration. Inset: DFB structure with its characteristic halogenated benzoyl groups (orange). This rules out PxGP as a direct target of DFB.

Glycogen phosphorylase activity is regulated by reversible phosphorylation, existing as a less-active dephosphorylated b form (PxGP-b) and a fully active phosphorylated a form (PxGP-a) [17]. Recombinant PxGP-b was activated by incubation with immobilized phosphorylase kinase (PhK) in the presence of ATP and Mg2+ for 2 h [18]. The activated enzyme showed robust catalytic activity in a coupled-enzyme spectrophotometric assay, with NADPH production (monitored by increase in A340) increasing linearly for the first 15 min and reaching a plateau by approximately 20 min, consistent with substrate depletion and/or product accumulation. The specific activity of activated PxGP-a was 294.25 ± 3.45 U/mg protein (Figure S1A), confirming that the recombinant enzyme is catalytically competent for inhibitor screening.

GPI potently inhibits PxGP in vitro and in vivo

Given the structural similarity between insecticidal BPUs and mammalian GPIs [15] (Figure 1), we tested whether PxGP could be a target of acylurea-scaffold compounds. GPI exhibited potent, dose-dependent inhibition of recombinant PxGP-a, with a sigmoidal dose-response curve typical of competitive or allosteric inhibition (Figure 2B). Non-linear regression analysis yielded an IC50 value of 2.96 nM (95% confidence interval [CI]: 1.66–5.35 nM, Hill slope = 0.74, R² = 0.877, n = 3). Near-complete inhibition (>95%) was achieved at 500 nM GPI, indicating saturable, high-affinity binding. The potency of GPI against PxGP is comparable to its effect on mammalian orthologs (see below) [15].

In striking contrast, the BPU insecticide Diflubenzuron (DFB), which shares the acylurea core with GPI (Figure 1), did not effectively inhibit PxGP-a. No standard sigmoidal dose-response curve was observed over 0.54 nM to 1700 nM (Figure 2C). Maximum inhibition observed did not exceed 61.96 ± 2.54% (mean ± SEM) even at 100 µM, with high variability at lower concentrations. These data indicate that DFB, unlike GPI, is not an effective inhibitor of PxGP, ruling out GP as the direct molecular target of this BPU insecticide.

Consistent with the recombinant enzyme results, GPI effectively inhibited native GP activity in crude lysates from third-instar larvae in a dose- and time-dependent manner (Figure S1B). After 45 min of pre-incubation, 10 µM GPI reduced native GP activity by 57.52 ± 1.88% (n = 3). DFB, however, showed no significant inhibitory effect on native GP activity at any concentration tested (10-7, 10-6, or 10-5 M) or time point (5, 10, 20, or 30 min), with activity remaining at 97-108% of control (Figure S1C).

Collectively, these results establish PxGP as a high-affinity molecular target for GPI, whereas DFB does not inhibit PxGP.

Paradoxical finding: potent PxGP inhibition by GPI lacks in vivo toxicity

Because GPI potently inhibit PxGP, the rate-limiting enzyme in glycogenolysis, which supplies precursors for chitin synthesis [19], we expected it to exhibit significant insecticidal activity. In leaf-dip bioassays with third-instar P. xylostella larvae continuously exposed to GPI (250 or 500 mg/L) or DFB (125 or 250 mg/L), GPI caused no significant toxicity or developmental disruption across the entire time course (24 to 120 h post-treatment) (Figure 3A). At 120 h, mortality with 500 mg/L GPI was only 12.78 ± 2.42% (n = 3), which was not significantly different from the solvent control (CK: 5.00 ± 5.00%, P = 0.21) (Figure 3B). Larvae exposed to 250 mg/L GPI exhibited even lower mortality (2.50 ± 2.50%).

The GP inhibitor (GPI) is non-toxic to Plutella xylostella, unlike the benzoylphenylurea Diflubenzuron (DFB).

(A) Developmental stage distribution over time for control, GPI (250, 500 mg/L), and DFB (125, 250 mg/L) treated larvae. DFB causes developmental arrest. (B) Quantitative effects on mortality (top), pupation (middle), and adult emergence (bottom). GPI groups did not differ from control (P > 0.05), whereas DFB caused significant, dose-dependent mortality and severely impaired development (*P < 0.05). Data are mean ± SEM (n=3). Different letters denote statistical significance. (C) Larval phenotypes at 96 h. Control and GPI-treated larvae are normal. DFB-treated larvae show catastrophic molting failure and “double cuticle” (red arrows). Scale bar, 1 mm. (D) Pupal and adult phenotypes. GPI-treated individuals develop normally. DFB treatment results in deformed pupae and trapped pharate adults (red arrows). Scale bar, 1 mm.

In stark contrast, DFB induced significant, dose-dependent mortality. At 120 h, DFB treatment at 125 mg/L caused 41.60 ± 4.85% mortality (P < 0.01 vs. control) and at 250 mg/L caused 61.36 ± 10.02% mortality (P < 0.01 vs. control). Probit analysis of the DFB dose-response data yielded calculated LC30 and LC50 values of 141.900 mg/L (95% CI: 67.893–206.576 mg/L) and 229.537 mg/L (95% CI: 145.287-331.784 mg/L), respectively, with an acceptable goodness-of-fit (χ² = 11.946, P = 0.683) (Table S2).

Furthermore, GPI-treated larvae showed no developmental abnormalities. Pupation rates were 64.17 ± 4.79% for the 500 mg/L GPI group and 69.72 ± 6.81% for the 250 mg/L group, which were statistically indistinguishable from the control group (64.72 ± 9.33%, P > 0.05 for both comparisons) (Figure 3B, middle panel). Similarly, adult emergence rates were 56.11 ± 5.64% and 56.39 ± 6.24% for the 500 and 250 mg/L GPI groups, respectively, compared to 54.10 ± 11.91 % for the control (P > 0.05, Figure 3B). Phenotypically, GPI-treated larvae molted normally, fed actively, and displayed no cuticle defects (Figure 3C). DFB-treated larvae exhibited the characteristic phenotypes of BPU intoxication: catastrophic molting failure, larval-pupal intermediates, and complete failure of adult emergence (Figure 3C and 3D). This striking discrepancy between biochemical potency and in vivo inefficiency presents a profound paradox that demands mechanistic investigation.

GPI exposure triggers a compensatory upregulation of PxGP transcription

To explore this discrepancy, we measured PxGP gene expression after GPI exposure. RT-qPCR revealed that 500 mg/L GPI rapidly upregulated PxGP mRNA transcription. As shown in Figure 4, its relative expression increased 3.24-fold at 24 h, peaked at 3.48-fold at 48 h, and remained elevated at 96 h. This indicates that the insect responds to the pharmacological inhibition by transcriptionally increasing the de novo synthesis of the target enzyme.

Ingestion of GP inhibitor (GPI) induces compensatory upregulation of PxGP transcription.

Relative PxGP mRNA levels in larvae after continuous exposure to 500 mg/L GPI or control (CK). Expression was normalized to PxRPS13 (control set to 1.0 at each time point) using the 2-ΔΔCt method. Data are mean ± SEM (n=3). GPI triggered a rapid and sustained transcriptional increase, peaking at 48 h (3.48-fold). Significance vs. control: *P < 0.05, **P < 0.01, ***P < 0.001 (two-way ANOVA with Sidak’s test).

Sequence homology and rationale for target site selection

To achieve effective gene knockdown coupled with functional ablation, precise selection of the interference target site is critical. Sequence analysis of PxGP (837 residues) revealed that residues 88-680 form the active site pocket of the GT35_Glycogen_Phosphorylase family, and the C-terminal region contains a phosphatase-pyridoxal phosphate linkage site (residues 672-684). Eleven residues within positions 42-318 are conserved and correspond to the AMP-binding site in human GP (Figure 5A) [20].

PxGP sequence conservation and domain architecture.

(A) Schematic of PxGP functional domains: pyridoxal phosphate (PLP) binding site (blue boxes), AMP allosteric site (yellow boxes), and glycogen-binding/active site pocket (blue shading). (B) Multiple sequence alignment of GP from Plutella xylostella (PxGP), other lepidopterans (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP). Identical residues are shaded red. The gray box indicates the 606 bp region (aa 22–224) targeted for RNAi, which overlaps the conserved AMP-binding domain. PxGP shares 82.29% identity with human GP.

Multiple sequence alignment of PxGP with GP orthologs from other key lepidopteran pests (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP) demonstrated high conservation across species (Figure 5B). PxGP shares 82.29% sequence identity with human GP (Figure S2). Since GP activity is allosterically regulated by AMP binding and phosphorylation at Ser14, triggering a conformational shift from the inactive T-state GPb (unphosphorylated) to the active R-state GP-a (phosphorylated) [17], we targeted the core AMP-binding site region (gray box in Figure 5B) for RNAi. Theoretically, interference in this region should prevent AMP-mediated activation of existing GPb while degrading the entire mRNA pool, drastically reducing total GP protein synthesis and ensuring complete functional ablation.

RNAi confirms PxGP is non-essential for larval development

The compensatory upregulation of PxGP mRNA expression in response to pharmacological inhibition (Section 3.4) indicated that the gene is under active transcriptional regulation and suggested that increased enzyme synthesis might partially mitigate the effects of GPI. However, this compensation alone could not fully explain the complete absence of phenotype. To determine whether a more profound loss-of-function—achieved by directly suppressing PxGP gene expression—is lethal, we performed RNAi to knock down PxGP at the transcriptional level.

Firstly, we quantified the expression profile of PxGP across developmental stages in P. xylostella, PxGP was highly expressed in fourth-instar larvae (L4) and adults (Figure 6), stages that correlate with the energy demands for metamorphosis and flight, respectively [8, 21]. Given the inherent time lag required for systemic dsRNA delivery to achieve effective post-transcriptional knockdown [22], dsRNA was administered during the mid-third-instar larval stage. Mid-third-instar larvae were microinjected with dsRNA targeting a conserved 606 bp region of PxGP (dsGP) or with a non-specific control dsGFP. Three concentrations of dsGP (6,000, 10,000, and 14,000 ng/µL) were tested, and knockdown efficiency was assessed by RT-qPCR at 24, 48, 72, and 96 h post-injection (Figure 7). The highest concentration (14,000 ng/µL) achieved maximal suppression at 48 h, reducing PxGP transcript levels to 12.47 ± 1.68% of the dsGFP control (87.59% knockdown, P < 0.001) (Figure 7B). Significant knockdown persisted at 72 h (69.86% suppression,P < 0.05), with partial recover by 96 h, likely due to dsRNA degradation and/or compensatory transcriptional upregulation. The 10,000 ng/µL dose also produced significant knockdown at 48 h (65.76% suppression, P < 0.01), whereas the 6,000 ng/µL dose showed more modest effects (Figure 7B).

Developmental expression profile of PxGP.

Relative PxGP mRNA levels across life stages (Egg, larval instars L1-L4, Pupa, Adult), normalized to PxRPS13 (L1 set to 1.0). Data are mean ± SEM (n=3). Expression peaks in adults. Different letters indicate significant differences (P < 0.05, one-way ANOVA with Tukey’s test).

RNAi-mediated knockdown of PxGP is dose- and time-dependent.

(A) Experimental workflow. Third-instar larvae were injected with dsRNA targeting PxGP (dsGP) at three concentrations or a control dsRNA (dsGFP), then sampled for analysis. (B) Time-course of PxGP mRNA levels post-injection, relative to dsGFP control (set to 1.0). The highest dsGP dose (14,000 ng/µL) caused maximal suppression (87.6%) at 48 h. Data are mean ± SEM (n=3). Significance vs. dsGFP at each time point: **P < 0.01, ***P < 0.001 (one-way ANOVA with Tukey’s test).

Despite this robust, transient silencing, no adverse phenotypic effects were observed. Developmental stage distribution and mortality rates in all dsGP-treated groups (6,000, 10,000, and 14,000 ng/µL) were statistically indistinguishable from the dsGFP control group across the entire 120 h observation period (P > 0.05 at all time points) (Figure 8A and 8B). At 120 h, cumulative mortality was 8.10 ± 4.23%, 9.82 ± 0.81%, and 3.33 ± 3.33% for the 6,000, 10,000, and 14,000 ng/µL dsGP groups, respectively, comparable to the dsGFP control (Figure 8B). Pupation and adult emergence also indistinguishable from controls (Figure 8B). Phenotypic examination revealed that dsGP-injected larvae developed normally, molted successfully, formed morphologically normal pupae, and emerged as viable adults without cuticle defects or developmental delays (Figure 8C and 8D). Thus, acute, profound suppression of PxGP expression is not lethal to P. xylostella during the critical larval-pupal transition, despite the enzyme’s established role in glycogenolysis and chitin precursor synthesis. This finding necessitates an investigation into the compensatory mechanisms that allow the insect to survive in the absence of functional GP.

PxGP knockdown does not impair development or survival.

(A) Developmental stage distribution post-injection shows similar progression across all groups. (B) Cumulative mortality (top), pupation (middle), and adult emergence (bottom) rates were not significantly different between dsGP and dsGFP control groups (P > 0.05). Data are mean ± SEM (n=3). (C) Representative larval phenotypes at 48 h and 96 h post-injection are normal in both dsGFP and high-dose dsGP groups. Scale bar, 1 mm. (D) Pupae and adults from both groups are morphologically normal. Scale bar, 1 mm.

PxGP knockdown coordinately suppresses the glycogenolysis-to-chitin pathway

To elucidate how larvae survive GP loss, we examined downstream genes linking glycogenolysis to chitin synthesis. Specifically, we measured the expression of trehalase (PxTre), which hydrolyzes circulating trehalose to glucose [8], and hexokinase (PxHex), which phosphorylates glucose to glucose-6-phosphate—a key branch point leading to UDP-GlcNAc synthesis for chitin production [23, 24].

RT-qPCR analysis revealed that suppression of PxGP triggered a coordinated transcriptional downregulation of both downstream genes (Figure 9A and 9B). At 48 h post-injection, coinciding with the peak of PxGP knockdown (87.59% suppression), PxTre expression was reduced to 30.46 ± 1.25% of the dsGFP control (69.54% suppression, P < 0.001,), and PxHex to 22.82 ± 3.39% of control (77.18% suppression, P < 0.001). This suppression persisted at 72 h (PxTre: 61.27 ± 2.21% suppression, P < 0.001; PxHex: 41.96 ±4.02% suppression, P < 0.001) and returned to near baseline levels by 96 h, mirroring the recovery kinetics of PxGP itself.

PxGP knockdown triggers biphasic metabolic reprogramming, culminating in gluconeogenic compensation.

(A, B) Concurrent suppression of downstream genes Trehalase (PxTre) and Hexokinase (PxHex) post-knockdown. (C) Upregulation of gluconeogenic genes PEPCK and G6Pase at 96 h. Data in A-C are mean ± SEM (n=3); *P < 0.05, **P < 0.01, ***P < 0.001 vs. dsGFP control. (D-H) Metabolite levels post-knockdown. (D) Total protein decreased at 72 h (P < 0.01). (E) Glycogen remained stable. (F) Trehalose surged 7.4-fold at 96 h (P < 0.001). (G) Glucose-6-phosphate (G6P) increased 6.5-fold at 96 h (P < 0.001). (H) Free glucose showed a transient dip at 72 h (P < 0.05) before recovery. Data are mean ± SEM (n=3); *P < 0.05, ***P < 0.001 vs. control at each time point.

This coordinated downregulation indicates that the insect responds to the loss of GP not by futilely attempting to maintain flux through a pathway that lacks its key upstream enzyme, but rather by implementing a broader transcriptional shutdown of the glycogenolysis-to-glucose-to-chitin pathway. This suggests a coherent metabolic response program, raising the question: if glycogenolysis is blocked, how does the insect obtain the glucose required for survival and chitin synthesis?

Biphasic metabolic response: transient depletion followed by gluconeogenic rescue

The absence of a lethal or developmental phenotype despite glycogenolysis pathway suppression suggested activation of alternative glucose synthesis pathway to compensate for GP loss. We therefore measured expression of key gluconeogenic enzymes and metabolite levels across the glycogeno-glucose axis.

Gene expression analysis reveals gluconeogenic activation

We assessed transcript levels of two rate-limiting gluconeogenic enzymes: phosphoenolpyruvate carboxykinase (PEPCK, encoded by PxPEPCK), which catalyzes the conversion of oxaloacetate to phosphoenolpyruvate; glucose-6-phosphatase (G-6-Pase, encoded by PxG-6-Pase), which catalyzes the dephosphorylation of glucose-6-phosphate to free glucose. These enzymes are reciprocally regulated relative to glycolysis/glycogenolysis and serve as molecular markers of gluconeogenic flux [25, 26].

At 96 h post-injection―a critical pre-pupal stage with high carbohydrate demand for pupal cuticle synthesis [27, 28]―both PxPEPCK and PxG-6-Pase were significantly upregulated in dsGP-treated larvae (3.95 ± 0.44-fold and 3.34 ± 0.20-fold versus dsGFP control, respectively; P < 0.001 for each) (Figure 9C and 9D). This indicates robust transcriptional activation of the gluconeogenesis pathway in response to GP loss.

Protein catabolism provides gluconeogenic substrate mobilization

To identify the carbon source for gluconeogenesis, we quantified total protein content in dsGP-treated larvae across the time course. Strikingly, a significant 29.39% reduction in protein was observed at 72 h post-injection (P < 0.01), coinciding precisely with the nadir of glucose insufficiency (Figure 9D).

This protein decline provides direct biochemical evidence for substrate mobilization: the degradation of cellular proteins liberates amino acids, which serve as the primary carbon source for gluconeogenesis in insects [25, 29–31]. The magnitude of protein loss (∼30%) is sufficient to provide substantial amino acid flux into the gluconeogenic pathway.

Metabolite profiling reveals a biphasic adaptation

Quantification of four key metabolites (glycogen, trehalose, glucose-6-phosphat [G6P], and free glucose) along the glycogen-glucose axis at 48, 72, and 96 h post-injection uncovered a dynamic, biphasic response (Figure 9E-9H).

Glycogen Content: Contrary to the expectation of accumulation upon GP blockade, glycogen levels remained unchanged relative to controls throughout the time course (P > 0.05) (Figure 9E). This paradox indicates that insects activate alternative, GP-independent pathways for glycogen catabolism―most likely via α-amylase or glycogen debranching enzymes [32–34]―to provide emergency glucose mobilization during the period of metabolic stress, suggesting that these compensatory pathways are sufficient for short-term energy needs but are eventually superseded by gluconeogenic glucose production.

Trehalose: Trehalose levels remained stable at 48-72 h but surged dramatically by 96 h (7.44-fold elevation, P < 0.001) (Figure 9F and 9G). This massive accumulation, coinciding precisely with the peak gluconeogenic gene (e.g., PEPCK and G-6-Pase) expression, indicates that newly synthesized glucose derived from gluconeogenesis is being actively channeled into trehalose synthesis. The magnitude of this accumulation indicates that gluconeogenic glucose production exceeds the immediate metabolic demands of the larva, creating a metabolic “overshoot” that ensures adequate carbohydrate reserves for the upcoming pupation [35].

G6P: G6P, which sits at the intersection of glycolysis, gluconeogenesis, the pentose phosphate pathway, and glycogen synthesis [36], exhibited a pattern mirroring that of trehalose (Figure 9G). G6P content declined gradually from 48 to 72 h, consistent with reduced glucose mobilization from glycogen, but then increased sharply at 96 h (6.49-fold elevation, P < 0.001). This surge directly reflects the massive influx of glucose generated via the gluconeogenic pathway, which enters the metabolic network at the G6P node. The coordination between G6P accumulation and trehalose accumulation confirms that gluconeogenesis-derived glucose is being efficiently converted into storage and transport carbohydrates.

Free Glucose: Free glucose levels showed the most subtle but perhaps most physiologically informative pattern (Figure 9H). At 48 h―coinciding with peak PxGP suppression―glucose content was statistically indistinguishable from controls (P > 0.05), indicating that initial homeostatic mechanisms (including GP-independent glycogen breakdown and dietary carbohydrate absorption) successfully buffered against acute GP loss. However, at 72 h, the nadir of the metabolic stress response, free glucose levels declined significantly, albeit modestly (18.15% reduction, P < 0.05). This transient dip reveals a critical window of metabolic insufficiency during which the insect’s compensatory mechanisms are insufficient to fully maintain glucose homeostasis. Importantly, by 96 h, glucose levels fully recovered to control values (P >0.05), demonstrating that the gluconeogenic pathway successfully restores glucose homeostasis before critical developmental transitions occur.

Integrated interpretation

Collectively, these metabolite and gene expression data reveal a sophisticated, biphasic metabolic adaptation strategy that fully explains the paradox of why profound GP loss (87.59% knockdown) causes no mortality or developmental defects:

Phase I (emergency compensation, 48-72 h): Metabolic stress and emergency compensation during the initial response to PxGP suppression, the insect experiences genuine metabolic stress, evidenced by the glycogen levels neither accumulated nor were depleted and the transient reduction in free glucose at 72 h. The static glycogen content indicates activation of GP-independent catabolism pathways, which provide emergency glucose but are insufficient for complete homeostatic maintenance.The modest glucose dip at 72 h demonstrates that this stress is real but tolerable, likely due to several factors: (i) the magnitude is modest (only 18.15% reduction), (ii) the duration is brief (∼24 h), (iii) the timing coincides with the mid-L4 larval stage when energy demands are moderate and active feeding allows dietary carbohydrate supplementation, and (iv) trehalose stores (which remain stable during this phase) can be mobilized to buffer hemolymph glucose.

Phase II (gluconeogenic rescue, 96 h): Robust transcriptional upregulation of PEPCK and G-6-Pase (3-4-fold increases) drives massive de novo glucose synthesis from non-carbohydrate precursors such as amino acids, lactate, and glycerol. The surge of gluconeogenesis-derived glucose is rapidly converted into trehalose and G6P, creating a metabolic overshoot that restores basal glucose levels. This overcompensation ensures adequate carbohydrate supply for the energetically demanding process of pupal cuticle synthesis, which begins shortly after 96 h.

This substrate-to-product chain for gluconeogenesis-mediated metabolic rescue—protein degradation → amino acid release → gluconeogenic conversion → glucose synthesis → trehalose storage—constitutes a complete compensatory mechanism that sustains development despite profound GP ablation.

RNAi-mediated suppression of PxGP reduces enzyme activity in vivo

To confirm that RNAi-mediated transcript suppression of PxGP translates to reduced enzyme function in vivo, we measured glycogen phosphorylase a (GP-a) activity in crude extracts from dsGP-treated and dsGFP-treated larvae at 24, 48, 72, and 96 h post-injection using a coupled-enzyme spectrophotometric assay (Figure 10A, B). At 48 h post-injection, PxGP mRNA levels were reduced by 54.66% (Figure S3).

Effects of PxGP knockdown on GP enzyme activity and the expression of glycogen metabolism-related genes.

(A) GP enzyme activity normalized to sample protein concentration. A significant decrease in activity was detected only at 48 h (P < 0.05). (B) GP enzyme activity per insect. Significant decreases were detected at 24 and 48 h (P < 0.05, **P < 0.01). (C) Expression levels of the GBE gene at different time points after RNAi treatment. (D) Expression levels of the α-Amylase gene at different time points after RNAi treatment. All data are presented as mean ± SEM (n = 3); *P < 0.05, **P < 0.01, ***P < 0.001, dsGFP vs. dsGP treatment groups. An independent samples t-test was used in (A) and (B), and two-way ANOVA with Sidak’s test was used in (C) and (D).

When GP activity was normalized to total protein content, a significant reduction was detected only at 48 h (−10.35%, P < 0.05; Figure 10A). However, total protein concentration itself declined substantially — by 30.78% at 24 h — before gradually recovering over subsequent time points (Figure S4). This concurrent protein decline explains the modest per-protein activity reduction: as the protein pool shrinks, the specific activity calculation is inflated even if total enzyme function is diminished.

To address this, we also calculated GP activity on a per-larva basis. Per-larva GP-a activity was significantly reduced at both 24 h (−27.57%, P < 0.05) and 48 h (−29.28%, P < 0.01; Figure 10B), confirming functional suppression of GP at the protein level. Activity partially recovered at 72 and 96 h, consistent with the transcript recovery pattern observed in our previous RNAi experiments.

The 30.78% reduction in total protein at 24 h post-injection provides independent biochemical evidence supporting protein catabolism — consistent with the mobilization of amino acids to fuel gluconeogenesis as a compensatory response to GP suppression.

PxGP knockdown selectively activates glycogen branching enzyme expression

To investigate whether alternative glycogen metabolic pathways are activated following GP suppression, we measured the expression of glycogen branching enzyme (GBE) and α-amylase by RT-qPCR in dsGP-treated versus dsGFP-treated larvae at 24, 48, 72, and 96 h post-injection (Figure 10C, D).

GBE, which catalyzes the formation of branch points in glycogen to increase its solubility and accessibility for degradation [37], was significantly upregulated at 24 h (+29.24%, P < 0.05), 48 h (+16.78%, P < 0.05), and 96 h (+44.46%, P < 0.001; Figure 10C). In contrast, α-amylase, a glycoside hydrolase that catalyzes hydrolysis of α-(1,4)-glycosidic bonds [38], showed no significant change at any time point (Figure 10D).

This differential response — selective GBE upregulation with unchanged α-amylase expression — indicates that the compensatory response to GP suppression involves targeted remodeling of glycogen structure rather than a generalized activation of all glycogen-degrading enzymes. The enhanced glycogen branching may facilitate alternative routes of glycogen mobilization even when GP-mediated phosphorolysis is impaired.

PxGP knockdown causes transient fitness costs that are fully compensated before pupation

To determine whether the proposed metabolic compensation carries physiological consequences, we assessed multiple fitness parameters following PxGP knockdown (Figure 11A–F).

Assessment of fitness cost of PxGP gene silencing by RNAi in Plutella xylostella.

(A) Schematic diagram of fitness cost evaluation at different developmental stages of P. xylostella after dsRNA injection. (B) Food consumption per 10 larvae at different time points. (C) Average body weight per larva at different time points. (D) Pupal weight on day 3 of pupation (n = 21). (E) Adult wing area (n = 20–21). (F) Number of eggs laid by females within 3 days (n = 10–11). All data are presented as mean ± SEM (n = 3); *P < 0.05, dsGP vs. dsGFP treatment group (independent samples t-test).

Feeding rate was not significantly different between dsGP and dsGFP groups at any time point from 24 to 120 h post-injection (Figure 11B), confirming that the metabolic changes observed are independent of food intake. Larval body weight was significantly reduced in the dsGP group at 24 h (−29.10%, P < 0.05) and 48 h (−25.38%, P < 0.05; Figure 11C), indicating a transient but measurable physiological cost associated with metabolic compensation. However, pupal weight measured on day 3 after pupation showed no significant difference between groups (Figure 11D), demonstrating that larvae fully recover from the transient weight deficit before the pupal transition. Adult wing area (Figure 11E, Figures S5 and S6) and female 3-day fecundity (Figure 11F) were also not significantly affected.

This pattern — transient larval weight loss followed by complete recovery of pupal weight, adult morphology, and reproductive performance — is consistent with our metabolic compensation model: GP suppression triggers protein catabolism to supply amino acids for gluconeogenesis, resulting in short-term weight loss, but the compensatory mechanism is sufficiently effective to restore metabolic homeostasis before developmentally critical transitions.

Molecular docking reveals the structural basis for selective inhibition of PxGP by GPI

To provide structural insight into the observed selectivity, GPI potently inhibits PxGP (IC50 = 2.96 nM) while DFB does not, we performed molecular docking and binding free energy analysis using a PxGP homology model (Figure 12, Table 1).

Binding mode analysis of GPI and DFB with PxGP.

(A) 3D binding model of PxGP with GPI. (B, C) The key intermolecular interactions for GPI (B) and DFB (C). 2D interaction profile analysis of GPI (D) and DFB (E) with PxGP.

The PxGP homodimer was modeled using SWISS-MODEL with the human liver GP–acyl urea co-crystal structure (PDB: 2ATI; [15]) as the template. Molecular docking and MM/GBSA calculations were performed using Cresset Flare V11.

GPI was predicted to bind at the allosteric site located at the dimer interface, engaging seven residues across both subunits: Asn44 and Val45 from subunit A, and Trp67, Gln71, Tyr75, Arg193, and Asp227 from subunit B (Figure 12B, D). Three strong hydrogen bonds ( ≤ 2.8 Å) were formed with Asp227. This cross-subunit binding mode is consistent with the experimentally determined binding site of acyl urea inhibitors in the mammalian GP crystal structure [15].

DFB contacted six residues (Val45 from chain A; Ile68, Gln72, Tyr155, Arg193, and Arg242 from chain B) and formed two hydrogen bonds with Arg193 (2.1 and 2.7 Å; Figure 12C, E). Critically, DFB’s difluorobenzoyl moiety remained entirely solvent-exposed without productive protein contacts, explaining its inability to achieve effective target engagement.

MM/GBSA analysis confirmed GPI’s substantially more favorable binding (ΔG = −34.63 vs. −29.29 kcal/mol; ΔΔG = −5.34 kcal/mol; Table 1). Energy decomposition revealed van der Waals interactions as the primary driver of selectivity (ΔVDW = −11.49 kcal/mol), consistent with the superior shape complementarity of GPI within the dimer interface pocket.

These structural data confirm that, consistent with established structure-activity relationships of acyl urea compounds [15], target selectivity is governed by side-chain substitution pattern rather than the shared acyl urea core.

Discussion

Biochemical evidence excluding GP as a DFB target

A central challenge in mechanism-based insecticide development is distinguishing potent in vitro biochemical inhibitors from physiologically effective in vivo toxicants. This study addresses this challenge by investigating the debated molecular target of BPU insecticides. For over 40 years, the field has contented with conflicting evidence: while genetic studies implicated chitin synthase [1], direct biochemical assays consistently failed to validate it as a direct BPU target [39–41]. It has been hypothesized that BPUs might inhibit insect GP, thereby limiting glucose-derived precursors for chitin synthesis. Our systematic investigation demonstrates that while a human GPI potently inhibits PxGP, the BPU insecticide diflubenzuron does not. This discrepancy is attributable to key substitutions in its aromatic halogen substituents, which preclude effective binding to GP’s active or allosteric sites. This provides the first direct biochemical evidence excluding GP as a candidate molecular target for BPUs. Consequently, BPU lethality must arise from a different mechanism, likely involving interference with chitin assembly, cuticle organization, or CHS trafficking rather than direct enzymatic inhibition, a mechanism that remains to be conclusively identified [19, 40, 41].

It is important to delimit the scope of this conclusion. Our enzymatic and structural analyses establish that diflubenzuron, specifically, does not productively engage PxGP; they do not, by themselves, exclude every member of the structurally diverse BPU class as a potential GP ligand. BPU insecticides differ substantially in their aromatic halogen substituents and, in several cases, in additional ether or alkyl side chains (Figure 1), and our MM/GBSA analysis indicates that it is precisely these side-chain architectures — rather than the shared acylurea core — that dictate whether a given acyl urea can occupy the GP allosteric pocket (Table 1). Extending the present conclusion from diflubenzuron to the BPU class as a whole would therefore require systematic testing of additional compounds. The structural framework established here (Table 1) provides a rational basis for prioritizing such an effort, by predicting which acyl urea side-chain structures are sterically and energetically compatible with productive GP binding.

Metabolic plasticity and gluconeogenic compensation underlie GP non-essentiality

The exclusion of GP as a DFB target raised a more fundamental question: despite being a potent GP inhibitor in vitro, the GPI exhibited no insecticidal activity in vivo. This paradox—wherein a nanomolar-potency inhibitor of a putatively essential metabolic enzyme causes no phenotype—demanded mechanistic investigation.

Our investigation shows that this tolerance stems from a robust, developmentally-coordinated metabolic compensation mechanism (Figure 13). Even acute RNAi-mediated suppression of PxGP expression (87.59% knockdown) induced no discernible phenotype, as insects activated a compensatory gluconeogenic pathway. This was evidenced by the concerted suppression of downstream glycogenolytic enzymes and the significant upregulation of the key gluconeogenic enzymes PEPCK (3.95-fold) and G-6-Pase (3.34-fold) at 96 h post-knockdown. Our new experimental data substantially strengthen this evidence: GP enzyme activity measurements confirm that RNAi-mediated transcript suppression translates to functional enzyme suppression in vivo (per-larva GP-a activity reduced by ∼27–30% at 24–48 h). The concurrent 30.78% decline in total protein provides independent biochemical confirmation of protein catabolism. Furthermore, the selective upregulation of glycogen branching enzyme (GBE) — but not α-amylase — reveals a targeted compensatory response within the glycogen remodeling pathway. Fitness assessment demonstrates that this compensation, while carrying a transient larval weight cost (24–48 h), is ultimately effective: pupal weight, adult wing morphology, and female fecundity are fully maintained. This response provides a metabolic “escape route,” synthesizing glucose de novo from non-carbohydrate precursors to bypass the blocked glycogenolytic pathway and maintain homeostasis, particularly during the high carbohydrate demand preceding metamorphosis. Future work utilizing transcriptomic and proteomic approaches will be essential to delineate the signaling pathways (potentially involving nutrient sensors like AMPK and transcription factors such as FOXO) that link GP suppression to this precise reprogramming of metabolic gene expression [42, 43].

Schematic of glycogen metabolism and compensatory glucose homeostasis.

Red arrows: blocked flux from glycogen to G6P; Green arrows: activated gluconeogenesis. Solid arrows represent single enzymatic steps, while dashed arrows indicate multi-step processes.

Direct metabolite quantification validated this compensation and revealed its sophisticated temporal dynamics. The most striking finding was the lack of glycogen accumulation upon GP knockdown, a paradox explained by the activation of GP-independent glycogen catabolism via alternative enzymes like α-amylase (which hydrolyzes internal α-1,4-bonds) and glycogen debranching enzymes (which hydrolyze α-1,6-branch points) [32–34]. However, the selective upregulation of GBE (rather than α-amylase) argues against the activation of alternative catabolic routes. Instead, reduced flux into glycogen synthesis—possibly through feedback inhibition of glycogen synthase—represents a more plausible explanation, and this will be a key focus of future investigations. Definitive metabolic rescue emerged at 96 h, characterized by a dramatic surge in trehalose and G6P, directly driven by the 3-4-fold upregulation of PEPCK and G-6-Pase, providing definitive biochemical proof that de novo glucose synthesis is the primary rescue mechanism. This biphasic response—initial metabolic stress followed by delayed (48-72 h), robust compensation—creates a temporal buffer (96 h), explaining the absence of phenotypes despite severe GP knockdown. This temporal coordination suggests a programmatically timed, possibly hormonally regulated response that anticipates developmental energy needs. Critically, it reveals a metabolic vulnerability window (∼72 h); strategies that disrupt both glycogenolysis and gluconeogenesis simultaneously could overcome this compensation.

Implications for insecticide target selection: metabolic redundancy and the dual-target strategy

Our findings offer direct and actionable insights for the rational design of metabolism-targeting insecticides. A key lesson is that the viability of a metabolic enzyme as an insecticidal target is determined not solely by its biochemical function or pathway position but by the presence of compensatory metabolic routes. GP fails as a standalone target due to compensation via gluconeogenesis. This contrasts sharply with other carbohydrate-metabolizing enzymes that have proven to be highly effective insecticidal targets. For example, trehalase inhibitors such as Validamycin A are potent insecticides [44, 45], suggests that inhibition of trehalase does not cause other pathways to initiate glucose compensation. This distinction informs a refined strategy: dual-target inhibition. Simultaneously blocking both glycogenolysis (via GP) and gluconeogenesis (e.g., via PEPCK or G-6-Pase) would create irreversible glucose starvation, a “metabolic pincer” attack that could overcome adaptive compensation—a strategy with precedent in cancer therapy targeting both glycolysis and oxidative phosphorylation [46, 47]. Ultimately, our work highlights the value of systems-level metabolic modeling for target prioritization. Enzymes at metabolic branch points with redundant outputs are less vulnerable than those in non-bypassable pathways (e.g., chitin synthase, trehalase). Prospective application of computational flux balance analysis and constraint-based modeling, as supported by existing literature, can systematically identify such non-compensatable nodes for effective target selection [48].

GPI concentrations and pharmacokinetic considerations

The GPI concentrations used in our larval exposure experiments (250–500 mg/L) encompass a wide dose range. Even at the highest concentration (500 mg/L), which is approximately 409,000-fold above the in vitro IC50 of 2.96 nM, no toxicity was observed. This apparent discrepancy between potent in vitro inhibition and absent in vivo toxicity could reflect several pharmacokinetic factors: limited oral bioavailability due to incomplete absorption across the midgut epithelium; metabolic inactivation by larval detoxification enzymes (cytochrome P450s, esterases, glutathione S-transferases); or sequestration by hemolymph binding proteins [49–51]. However, the metabolic phenotype we observe — elevated trehalose, reduced protein, and upregulation of gluconeogenic enzymes — provides indirect evidence that GPI does reach its target, as these changes are consistent with GP inhibition. The most parsimonious interpretation is that GPI achieves sufficient target engagement to partially suppress GP activity in vivo, but that the resulting metabolic compensation renders this suppression non-lethal. This interpretation is further supported by our RNAi data, in which direct genetic suppression of GP (bypassing all pharmacokinetic barriers) similarly fails to cause mortality.

Study limitations and future directions

We acknowledge several limitations that point to productive future investigations. First, the transient nature of RNAi-mediated gene silencing in Lepidoptera precluded assessment of long-term or stage-specific consequences of GP loss; its role in adult flight or reproduction warrants investigation via CRISPR/Cas9 knockout or advanced dsRNA delivery systems [52–54]. Second, the generalizability of the this gluconeogenic compensation mechanism across insect orders (e.g., Coleoptera, Diptera) remains to be tested, as metabolic regulation can vary with ecology and life history. Third, while our multi-layered evidence—from transcriptional upregulation to phenotypic rescue—provides robust qualitative support for the activation of gluconeogenesis, future studies employing 13C-isotope tracer analysis would be valuable to quantify the precise metabolic rates, identify primary carbon precursors, and further refine potential metabolic targets for insecticide development [55]. Extending this metabolic profiling to include key enzymes in amino acid metabolism (e.g., aspartate aminotransferase, glutamate aminotransferase) and the dynamics of lipid content would offer a more comprehensive understanding of the systemic metabolic reprogramming that accompanies GP loss.

A further consideration concerns the contribution of dietary carbohydrates to the glucose homeostasis we observed. Because our assays were performed on actively feeding larvae — and food consumption did not differ between dsGP and dsGFP groups (Figure 11B) — ingested dietary carbohydrate was continuously available and may have helped sustain hemolymph glucose alongside endogenous gluconeogenesis. Although the equivalent feeding rates argue that the metabolic phenotype is not a secondary consequence of altered food intake, our data cannot fully partition the relative contributions of de novo glucose synthesis and dietary carbohydrate absorption to the homeostatic buffering we documented. Starvation-challenge experiments, in which GP-suppressed larvae are deprived of dietary carbohydrate, would help isolate the endogenous gluconeogenic contribution and provide additional insight into the sufficiency of metabolic compensation under nutrient-limited conditions.

Concluding remarks

In conclusion, our biochemical data demonstrate that DFB does not inhibit PxGP, adding enzyme-level evidence to the existing genetic framework establishing CHS involvement in BPU activity. More importantly, it uncovers a fundamental principle of insect metabolic biology: robust compensatory pathways, specifically gluconeogenesis, can buffer against the loss of key catabolic enzymes. This metabolic plasticity renders GP non-viable as a standalone insecticidal target but illuminates a strategic path forward: the development of dual-target inhibitors that simultaneously block glycogenolysis and gluconeogenesis. Our structural analysis provides a molecular framework for understanding why specific acyl urea side-chain structures succeed or fail as GP inhibitors, which may inform future inhibitor design. Our findings underscore the necessity of adopting a systems-level perspective in target-based insecticide discovery. Biochemical potency in vitro is necessary but not sufficient for in vivo efficacy; the latter requires consideration of metabolic network architecture, pathway redundancy, and compensatory mechanisms. As the pest management community confronts the twin challenges of insecticide resistance and environmental sustainability, this mechanistic understanding of metabolic robustness will be essential for designing the next generation of selective, effective and durable insect control agents.

Materials and Methods

Insects and rearing conditions

A susceptible laboratory strain of P. xylostella (L.), maintained for multiple generations at the College of Plant Protection, Northwest A&F University, without prior insecticide exposure, was used for all experiments. Larvae were reared on radish (Raphanus sativus) seedlings grown in vermiculite. Adults were provisioned with a 10% (w/v) honey solution. All insect stages were maintained in a controlled environment at 25 ± 1°C, 60 ± 5% relative humidity (RH), and under a 16:8 h (light:dark) photoperiod.

Cell culture and maintenance

Spodoptera frugiperda (Sf9) cells (Gibco, CA, USA) were cultured in Sf-900™ III SFM medium (Gibco) supplemented with 10% fetal bovine serum (FBS; Gibco) and 1% (v/v) Antibiotic-Antimycotic solution (Gibco, Cat# 15240062). Adherent monolayer cultures were maintained in T-flasks at 27°C. For large-scale protein expression, cells were adapted to suspension culture in baffled flasks, agitated at 120 rpm at 27°C. Cells were harvested by centrifugation (4,000 rpm, 10 min, 4°C) when the density reached 5.0 × 106 cells/mL. Cell pellets were flash-frozen in liquid nitrogen and stored at -80°C until use.

Chemicals and reagents

Diflubenzuron (DFB, 1-(4-chlorophenyl)-3-(2,6-difluorobenzoyl) urea, ≥98% purity) was purchased from Shanghai Topscience Co., Ltd. (Shanghai, China, Cat# T0109).

The Glycogen phosphorylase-IN-1 (GPI), 1-(3-(3-(2-Chloro-4,5-difluorobenzoyl)ureido)-4-methoxyphenyl)-3-methylurea (also known as compound 42 in ref [15]), was purchased from TargetMol (Shanghai, China, Cat# T37577, ≥98% purity).

Glycogen (from bovine liver, Type III), β-glycerophosphate, HEPES, imidazole, EDTA, DTT, ATP, MgCl2, protease inhibitor cocktail components (AEBSF, Pepstatin A, Leupeptin), and all other biochemical reagents were purchased from Sigma-Aldrich (Shanghai, China) or Sangon Biotech (Shanghai, China) and were of molecular biology grade (≥95% purity) unless otherwise specified.

Recombinant PxGP expression and purification

The full-length coding sequence of PxGP (2,511 bp, encoding 837 amino acids; GenBank Accession No. XM_038119125.2 was amplified from P. xylostella third-instar larval cDNA by PCR using gene-specific primers (Table S1), and cloned into the pFastBac1 vector (Invitrogen, Carlsbad, CA, USA) using homology-based cloning (ClonExpress II One Step Cloning Kit, Vazyme, Nanjing, China). The construct was engineered to express PxGP with an N-terminal 6×His tag (HHHHHH-) immediately following the initiating methionine, under the control of the polyhedrin (PH) promoter. The sequence fidelity of the construct was verified by Sanger sequencing (Tsingke Biotechnology, Beijing, China).

The sequence-verified pFastBac1-PxGP construct was transformed into DH10Bac™ competent cells (Invitrogen) to generate recombinant bacmid DNA via Tn7-mediated transposition, following the manufacturer’s Bac-to-Bac® protocol. Recombinant bacmid DNA was isolated, verified by PCR using M13 forward and reverse primers (Table S1), and transfected into Sf9 cells using Cellfectin® II Reagent (Invitrogen) to generate P0 viral stock. The viral titer was amplified through two additional passages (P1, P2) to obtain high-titer P3 viral stock.

For large-scale protein production, Sf9 cells in suspension culture (2.0 × 106 cells/mL, 500 mL culture volume in a 2-L baffled Erlenmeyer flask) were infected with P3 viral stock at a Multiplicity of Infection (MOI) of 5. Infected cultures were maintained at 27°C with agitation at 120 rpm. Cells were harvested 72 h post-infection by centrifugation (4,000 rpm, 10 min, 4°C), yielding approximately 8.25 g of cell pellet. Cell pellets were flash-frozen in liquid nitrogen and stored at -80°C until use.

For protein purification, frozen cell pellets were thawed on ice and resuspended in ice-cold Lysis Buffer (50 mM β-glycerophosphate [pH 7.5], 150 mM NaCl, 0.2 mM DTT, 0.5 mM EDTA, and 1× protease inhibitor cocktail [1 mM AEBSF, 15 µM Pepstatin A, 20 µM Leupeptin]) at a ratio of 4 mL buffer per gram of cell pellet. Cells were lysed by sonication on ice using a Scientz-IID sonicator (Ningbo Scientz Biotechnology, China) with the following parameters: 6 cycles of 30 s ON / 60 s OFF at 40% amplitude. The lysate was clarified by centrifugation (16,000 × g, 30 min, 4°C).

The clarified supernatant was loaded onto a 5-mL HisTrap™ HP column (GE Healthcare, Uppsala, Sweden) pre-equilibrated with Lysis Buffer containing 10 mM imidazole. The column was first washed with 10 column volumes (CV) of Equilibration Buffer (Lysis Buffer + 10 mM imidazole), followed by 5 CV of Wash Buffer (Lysis Buffer + 30 mM imidazole). Bound 6×His-tagged PxGP protein was eluted with a 15 CV linear gradient from 30 to 200 mM imidazole in Lysis Buffer. Fractions (2 mL each) were collected and analyzed by 13% SDS-PAGE with Coomassie Brilliant Blue R-250 staining. Fractions containing PxGP, identified as a prominent band at ∼100 kDa, were pooled and dialyzed overnight at 4°C against Storage Buffer (50 mM HEPES [pH 7.4], 150 mM NaCl, 1 mM DTT, 10% [v/v] glycerol) using a 10 kDa molecular weight cut-off dialysis membrane (Spectrum Labs, Rancho Dominguez, CA, USA), with three buffer changes. Protein concentration was determined using the Pierce™ BCA Protein Assay Kit (Thermo Fisher Scientific, Waltham, MA, USA) with bovine serum albumin (BSA) as the standard. Aliquots of purified PxGP were flash-frozen in liquid nitrogen and stored at -80°C. The typical yield was 7.5 mg of purified protein from 8.25 g of cell pellet.

In vitro activation of recombinant PxGP

Purified recombinant PxGP exists predominantly in the inactive dephosphorylated b form (PxGP-b) and requires phosphorylation at a conserved serine residue (corresponding to Ser14 in mammalian GP) for conversion to the active a form (PxGP-a) [17]. Activation was performed in vitro using immobilized Phosphorylase Kinase (PhK) (Cat# HY-P2757, MedChemExpress, Monmouth Junction, NJ, USA).

Immobilized PhK beads (100 µL slurry) were washed three times with 1 mL of Kinase Buffer (25 mM β-glycerophosphate, pH 7.5, 0.5 mM EDTA, 0.25 mM DTT, 1× protease inhibitor cocktail) and resuspended in 500 µL of dialyzed PxGP-b solution (final protein concentration: 1 mg/mL). The activation reaction was initiated by adding ATP (sodium salt) to a final concentration of 7 mM and MgCl2 to 22 mM. The suspension was incubated for 2 h at room temperature (22-25°C) with gentle end-over-end rotation.

Following incubation, the activated PxGP-a (present in the supernatant) was separated from the immobilized PhK beads using a magnetic separation stand. The supernatant was passed through a 0.22 µm syringe filter (Millipore) to remove any residual beads and was used immediately for enzymatic assays. The extent of activation was assessed by measuring specific activity (see Section 2.7). Activated PxGP-a was used fresh and not stored, as phosphorylated GP is subject to dephosphorylation by endogenous phosphatases over time.

Western blot analysis

Protein samples were separated by 13% SDS-PAGE (120 V, 90 min) and electrophoretically transferred to a nitrocellulose (NC) membrane (0.45 µm pore size, Millipore, Burlington, MA, USA) at 300 mA for 90 min at 4°C using a Bio-Rad Mini Trans-Blot® system. The membrane was blocked for 1 h at room temperature in 5% (w/v) non-fat dry milk in TBST (10 mM Tris-HCl, pH 7.4, 150 mM NaCl, 0.1% [v/v] Tween-20).

The membrane was incubated overnight at 4°C with an anti-His mouse monoclonal antibody (1:1500 dilution in blocking buffer; Cat# ABT2050, Abbkine, Wuhan, China). After three washes with TBST (10 min each), the membrane was incubated for 1 h at room temperature with an alkaline phosphatase (AP)-conjugated goat anti-mouse IgG secondary antibody (1:4000 dilution in blocking buffer; Cat# A0208, Beyotime Biotechnology, Shanghai, China). Following three additional TBST washes, immunoreactive bands were visualized using BM Purple AP Substrate (Roche, Basel, Switzerland) according to the manufacturer’s instructions. Images were captured using a ChemiDoc™ XRS+ imaging system (Bio-Rad, Hercules, CA, USA).

Glycogen phosphorylase activity assay

GP enzymatic activity was measured using a coupled-enzyme spectrophotometric assay kit (Cat# BC3345, Solarbio Life Science, Beijing, China). This assay quantifies GP activity by measuring the GP-catalyzed phosphorolysis of glycogen to glucose-1-phosphate (G-1-P). The product G-1-P is enzymatically converted to glucose-6-phosphate (G-6-P) by phosphoglucomutase (PGM), and G-6-P is subsequently oxidized by glucose-6-phosphate dehydrogenase (G6PDH), which stoichiometrically reduces NADP+ to NADPH. The rate of NADPH formation, which is directly proportional to GP activity, was monitored by measuring the increase in absorbance at 340 nm (A340) using a Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland).

The reaction mixture (200 µL total volume per well in a 96-well UV-transparent microplate) contained the following components (as supplied in the kit): 160 µL of Reaction Buffer (containing glycogen [2 mg/mL final], inorganic phosphate [10 mM], AMP [1 mM], NADP+ [0.4 mM], PGM, G6PDH, MgCl2, and appropriate buffer), 30 µL of deionized water, and 10 µL of enzyme sample (purified PxGP-a or crude larval lysate). The reaction was initiated by adding the enzyme sample, and A340 was recorded continuously at 15 s intervals for 30 min at 25°C.

Specific activity was calculated using the following formula: Specific Activity(U/mg)=(ΔA340/min)/(ε×l×[Protein])×Vt/Vs×1000.

where:

ΔA340 represents the change in absorbance at 340 nm,

ε = molar extinction coefficient of NADPH at 340 nm (6.22 mM-1 · cm-1),

l = light path length (0.6 cm for 200 µL in a 96-well microplate),

[Protein] = protein concentration in the assay (mg/mL),

Vt = total reaction volume,

Vs = sample volume added.

One Unit (U) of GP activity is defined as the amount of enzyme that catalyzes the formation of 1 µmol of NADPH per minute under the assay conditions.

Inhibitor screening and IC50 determination

For in vitro enzyme inhibition assays, purified activated PxGP-a was diluted to 10 µg/mL in Assay Dilution Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl2. Test compounds (GPI, DFB) were prepared as 100× stock solutions in DMSO and serially diluted in Assay Dilution Buffer to achieve final assay concentrations ranging from 0.16 nM to 500 nM (for GPI) and 0.54 nM to 1700 nM (for DFB), respectively (final DMSO concentration ≤ 1% [v/v] in all wells, including vehicle controls).

The enzyme (10 µL of 10 µg/mL PxGP-a, final concentration 10 µg/mL in the assay) was pre-incubated with 10 µL of inhibitor solution (or vehicle control) for 15 min at room temperature in a 96-well microplate. The reaction was then initiated by adding 180 µL of the Reaction Buffer (as described in Section 2.7), and A340 was monitored continuously for 30 min at 25°C.

Residual enzyme activity at each inhibitor concentration was calculated as a percentage of the vehicle control (1% DMSO alone). Dose-response curves were generated by plotting percent residual activity versus log[inhibitor]. IC50 values (the concentration of inhibitor required to reduce enzyme activity by 50%), Hill slopes, and 95% confidence intervals (CI) were calculated by non-linear regression analysis using a four-parameter logistic equation in GraphPad Prism 9.0 (GraphPad Software Inc., San Diego, CA, USA): Y=Bottom+(Top−Bottom)/(1+10(LogIC50−X)×Hill slope). where Y is percent residual activity, X is log[inhibitor], Top and Bottom are the upper and lower plateaus of the curve (constrained to 100% and 0%, respectively), and Hill slope describes the steepness of the curve.

For each compound, at least three independent experiments were performed, each with triplicate technical replicates.

Preparation of larval crude lysate for in vivo activity

To assess the inhibitory effects of GPI and DFB on native GP activity in vivo, crude enzyme extracts were prepared from P. xylostella larvae. Healthy third-instar (L3) larvae (n = 30 per biological replicate, three replicates total) were collected, briefly washed in ice-cold phosphate-buffered saline (PBS, pH 7.4) to remove surface contaminants, and gently blotted dry on filter paper. Larvae were then homogenized on ice in 3 volumes (w/v) of ice-cold Homogenization Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl2, 1× protease inhibitor cocktail).

The homogenate was centrifuged at 15,000 × g for 20 min at 4°C, and the supernatant (crude enzyme extract) was carefully collected, avoiding the lipid layer and pellet. Protein concentration was determined using the BCA assay, and extracts were either used immediately for activity assays or aliquoted and stored at -80°C (activity remained >90% for up to 2 weeks).

GP activity in crude lysates was measured using the same coupled-enzyme assay described in Section 2.7, with the following modifications: crude lysate (diluted to 2 mg/mL total protein) was used in place of purified enzyme, and pre-incubation with inhibitors was performed for 5 to 45 min to assess time-dependent inhibition. Final inhibitor concentrations tested were 10-5, 10-6 or 10-7 M.

Sequence alignment and bioinformatic analysis

The amino acid sequence of PxGP (XP_037975053.2) was obtained from the NCBI database. Multiple sequence alignment with orthologs from Homo sapiens (NP_002854.3), Oryctolagus cuniculus (NP_001075653), Helicoverpa armigera (XP_021190210.1), Spodoptera exigua (ACN78408.1), and Ostrinia furnacalis (AFO54708.2) was performed using ClustalX. Conserved domains and motifs were analyzed using the ScanProsite and NCBI CD-Search web servers. Protein pairwise similarity matrix was visualized as a heatmap using the “HeatMap” module in TBtools-II (v2.390) [56].

Larval bioassays

Larval bioassays were conducted using a standard leaf-dip method adapted from established protocols [57]. Stock solutions of GPI and DFB were prepared in HPLC-grade acetone at 10,000 mg/L. Serial dilutions were prepared in deionized water containing 0.1% (v/v) Triton X-100 as a wetting agent to achieve final test concentrations of 250 and 500 mg/L for GPI, and 125 and 250 mg/L for DFB. The final acetone concentration in all treatment solutions was 1% (v/v). A solvent control group (CK) consisted of 1% acetone and 0.1% Triton X-100 in water without any test compound.

Fresh cabbage (Brassica oleracea var. capitata) leaf discs (5 cm diameter) were cut using a cork borer, dipped in the respective treatment solutions for 6-8 s with gentle agitation, and then air-dried for 30 min at room temperature in a fume hood. Once dry, each leaf disc was placed in a sterile 90 mm petri dish lined with moistened filter paper.

Ten synchronized third-instar larvae of P. xylostella (24-36 h post-molt to L3) were transferred to each treated leaf disc using a soft brush. Each treatment consisted of three biological replicates (i.e., three Petri dishes, 30 larvae total per treatment). Petri dishes were sealed with Parafilm® and maintained under standard rearing conditions (25 ± 1°C, 60 ± 5% RH, 16:8 h L:D photoperiod). Leaf discs were replaced with freshly treated discs every 12 h to ensure continuous exposure.

Mortality was recorded every 24 h for 120 h (5 days). Larvae that failed to respond to gentle prodding or showed moribund behavior (inability to right themselves or move coordinately) were considered dead. Corrected mortality was calculated using Abbott’s formula [58]: Pupation rates (number of pupae formed / initial number of larvae × 100) and adult emergence rates (number of emerged adults / number of pupae formed × 100) were recorded at 8 and 10 days post-treatment, respectively.

Lethal concentration values (LC30 and LC50) and 95% confidence intervals for DFB were calculated using probit analysis in POLO-Plus software (LeOra Software, Berkeley, CA, USA).

GPI exposure for gene expression analysis

For time-course gene expression analysis, synchronized third-instar larvae were exposed to 500 mg/L GPI (the highest concentration tested in bioassays) or solvent control (1% acetone + 0.1% Triton X-100) using the leaf-dip method described in Section 2.11.

Surviving larvae were collected at 24, 48, 72, and 96 h post-treatment. At each time point, three biological replicates (10 pooled larvae each) were flash-frozen in liquid nitrogen and stored at -80°C until RNA extraction. This sampling scheme ensured sufficient RNA yield while minimizing inter-individual variability.

Quantitative real-time PCR (RT-qPCR) analysis

Total RNA was extracted from frozen samples using RNAiso Plus reagent (TaKaRa, Kusatsu, Japan). RNA concentration and purity were assessed by NanoDrop 2000c spectrophotometer (Thermo Fisher Scientific). Only samples with A260/A280 ratios between 1.9 and 2.1 and A260/A230 ratios >1.8 were used for downstream analysis. RNA integrity was confirmed by agarose gel electrophoresis (sharp 28S/18S rRNA bands, ratio ≈c2:1). cDNA was synthesized from 1cµg RNA using PrimeScript™ RT Reagent Kit with gDNA Eraser (TaKaRa).

Gene-specific primers for PxGP, trehalase (PxTre), hexokinase (PxHex), phosphoenolpyruvate carboxykinase (PxPEPCK), PxG-6-Pase, and the reference gene PxRPS13 (ribosomal protein S13) were designed using Primer-BLAST (NCBI) and are listed in Table S1. PxRPS13 was selected based on its stable expression across developmental stages and treatment conditions (geNorm M value < 0.5) [59]. All primers were validated for specificity (single peak in melt-curve), efficiency (90-110%, R² > 0.98), and absence of primer-dimers (no-template controls).

RT-qPCR reactions were performed on a QuantStudio™ 5 Real-Time PCR System (Thermo Fisher Scientific) using TB Green® Premix Ex Taq™ II (Tli RNaseH Plus) (2×) (TaKaRa) in 20 µL reactions. The thermal cycling conditions were as follows: initial denaturation at 95°C for 30 s; 40 cycles of 95°C for 15 s (denaturation) and 60°C for 30 s (annealing/extension); followed by melt-curve analysis. Each sample was analyzed in technical triplicate, and each biological replicate consisted of three independent samples. Relative gene expression was calculated using the 2-ΔΔCt method [60].

RNA interference (RNAi)

A conserved 606 bp region of PxGP (nucleotides 66-671, part of the AMP-binding domain) was selected for RNAi. The fragment was amplified from larval cDNA using primers flanked by T7 promoters (Table S1). Double-stranded RNA (dsRNA) was synthesized with the T7 RiboMax™ Express RNAi System (Promega, Madison, WI, USA). The dsRNA of green fluorescent protein (dsGFP, 541 bp) was synthesized using a pGFP plasmid template as a non-specific negative control. Following synthesis, dsRNA was treated with RNase-free DNase I (30 min at 37°C), purified by isopropanol precipitation, and integrity verified by 1.5% agarose gel electrophoresis. Concentration and purity (A260/A280=1.9-2.1) was measured using a NanoDrop spectrophotometer. Purified dsRNA was resuspended in nuclease-free water at concentrations of 6,000, 10,000, or 14,000 ng/µL and stored in single-use aliquots at -80°C.

For microinjection, synchronized early third-instar larvae (6-12 h post-molt) were briefly chilled on ice for 2-3 min and injected laterally using a DMP-300 Injector (Micrology Precision Instruments, Ltd, Wuhan, China). After recovery on fresh cabbage leaves for 30 min at room temperature, larvae were maintained under standard rearing conditions. Phenotypic parameters (mortality, molting success, pupation rate, adult emergence) were monitored daily for 10 days. Larvae were collected at 24, 48, 72, and 96 h post-injection for RT-qPCR analysis to confirm target gene knockdown (n = 3 biological replicates, each containing 10 pooled larvae per time point).

Metabolite quantification

To directly assess metabolic flux alterations following PxGP knockdown, we quantified four key metabolites in the glycogen-glucose metabolic axis using commercial enzymatic assay kits: glycogen and glucose (Glycogen Assay Kit, Grace Biotechnology, Suzhou, China, Cat# G0590W96), trehalose (Trehalose Assay Kit, Grace Biotechnology, Cat# G0553W96), and glucose-6-phosphate (G6P Assay Kit, Beyotime Biotechnology, Shanghai, China, Cat# S0185). All kits utilize colorimetric or fluorometric enzymatic reactions and were performed according to the manufacturers’ protocols with minor modifications as described below.

Third-instar larvae injected with dsGP (14,000 ng/µL, the highest dose achieving maximal knockdown) or dsGFP control were collected at 48, 72, and 96 h post-injection. At each time point, three biological replicates were collected, each consisting of 10 pooled larvae. Samples were immediately flash-frozen in liquid nitrogen and stored at -80°C until analysis.

For metabolite extraction, frozen larvae were homogenized on ice in ice-cold extraction buffer specific to each metabolite assay (typically 100 µL buffer per 10 mg tissue). For glycogen extraction, samples were homogenized in deionized water and immediately boiled for 3 min to inactivate endogenous glycogenolytic enzymes. For trehalose, glucose, and G6P extraction, samples were homogenized in the respective assay buffers provided in the kits. Homogenates were centrifuged at 12,000 × g for 10 min at 4°C, and supernatants were immediately used for metabolite quantification or stored at -80°C for no more than 24 h.

Metabolite quantification was performed in 96-well microplates using Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland) at the wavelengths specified in each kit’s protocol. Sample metabolite concentrations were calculated by comparison with the absorbance of the kit’s standard. Metabolite concentrations were normalized to tissue protein content to account for variation in tissue input. Protein concentration in the crude lysates was determined using the Protein Quantification Kit (BCA Assay) (Abbkine Scientific) with bovine serum albumin as the standard. All samples were analyzed in technical triplicate. Final metabolite levels were expressed as: Glycogen: mg/mg protein; Glucose: µmol/mg protein; Trehalose: µmol/mg protein; G6P: µmo/mg protein. Because metabolite levels are expressed per unit total protein, the changes in total protein content that accompany PxGP knockdown bear directly on their interpretation. Total protein declined substantially at the earlier time points (up to ∼30%; Figure 9D, Figure S4) before recovering, which means that for any metabolite whose absolute amount was unchanged, the protein-normalized value would be correspondingly inflated during the protein-depleted window. The protein-normalized metabolite data reported here should therefore be interpreted with this denominator effect in mind. We retained protein normalization because it remains the most appropriate correction for the variation in tissue input inherent to pooled-larvae sampling; we note, however, that absolute (per-larva) quantification — as we additionally applied to GP enzyme activity (Figure 10A, B) — provides a complementary, denominator-independent perspective and represents a useful refinement for future metabolite profiling.

Glycogen phosphorylase activity assay in RNAi-treated larvae

To assess GP enzyme activity following RNAi, early third-instar larvae were injected with 14,000 ng/μL dsGP or dsGFP. At 24, 48, 72, and 96 h post-injection, 10 larvae per replicate (three biological replicates per time point) were collected. Active glycogen phosphorylase a (GP-a) activity was measured using a coupled-enzyme spectrophotometric assay kit (catalog no. BC3345, Solarbio, Beijing, China) following the manufacturer’s protocol as described above. GP activity was normalized both to total protein content (U/mg protein) and to individual larva (U/larva).

RT-qPCR analysis of alternative glycogen metabolic enzymes

Homologs of human glycogen branching enzyme (GBE, CAG9111719.1), Pieris rapae α-amylase (XP_011548197.3), and glycogen debranching enzyme (GDE, XP_048484812.1) were identified in the P. xylostella genome database by sequence homology search. RT-qPCR primers were designed using Primer-BLAST (NCBI) (Table S1). Multiple primer pairs designed for GDE failed to yield amplification products; therefore, only GBE and α-amylase were analyzed. PxRPS13 (ribosomal protein S13) was used as the reference gene. Total RNA was extracted from larvae collected at 24, 48, 72, and 96 h post-injection (three biological replicates, 10 larvae per replicate). Relative expression levels were calculated using the 2−ΔΔCt method and analyzed by two-way ANOVA with Šidák’s multiple comparisons test [61].

Fitness cost assessment

To evaluate fitness consequences of PxGP knockdown, the following parameters were measured. Feeding rate: cabbage leaf discs of uniform size were provided every 12 h, and remaining leaf area was measured by ImageJ at 24 h intervals from 24 to 120 h post-injection. Larval weight: 10 larvae per replicate were weighed at 24 h intervals. Pupal weight: all pupae were weighed on day 3 after pupation. Adult wing area: forewings were photographed using a SteREO Discovery V20 stereomicroscope (Zeiss, Germany) and measured using ImageJ. Female fecundity: newly emerged females were paired with untreated males, and eggs deposited on leaf discs were counted for 3 consecutive days. Three biological replicates were performed. Data were analyzed by independent-samples t-test.

Homology modeling and molecular docking

The three-dimensional structure of PxGP was generated by homology modeling using SWISS-MODEL (https://swissmodel.expasy.org [62]) with the crystal structure of human liver glycogen phosphorylase a in complex with an acyl urea inhibitor (PDB: 2ATI [15]) as the template. Since PxGP functions as a homodimer, the dimeric biological assembly was constructed based on the template’s quaternary structure. The three-dimensional structures of GPI (PubChem CID: 10070301) and DFB (PubChem CID: 37123) were obtained from PubChem in SDF format. Molecular docking was performed using Cresset Flare V11 (Cresset, Litlington, UK). Binding free energies were estimated by MM/GBSA (Molecular Mechanics with Generalized Born and Surface Area solvation) single-point calculations. Two-dimensional protein-ligand interaction diagrams were generated within Flare V11.

Use of artificial intelligence tools

AI-based language editing tools were used to assist with English grammar and style refinement during manuscript preparation. All scientific content, including experimental design, data collection, analysis, and interpretation, was generated entirely by the authors.

Data are presented as mean ± standard error of the mean (SEM). Statistical significance was set at P < 0.05 (*), P < 0.01 (**) or P < 0.001 (***).

eLife Assessment

This study provides valuable evidence that glycogen phosphorylase is unlikely to represent an effective insecticidal target in Plutella xylostella and that diflubenzuron does not directly inhibit this enzyme. The combination of biochemical, molecular and physiological approaches provides convincing data to support these conclusions, although the proposed metabolic compensation mechanism is supported primarily by indirect evidence and would benefit from direct demonstration.

Reviewer #1 (Public review):

Summary:

In this study, the authors investigate whether glycogen phosphorylase represents a molecular target of benzoylphenylurea insecticides and evaluate the physiological consequences of suppressing glycogen phosphorylase activity in the diamondback moth Plutella xylostella. The authors combine recombinant protein biochemistry, enzyme inhibition assays, RNA interference, structural modelling, metabolite profiling, gene expression analyses, and physiological measurements to determine whether diflubenzuron directly inhibits glycogen phosphorylase and whether suppression of this enzyme is sufficient to impair insect development. Based on these experiments, the authors conclude that diflubenzuron does not directly inhibit glycogen phosphorylase and that insects tolerate substantial suppression of this enzyme through compensatory metabolic responses.

Strengths:

This study addresses an important question in insect toxicology by systematically evaluating glycogen phosphorylase as a potential insecticidal target. The authors combine complementary biochemical, molecular, physiological, and structural approaches, including recombinant enzyme characterization, inhibitor assays, RNA interference, metabolite profiling, structural modelling, and measurements of fitness-related traits. This integrative approach provides a comprehensive evaluation of the biological consequences of glycogen phosphorylase suppression. In particular, the biochemical evidence that diflubenzuron does not inhibit glycogen phosphorylase, together with the observation that strong suppression of glycogen phosphorylase produces only transient physiological effects without measurable impacts on development or reproduction, provides strong support for the conclusion that glycogen phosphorylase is unlikely to represent an effective standalone insecticidal target.

Weaknesses:

The main limitation concerns the proposed mechanism underlying metabolic compensation. The observed increases in gluconeogenic gene expression, changes in metabolite abundance, and reductions in total protein are consistent with activation of compensatory metabolism, but are insufficient to directly demonstrate increased gluconeogenic flux or establish that amino acid-derived carbon is incorporated into newly synthesized glucose. Similarly, although the analyses of glycogen-associated enzymes strengthen the discussion of alternative metabolic pathways, changes in gene expression alone do not demonstrate that these pathways contribute to glycogen utilization in vivo.

Some mechanistic interpretations therefore extend beyond the data presented. For example, decreases in total protein are interpreted as evidence of protein catabolism fuelling gluconeogenesis, yet they do not directly demonstrate amino acid mobilization or incorporation into glucose. Likewise, increased expression of gluconeogenic genes is interpreted as evidence of increased pathway activity, although transcriptional changes do not necessarily reflect metabolic flux. Finally, the absence of major developmental defects following glycogen phosphorylase suppression is attributed primarily to metabolic compensation, but an alternative explanation is not fully considered. Such explanation could be that glycogen phosphorylase is not rate-limiting for glucose homeostasis under the nutrient-rich experimental conditions, where dietary carbohydrates are continuously available. Consequently, the proposed compensatory mechanism remains plausible and well supported by indirect evidence, but several aspects would benefit from more cautious interpretation.

Overall, the authors successfully achieve their primary objective of evaluating glycogen phosphorylase as a candidate insecticidal target. The study provides useful biochemical and physiological evidence that this enzyme is unlikely to represent an effective target for insecticide development in P. xylostella, while highlighting the importance of metabolic plasticity when assessing metabolic targets. The experimental approaches and datasets presented here should be valuable to researchers studying insect metabolism, insecticide mode of action, and target validation, although the precise mechanisms underlying the proposed metabolic compensation remain an important subject for future investigation.

Author response:

The following is the authors’ response to the previous reviews.

Public Reviews:

Reviewer #1 (Public review):

Strengths:

This study addresses an important question in insect toxicology by systematically evaluating glycogen phosphorylase as a potential insecticidal target. The authors combine complementary biochemical, molecular, physiological, and structural approaches, including recombinant enzyme characterization, inhibitor assays, RNA interference, metabolite profiling, structural modelling, and measurements of fitness-related traits. This integrative approach provides a comprehensive evaluation of the biological consequences of glycogen phosphorylase suppression. In particular, the biochemical evidence that diflubenzuron does not inhibit glycogen phosphorylase, together with the observation that strong suppression of glycogen phosphorylase produces only transient physiological effects without measurable impacts on development or reproduction, provides strong support for the conclusion that glycogen phosphorylase is unlikely to represent an effective standalone insecticidal target.

We thank the reviewer for the positive assessment of our study and for recognizing the value of our integrative approach, including recombinant enzyme characterization, RNAi, metabolite profiling, structural modelling, and fitness measurements. We are also grateful for the acknowledgement that our biochemical evidence and phenotypic observations provide strong support for the conclusion that GP is unlikely to be an effective standalone insecticidal target.

Weaknesses:

(1) The proposed metabolic compensation mechanism is supported by indirect evidence.

We agree with the reviewer that our data demonstrate correlation with, rather than direct proof of, increased gluconeogenic flux. We have revised the manuscript throughout to moderate our interpretations and to clearly frame the compensation model as a plausible interpretation supported by multiple lines of indirect evidence, rather than an established mechanism. Specific revisions are detailed below.

Specific revisions:

(1) Abstract (Lines 35–39):

“We provide evidence that insects compensate through a multi-layered metabolic response: upregulation of gluconeogenic enzymes (PEPCK, G-6-Pase), selective upregulation of glycogen branching enzyme (GBE) but not α-amylase, and changes in protein content suggestive of catabolic substrate mobilization”

(2) Abstract (Lines 41–42):

“Indicating that the metabolic compensation response is ultimately effective in sustaining development”

(3) Abstract (Lines 42–44):

“These findings suggest that GP is functionally non-essential under these conditions, likely through a combination of gluconeogenic compensation and the availability of alternative carbon sources”

(4) Introduction (Lines 88–90):

“Our findings reveal that PxGP is functionally non-essential for larval development under the tested conditions, reflecting a previously uncharacterized gluconeogenic compensation mechanism”

(5) Results – Gene expression (Line 265):

“Gene expression analysis is consistent with gluconeogenic activation”

(6) Results – Protein content (Line 278):

“Changes in protein content suggestive of catabolic substrate mobilization”

(7) Results – Protein decline (Line 283):

“This protein decline is consistent with substrate mobilization”

(8) Results – GBE expression (Line 389):

“PxGP knockdown selectively upregulates glycogen branching enzyme expression”

(9) Results – GBE differential response (Lines 400–403):

“This differential expression pattern—selective GBE upregulation with unchanged α-amylase expression—indicates that the compensatory response to GP suppression involves targeted remodeling of glycogen structure rather than a generalized upregulation of all glycogen-degrading enzymes.”

(10) Results – Fitness assessment (Lines 411–412):

“To determine whether the metabolic changes observed following PxGP knockdown are associated with measurable physiological consequences”

(11) Results – Fitness interpretation (Lines 426–430):

“GP suppression likely triggers protein catabolism to supply amino acids for gluconeogenesis, contributing to transient weight loss. In the presence of continuous dietary carbohydrate supply, the compensatory response appears sufficiently effective to restore metabolic homeostasis before developmentally critical transitions”

(12) Discussion – Glycogen accumulation paradox (Line 520-521):

“a paradox that could be explained by the activation of GP-independent glycogen catabolism via alternative enzymes like...”

(13) Results – Trehalose and G6P (Lines 299–303, 315–317):

“Trehalose levels remained stable at 48-72 h but increased substantially by 96 h (7.44-fold elevation, P < 0.05) (Figure 9F). This increase coincided with the significant upregulation of gluconeogenic genes (e.g., PEPCK and G-6-Pase), consistent with the idea that gluconeogenesis contributes to trehalose synthesis and ensures adequate carbohydrate reserves for the upcoming pupation”

And: “Together, the coordinated behavior of G6P and trehalose is consistent with gluconeogenesis-derived glucose being converted into storage and transport carbohydrates”

(14) Results – Integrated interpretation (Lines 330–332):

“Collectively, these metabolite and gene expression data reveal a biphasic metabolic adaptation that provides a coherent explanation for why substantial GP suppression causes no mortality or developmental defects”

(15) Discussion – Definitive proof (Lines 527–531):

“At 96 h, trehalose and G6P levels in dsGP-treated larvae were maintained at markedly higher levels than in dsGFP controls, which had declined by this time point, coinciding with 3–4‑fold upregulation of PEPCK and G‑6‑Pase. These changes strongly support, albeit indirectly, de novo glucose synthesis as the primary rescue mechanism.”

(16) Discussion – Protein catabolism confirmation (Lines 505):

“provides independent biochemical support for protein catabolism”

(2) Some mechanistic interpretations extend beyond the data presented.

We accept this criticism and have systematically revised the manuscript to distinguish more carefully between observed transcriptional/metabolic changes and functional interpretations. We now consistently frame these as correlative evidence consistent with—but not proving—the proposed mechanisms.

Specific revisions:

The revisions listed under Weakness 1 above—particularly those modifying language around protein decline (Lines 278, 283), gene expression (Lines 265, 389), and trehalose/G6P changes (Lines 299–303, 314–317)—also directly address this weakness by removing the implication that transcriptional or protein-level changes constitute functional proof of pathway activation. Additionally, we have made the following revisions:

(17) Additional discussion of GBE/α-amylase limitation (Line 405-408)

We have added explicit acknowledgement that transcriptional changes alone do not demonstrate functional pathway activation:

“However, as these observations are limited to the transcript level, further enzymatic activity assays or glycogen structure analyses would be required to determine whether this transcriptional change translates into functional alterations in glycogen mobilization.”

(18) Fitness section – clarification (Lines 426–430):

We have revised the interpretation of transient larval weight loss as described above, introducing “likely” and clarifying that the data are consistent with the model rather than conclusively demonstrating it.

(3) An alternative explanation for the limited phenotype is not fully considered.

We thank the reviewer for this important point. We agree that the continuous availability of dietary carbohydrates under our experimental conditions represents a plausible alternative explanation for the lack of phenotype. We have added explicit discussion of this alternative interpretation in the Abstract, Introduction, and Discussion sections.

Specific revisions:

(19) Abstract (Lines 42–44):

As shown above, we have revised the abstract to include: “These findings suggest that GP is functionally non-essential under these conditions, likely through a combination of gluconeogenic compensation and the availability of alternative carbon sources.”

(20) Introduction (Lines 88–90):

As shown above, we have revised the introduction to include: “under the tested conditions” and to acknowledge that the observed phenotype reflects a combination of factors.

(21) Discussion – Biphasic response (Lines 531–539):

We have revised the discussion of the biphasic response to moderate the causal language and acknowledge alternative interpretations:

“This biphasic response—initial metabolic stress followed by delayed (48-72 h) and robust compensation—creates a temporal buffer by 96 h and is consistent with the absence of overt phenotypes despite substantial GP suppression. This temporal coordination raises the possibility of a developmentally programmed response, potentially involving hormonal regulation, that anticipates energy demands at critical transitions. Critically, this pattern suggests a potential vulnerability window (~72 h) during which combined inhibition of glycogenolysis and gluconeogenesis might overcome compensation, although this remains speculative and would require experimental validation.”

(22) Discussion – Alternative explanation paragraph (Lines 540–547):

We have added a following paragraph explicitly addressing the alternative explanation:

“We also acknowledge that the absence of a severe phenotype may reflect an additional, non‑mutually exclusive explanation: under our experimental conditions, with continuous dietary carbohydrate availability, GP activity may not be rate‑limiting for maintaining glucose homeostasis. The observed transient larval weight loss could thus reflect both active metabolic compensation and the inherently low demand on GP for glucose supply in a feeding larva. Distinguishing between active compensation and the non‑limiting nature of GP will require future studies under nutrient‑restricted conditions or during fasting intervals, where GP's role is likely to become more critical.”

Additional revisions to moderate language throughout the manuscript

Beyond the specific revisions addressing each weakness, we have also made the following broader language modifications to ensure consistent and cautious interpretation throughout:

(23) Changes from “ablation” to “suppression” (Lines 180, 195–197, 357):

“Ablation” to “suppression” throughout.

(24) Conclusions – Fundamental principle (Lines 619):

“our findings uncover a fundamental principle of...”

(25) Conclusions – Strategic path (Lines 621–623):

“This metabolic plasticity renders GP non-viable as a standalone insecticidal target but illuminates a potential strategic path forward”

(26) Discussion – Trehalase inhibitors (Lines 556–558):

“For example, trehalase inhibitors such as Validamycin A are potent insecticides [44, 45], suggesting that trehalase inhibition may not be subject to the same degree of metabolic compensation”

(27) Discussion – Metabolic pincer (Line 561):

“a 'metabolic pincer' attack that might overcome adaptive compensation”

(28) MM/GBSA and structural predictions (Lines 27–32, 450, 455–457):

Abstract (Lines 27–32)

“Molecular docking and MM/GBSA analysis predict that this selectivity reflects differential side-chain engagement: GPI is predicted to occupy the allosteric site at the dimer interface via contacts with seven residues spanning both subunits (ΔG = −34.63 kcal/mol), whereas DFB's difluorobenzoyl moiety is predicted to remain solvent-exposed without productive protein contacts (ΔG = −29.29 kcal/mol).”

Line 450

“MM/GBSA analysis indicated”

Lines 455–457

“These structural predictions are consistent with established structure–activity relationships of acyl urea compounds [15], in which target selectivity is governed by the side-chain substitution pattern rather than by the shared acyl urea core”

(29) Scope of conclusion (Lines 471–472):

“This provides the first direct biochemical evidence excluding GP as a candidate molecular target for diflubenzuron”

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

Major recommendations

(1) Moderate the interpretation of the proposed compensatory mechanism throughout the manuscript.

We fully agree with the reviewer's assessment. Throughout the manuscript, we have systematically moderated the language used to describe the compensatory mechanism, ensuring that our interpretations are framed as plausible inferences supported by multiple lines of indirect evidence rather than as established conclusions. The specific revisions are detailed in items (1)–(7), items (10)–(16) of our response to Weakness 1 above, and items (24)–(25) of our response to Weakness 3 above.

Additional language modifications addressing the reviewer's concern about assertive terminology are detailed in our response to Minor Recommendation 1 below.

Line 495

“Our investigation shows that this tolerance reflects a robust”

Line 498

“As insects elicited a compensatory gluconeogenic pathway.”

Line 518

“Direct metabolite quantification reflected this compensation”

(2) Clarify the evidence supporting alternative glycogen degradation pathways.

We thank the reviewer for this valuable suggestion. We have revised the GBE and α-amylase sections to more clearly distinguish between transcriptional changes and functional pathway activation. The revised wording now explicitly acknowledges that our observations are limited to the transcript level and that functional confirmation would require additional experiments. The specific revisions are detailed in items (8)–(9) and item (12) of our response to Weakness 1 above. We have also added a sentence acknowledging that these observations are limited to the transcript level (item 17) of our response to Weakness 2 above.

(3) Discuss alternative explanations for the limited physiological phenotype.

We thank the reviewer for this important point. We agree that the continuous availability of dietary carbohydrates under our experimental conditions represents a plausible alternative explanation for the lack of phenotype. We have incorporated this alternative interpretation into the Abstract, Introduction, and Discussion sections, as detailed in items (3)–(4) of our response to Weakness 1 above, and items (21)–(22) of our response to Weakness 3 above.

We have also revised the subsequent conclusion (Lines 553–554) to align with this alternative interpretation, changing “GP fails as a standalone target due to compensation via gluconeogenesis” to “GP appears to fail as a standalone target, at least in part because of compensation via gluconeogenesis.” This ensures consistency with the preceding discussion of GP’s potentially non‑rate‑limiting role.

Minor recommendations

(1) Review the manuscript for statements using terms such as "demonstrates," "confirms," "proves," or "establishes."

We have systematically reviewed the entire manuscript and replaced overly assertive terms (e.g., “demonstrates,” “confirms,” “proves,” “establishes”) with more cautious language (e.g., “suggests,” “is consistent with,” “provides evidence that,” “indicates”) wherever they refer to the proposed metabolic mechanism. Specific revisions are listed under Major Recommendation 1 above. In addition:

MM/GBSA and structural predictions (Lines 27–32, 450, 455–457): Abstract (Lines 27–32): Changed “Molecular docking and MM/GBSA analysis reveal that...” to “Molecular docking and MM/GBSA analysis predict that...”

Results (Line 450): Changed “MM/GBSA analysis confirmed” to “MM/GBSA analysis indicated.”

Results (Lines 455–457): Changed “These structural data confirm that...” to “These structural predictions are consistent with...”

Conclusion scope (Lines 471–472): Changed “This provides the first direct biochemical evidence excluding GP as a candidate molecular target for BPUs” to “This provides the first direct biochemical evidence excluding GP as a candidate molecular target for diflubenzuron.”

Changes from “ablation” to “suppression” (Lines 180, 195–197, 357): Changed “ablation” to “suppression” throughout the manuscript where referring to GP knockdown.

(2) Ensure that the distinction between changes in transcript abundance, enzyme activity, and metabolic flux is maintained consistently throughout the Results and Discussion.

We have carefully reviewed the Results and Discussion sections to ensure that we consistently distinguish between transcript abundance (measured by RT-qPCR), enzyme activity (measured by activity assays), and inferred metabolic flux (not directly measured). The revisions listed under Major Recommendations 1 and 2 above directly address this point. Key changes include:

Consistently using “transcriptional upregulation” or “expression” when referring to qPCR data, rather than “activation” or “pathway activity.”

Explicitly acknowledging that transcriptional changes do not necessarily reflect metabolic flux.

Using “is consistent with” rather than “demonstrates” when linking gene expression changes to functional outcomes.

Additional clarifications and corrections to data presentation

During the preparation of the revised manuscript, we also made several corrections and clarifications to the data presentation:

RNAi knockdown efficiency (Line 364–367): We have added a clarification that the knockdown efficiency in the cohort used for enzyme activity and fitness assays (54.66% at 48 h) was lower than that achieved in the dose–response experiment (87.59% at 48 h), reflecting batch-to-batch variation between independently injected cohorts. We have also noted that the enzyme-activity data should be interpreted against the transcript reduction measured in this same cohort.

Total protein data comparability (Lines 374–380): We have added a note clarifying that the total-protein data shown in Figure 9D and Figure 10–figure supplement 2 derive from independent experimental cohorts processed at different homogenization ratios; absolute protein concentrations are therefore not directly comparable between the two panels. Both datasets nonetheless show a transient decline of approximately 30% in total protein in dsGP-treated larvae within the first 72 h.

Trehalose and G6P data interpretation (Lines 304–309, 314–315, 345–353): We have revised the description of trehalose and G6P levels at 96 h to clarify that the large fold-differences primarily reflect a pronounced decline in dsGFP control values at this time point, rather than a net increase in absolute metabolite content in dsGP-treated larvae. The revised wording now indicates that GP-suppressed larvae maintain their trehalose and G6P pools at a stage when control larvae are actively depleting them.

Gene expression recovery kinetics (Lines 251–252): We have corrected the description of PxTre and PxHex expression at 96 h to state that they “returned to, and modestly exceeded, control levels” rather than “returned to near baseline levels.”

Cell line description update (Lines 644–654): In response to editorial requirements, we have updated the cell line description to include supplier authentication, mycoplasma testing status, and confirmation that the cells were used experimentally within one year of purchase.

Figure corrections:

Figure 2 legend: revised to state “maximum inhibition did not exceed 55.07 ± 7.05%” to match the main text.

Figure 4 legend: corrected normalization description from “control set to 1.0 at each time point” to “calibrated to the 24 h control sample (set to 1.0).”

Figure 6 legend: corrected reference sample from “L1 set to 1.0” to “Egg set to 1.0.”

Figure 8D: corrected y-axis label from “Adult emergency (%)” to “Adult emergence (%).”

Figure 9C: corrected in-figure label from “G-6-P” to “G-6-Pase.”

Terminology correction: We have standardized the use of “similarity” (rather than “identity”) when referring to sequence comparisons (Lines 190, 1075).

PROSITE motif name correction (Lines 181–184): We have corrected “phosphatase-pyridoxal phosphate linkage site” to “phosphorylase pyridoxal-phosphate attachment site.”

Figure 11 statistical description (Lines 1138–1142): We have revised the figure legend to specify that data in (B) and (C) are mean ± SEM of three biological replicates, while data in (D–F) are shown as box plots with sample sizes indicated, and the statistical comparison is dsGP vs. dsGFP (independent samples t‑test). These revisions do not affect any results or conclusions.

Gene expression peak description (Line 301): Changed “coincided temporally with peak expression of gluconeogenic gene” to “coincided with the significant upregulation of gluconeogenic genes” to avoid implying a single defined peak.

Minor textual corrections

We have also corrected the following textual issues:

Sentence structure (Lines 433–435): We have revised the sentence structure to correct a comma splice and clarify the logical relationship. The original sentence “To provide structural insight into the observed selectivity, GPI potently inhibits PxGP (IC50 = 2.96 nM) while DFB does not, we performed...” has been revised to “To provide structural insight into the observed selectivity—in which GPI potently inhibits PxGP (IC50 = 2.96 nM) while DFB does not—we performed...”

Additional minor corrections: We have corrected a small number of typographical errors and stylistic inconsistencies throughout the manuscript.

All of these corrections are limited to data presentation, figure labeling, and textual clarity. They do not alter any experimental results, quantitative conclusions, or the overall interpretation of the study.

Version 3

Abstract

Glycogen phosphorylase (GP) catalyzes the rate-limiting step of glycogenolysis and occupies a central position in insect carbohydrate metabolism, supplying precursors for chitin biosynthesis. Acyl urea compounds structurally related to benzoylphenylurea (BPU) insecticides are potent inhibitors of mammalian GP, raising the question of whether insect GP could serve as an independent insecticidal target. Here, we systematically evaluate this possibility in the diamondback moth Plutella xylostella. We show that a mammalian GP inhibitor (GPI) potently inhibits recombinant PxGP (IC50 = 2.96 nM), while the BPU insecticide diflubenzuron (DFB) does not. Molecular docking and MM/GBSA analysis predict that this selectivity reflects differential side-chain engagement: GPI is predicted to occupy the allosteric site at the dimer interface via contacts with seven residues spanning both subunits (ΔG = −34.63 kcal/mol), whereas DFB’s difluorobenzoyl moiety is predicted to remain solvent exposed without productive protein contacts (ΔG = −29.29 kcal/mol). Despite potent in vitro inhibition, GPI exhibits no insecticidal activity, and RNAi-mediated knockdown of PxGP (confirmed by enzyme activity measurements showing ∼29% reduction in per-larva GP-a activity) does not impair development or survival. We provide evidence that insects compensate through a multi-layered metabolic response: upregulation of gluconeogenic enzymes (PEPCK, G-6-Pase), selective upregulation of glycogen branching enzyme (GBE) but not α-amylase, and changes in protein content suggestive of catabolic substrate mobilization. Fitness assessment reveals transient larval weight loss (24–48 h) with complete recovery of pupal weight, adult wing morphology, and female fecundity, indicating that the metabolic compensation response is ultimately effective in sustaining development. These findings suggest that GP is functionally non-essential under these conditions, likely through a combination of gluconeogenic compensation and the availability of alternative carbon sources, and highlight the importance of considering metabolic network redundancy in target-based insecticide design.

Introduction

Benzoylphenylurea (BPU) insecticides, including diflubenzuron (DFB), have been widely used for pest control for over four decades. The molecular basis of BPU-mediated insecticidal activity has been the subject of extensive investigation. Genetic studies, most notably the demonstration through CRISPR/Cas9 gene editing that amino acid substitutions at a single site in chitin synthase (CHS) confer high-level resistance to DFB in Drosophila melanogaster [1], provide compelling evidence that CHS is the primary target gene associated with BPU resistance. These findings establish a clear causal genetic link between CHS and BPU activity.

Despite this genetic evidence, the biochemical mechanism by which BPUs inhibit CHS remains incompletely characterized. Multiple in vitro studies of chitin synthase, including those using cell-free chitin synthase preparations, have failed to demonstrate direct enzymatic inhibition by BPU compounds: DFB did not inhibit Tribolium gut chitin synthetase [2], and a systematic evaluation concluded that chitin synthesis-inhibiting insect growth regulators do not inhibit chitin synthase in cell-free assays [3, 4]. More recently, Zhang and Zhu (2013) reported only slight in vitro inhibition of Anopheles gambiae CHS activity by DFB at the highest concentration tested, with no in vivo inhibition detected [5]. These findings led to the recognition that a key obstacle in identifying the action site of BPU insecticides has been the lack of in vitro demonstration of their ability to inhibit insect chitin synthesis under cell-free conditions. This gap between genetic and biochemical evidence does not diminish the importance of the genetic findings, but it does highlight the need for continued investigation at the enzyme level [6, 7].

Glycogen Phosphorylase (GP) catalyzes the rate-limiting step in glycogenolysis and plays a central role in insect carbohydrate metabolism, providing glucose-1-phosphate that feeds into the hexosamine pathway and ultimately supplies UDP-N-acetylglucosamine for chitin biosynthesis [8, 9]. In insects, GP-derived glucose is converted to trehalose (the primary hemolymph sugar), which is subsequently channeled through a series of enzymatic steps to generate UDP-GlcNAc, the substrate monomer for chitin synthesis. GP also plays key roles in carbohydrate homeostasis, osmotic regulation, stress adaptation [10–13]. Acyl urea compounds structurally related to BPUs have been identified as potent inhibitors of mammalian GP (Figure 1) [14, 15], raising the question of whether insect GP could itself represent a viable insecticidal target — independent of the BPU mechanism-of-action question. However, the feasibility of targeting insect GP for pest control has not been systematically evaluated.

Structural comparison of benzoylphenylurea (BPU) insecticides and a mammalian glycogen phosphorylase inhibitor (GPI) reveals a shared acylurea scaffold.

Chemical structures of four BPU insecticides (Diflubenzuron, Teflubenzuron, Novaluron, and Flufenoxuron) and a human glycogen phosphorylase inhibitor (GPI) are shown. All share a central N,N’-diphenylurea (acylurea) core (dashed boxes). BPUs are characterized by halogenated (F, Cl) aromatic substituents and in some cases by additional ether or alkyl side chains. The GPI contains a methoxyphenyl group bearing an additional methylurea side chain.

In this study, we address two distinct questions. First, we test whether DFB directly inhibits Plutella xylostella GP (PxGP) at the enzyme level, adding biochemical data to the existing genetic framework. Second, and more importantly, we systematically evaluate GP as an independent insecticidal target by characterizing the metabolic consequences of GP inhibition and knockdown in P. xylostella. Our findings reveal that PxGP is functionally non-essential for larval development under the tested conditions, reflecting a previously uncharacterized gluconeogenic compensation mechanism, providing new insight into the metabolic plasticity of insects and its implications for target-based insecticide design.

Results

Recombinant P. xylostella glycogen phosphorylase (PxGP) is a functional, activatable enzyme

To investigate its biochemical properties and evaluate its potential as an insecticidal target, the full-length PxGP cDNA (predicted molecular mass of 98.1 kDa) was cloned and expressed in Sf9 cells using the Bac-to-Bac baculovirus expression system. Following purification via immobilized metal-affinity chromatography (IMAC) of the N-terminal 6×His tagged protein [16], SDS-PAGE analysis and Western blot analysis confirmed a highly pure protein band migrating at approximately 100 kDa, consistent with its theoretical mass (Figure 2A). A total yield of 7.5 mg of purified protein was obtained from 8.25 g of cell pellet (specific yield=0.91 mg/g cells).

Recombinant PxGP is inhibited by a mammalian GP inhibitor (GPI) but not by the benzoylphenylurea Diflubenzuron.

(A) Purification of 6×His-tagged PxGP from Sf9 cells. Left panel: Coomassie-stained SDS-PAGE. Right: Western blot with anti-His antibody. The expected ∼100 kDa band is indicated (red box). M, marker; S, soluble fraction; P, pellet; F, flow-through; lanes 1–4, imidazole elution fractions. (B) Dose-response inhibition by GPI. Activated PxGP-a activity was measured after pre-incubation with GPI (0.16–500 nM). Data (mean ± SEM, n=3) were fit to a four-parameter logistic curve. The calculated IC50 is 2.96 nM. Inset: GPI structure with its characteristic asymmetric side chains (orange). (C) Inhibition assay with DFB (0.54–1700 nM). DFB did not yield a complete sigmoidal inhibition curve; maximum inhibition did not exceed 55.07 ± 7.05% at the highest concentration. Inset: DFB structure with its characteristic halogenated benzoyl groups (orange). This rules out PxGP as a direct target of DFB.

Glycogen phosphorylase activity is regulated by reversible phosphorylation, existing as a less-active dephosphorylated b form (PxGP-b) and a fully active phosphorylated a form (PxGP-a) [17]. Recombinant PxGP-b was activated by incubation with immobilized phosphorylase kinase (PhK) in the presence of ATP and Mg2+ for 2 h [18]. The activated enzyme showed robust catalytic activity in a coupled-enzyme spectrophotometric assay, with NADPH production (monitored by increase in A340) increasing linearly for the first 15 min and reaching a plateau by approximately 20 min, consistent with substrate depletion and/or product accumulation. The specific activity of activated PxGP-a was 294.25 ± 3.45 U/mg protein (Figure 2–figure supplement 1A), confirming that the recombinant enzyme is catalytically competent for inhibitor screening.

GPI potently inhibits PxGP in vitro and in vivo

Given the structural similarity between insecticidal BPUs and mammalian GPIs [15] (Figure 1), we tested whether PxGP could be a target of acylurea-scaffold compounds. GPI exhibited potent, dose-dependent inhibition of recombinant PxGP-a, with a sigmoidal dose-response curve typical of competitive or allosteric inhibition (Figure 2B). Non-linear regression analysis yielded an IC50 value of 2.96 nM (95% confidence interval [CI]: 1.66–5.35 nM, Hill slope = 0.74, R² = 0.877, n = 3). Near-complete inhibition (>95%) was achieved at 500 nM GPI, indicating saturable, high-affinity binding. The potency of GPI against PxGP is comparable to its effect on mammalian orthologs (see below) [15].

In striking contrast, the BPU insecticide Diflubenzuron (DFB), which shares the acylurea core with GPI (Figure 1), did not effectively inhibit PxGP-a. No standard sigmoidal dose-response curve was observed over 0.54 nM to 1700 nM (Figure 2C). Maximum inhibition observed did not exceed 55.07 ± 7.05% (mean ± SEM), with high variability at lower concentrations. These data indicate that DFB, unlike GPI, is not an effective inhibitor of PxGP, ruling out GP as the direct molecular target of this BPU insecticide.

Consistent with the recombinant enzyme results, GPI effectively inhibited native GP activity in crude lysates from third-instar larvae in a dose- and time-dependent manner (Figure 2–figure supplement 1B). After 45 min of pre-incubation, 10 µM GPI reduced native GP activity by 57.52 ± 1.88% (n = 3). DFB, however, showed no significant inhibitory effect on native GP activity at any concentration tested (10-7, 10- 6, or 10-5 M) or time point (5, 10, 20, or 30 min), with activity remaining at 97-108% of control (Figure 2–figure supplement 1C).

Collectively, these results establish PxGP as a high-affinity molecular target for GPI, whereas DFB does not inhibit PxGP.

Paradoxical finding: potent PxGP inhibition by GPI lacks in vivo toxicity

Because GPI potently inhibits PxGP, the rate-limiting enzyme in glycogenolysis, which supplies precursors for chitin synthesis [19], we expected it to exhibit significant insecticidal activity. In leaf-dip bioassays with third-instar P. xylostella larvae continuously exposed to GPI (250 or 500 mg/L) or DFB (125 or 250 mg/L), GPI caused no significant toxicity or developmental disruption across the entire time course (24 to 120 h post-treatment) (Figure 3A). At 120 h, mortality with 500 mg/L GPI was only 12.78 ± 2.42% (n = 3), which was not significantly different from the solvent control (CK: 5.00 ± 5.00%, P = 0.21) (Figure 3B). Larvae exposed to 250 mg/L GPI exhibited even lower mortality (2.50 ± 2.50%).

The GP inhibitor (GPI) is non-toxic to Plutella xylostella, unlike the benzoylphenylurea Diflubenzuron (DFB).

(A) Developmental stage distribution over time for control, GPI (250, 500 mg/L), and DFB (125, 250 mg/L) treated larvae. DFB causes developmental arrest. (B) Quantitative effects on mortality (top), pupation (middle), and adult emergence (bottom). GPI groups did not differ from control (P > 0.05), whereas DFB caused significant, dose-dependent mortality and severely impaired development (P < 0.05). Data are mean ± SEM (n=3). Different letters denote statistical significance. (C) Larval phenotypes at 96 h. Control and GPI-treated larvae are normal. DFB-treated larvae show catastrophic molting failure and “double cuticle” (red arrows). Scale bar, 1 mm. (D) Pupal and adult phenotypes. GPI-treated individuals develop normally. DFB treatment results in deformed pupae and trapped pharate adults (red arrows). Scale bar, 1 mm.

In stark contrast, DFB induced significant, dose-dependent mortality. At 120 h, DFB treatment at 125 mg/L caused 41.60 ± 4.85% mortality (P < 0.01 vs. control) and at 250 mg/L caused 61.36 ± 10.02% mortality (P < 0.01 vs. control). Probit analysis of the DFB dose-response data yielded calculated LC30 and LC50 values of 141.900 mg/L (95% CI: 67.893–206.576 mg/L) and 229.537 mg/L (95% CI: 145.287-331.784 mg/L), respectively, with an acceptable goodness-of-fit (χ² = 11.946, P = 0.683) (Supplementary file 2).

Furthermore, GPI-treated larvae showed no developmental abnormalities. Pupation rates were 64.17 ± 4.79% for the 500 mg/L GPI group and 69.72 ± 6.81% for the 250 mg/L group, which were statistically indistinguishable from the control group (64.72 ± 9.33%, P > 0.05 for both comparisons) (Figure 3B, middle panel). Similarly, adult emergence rates were 56.11 ± 5.64% and 56.39 ± 6.24% for the 500 and 250 mg/L GPI groups, respectively, compared to 54.10 ± 11.91 % for the control (P > 0.05, Figure 3B). Phenotypically, GPI-treated larvae molted normally, fed actively, and displayed no cuticle defects (Figure 3C). DFB-treated larvae exhibited the characteristic phenotypes of BPU intoxication: catastrophic molting failure, larval-pupal intermediates, and complete failure of adult emergence (Figure 3C and 3D). This striking discrepancy between biochemical potency and in vivo inefficiency presents a profound paradox that demands mechanistic investigation.

GPI exposure triggers a compensatory upregulation of PxGP transcription

To explore this discrepancy, we measured PxGP gene expression after GPI exposure. RT-qPCR revealed that 500 mg/L GPI rapidly upregulated PxGP mRNA transcription. As shown in Figure 4, its relative expression increased 3.24-fold at 24 h, peaked at 3.48-fold at 48 h, and remained elevated at 96 h. This indicates that the insect responds to the pharmacological inhibition by transcriptionally increasing the de novo synthesis of the target enzyme.

Ingestion of GP inhibitor (GPI) induces compensatory upregulation of PxGP transcription.

Relative PxGP mRNA levels in larvae after continuous exposure to 500 mg/L GPI or control (CK). Expression was normalized to PxRPS13 and calibrated to the 24 h control sample (set to 1.0) using the 2-ΔΔCt method. Data are mean ± SEM (n=3). GPI triggered a rapid and sustained transcriptional increase, peaking at 48 h (3.48-fold). Significance vs. control: *P < 0.05, **P < 0.01, ***P < 0.001 (two-way ANOVA with Sidak’s test).

Sequence homology and rationale for target site selection

To achieve effective gene knockdown coupled with functional suppression, precise selection of the interference target site is critical. Sequence analysis of PxGP (837 residues) revealed that residues 88-680 form the active site pocket of the GT35_Glycogen_Phosphorylase family, and the C-terminal region contains a phosphorylase pyridoxal-phosphate attachment site (residues 672-684). Eleven residues within positions 42-318 are conserved and correspond to the AMP-binding site in human GP (Figure 5A) [20].

PxGP sequence conservation and domain architecture.

(A) Schematic of PxGP functional domains: pyridoxal phosphate (PLP) binding site (blue boxes), AMP allosteric site (yellow boxes), and glycogen-binding/active site pocket (blue shading). (B) Multiple sequence alignment of GP from Plutella xylostella (PxGP), other lepidopterans (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP). Identical residues are shaded red. The gray box indicates the 606 bp region (aa 22–224) targeted for RNAi, which overlaps the conserved AMP-binding domain. PxGP shares 82.29% similarity with human GP.

Multiple sequence alignment of PxGP with GP orthologs from other key lepidopteran pests (Helicoverpa armigera HaGP, Spodoptera exigua SeGP, Ostrinia furnacalis OfGP), human (HsGP), and rabbit (OcGP) demonstrated high conservation across species (Figure 5B). PxGP shares 82.29% sequence similarity with human GP (Figure 5–figure supplement 1). Since GP activity is allosterically regulated by AMP binding and phosphorylation at Ser14, triggering a conformational shift from the inactive T-state GPb (unphosphorylated) to the active R-state GP-a (phosphorylated) [17], we targeted the core AMP-binding site region (gray box in Figure 5B) for RNAi. Theoretically, interference in this region should prevent AMP-mediated activation of existing GPb while degrading the mRNA pool, substantially reducing de novo GP protein synthesis and maximizing functional suppression.

RNAi confirms PxGP is non-essential for larval development

The compensatory upregulation of PxGP mRNA expression in response to pharmacological inhibition (GPI exposure experiments) indicated that the gene is under active transcriptional regulation and suggested that increased enzyme synthesis might partially mitigate the effects of GPI. However, this compensation alone could not fully explain the complete absence of phenotype. To determine whether a more profound loss-of-function—achieved by directly suppressing PxGP gene expression—is lethal, we performed RNAi to knock down PxGP at the transcriptional level.

Firstly, we quantified the expression profile of PxGP across developmental stages in P. xylostella, PxGP was highly expressed in fourth-instar larvae (L4) and adults (Figure 6), stages that correlate with the energy demands for metamorphosis and flight, respectively [8, 21]. Given the inherent time lag required for systemic dsRNA delivery to achieve effective post-transcriptional knockdown [22], dsRNA was administered during the mid-third-instar larval stage. Mid-third-instar larvae were microinjected with dsRNA targeting a conserved 606 bp region of PxGP (dsGP) or with a non-specific control dsGFP. Three concentrations of dsGP (6,000, 10,000, and 14,000 ng/µL) were tested, and knockdown efficiency was assessed by RT-qPCR at 24, 48, 72, and 96 h post-injection (Figure 7). The highest concentration (14,000 ng/µL) achieved maximal suppression at 48 h, reducing PxGP transcript levels to 12.47 ± 1.68% of the dsGFP control (87.59% knockdown, P < 0.001) (Figure 7B).

Developmental expression profile of PxGP.

Relative PxGP mRNA levels across life stages (Egg, larval instars L1-L4, Pupa, Adult), normalized to PxRPS13 (Egg set to 1.0). Data are mean ± SEM (n=3). Expression peaks in adults. Different letters indicate significant differences (P < 0.05, one-way ANOVA with Tukey’s test).

RNAi-mediated knockdown of PxGP is dose- and time-dependent.

(A) Experimental workflow. Third-instar larvae were injected with dsRNA targeting PxGP (dsGP) at three concentrations or a control dsRNA (dsGFP), then sampled for analysis. (B) Time-course of PxGP mRNA levels post-injection, relative to dsGFP control (set to 1.0). The highest dsGP dose (14,000 ng/µL) caused maximal suppression (87.6%) at 48 h. Data are mean ± SEM (n=3). Significance vs. dsGFP at each time point: * P < 0.05, **P < 0.01, ***P < 0.001 (one-way ANOVA with Tukey’s test).

Significant knockdown persisted at 72 h (69.86% suppression, P < 0.05), with partial recovery by 96 h, likely due to dsRNA degradation and/or compensatory transcriptional upregulation. The 10,000 ng/µL dose also produced significant knockdown at 48 h (65.76% suppression, P < 0.001), whereas the 6,000 ng/µL dose showed more modest effects (Figure 7B).

Despite this robust, transient silencing, no adverse phenotypic effects were observed. Developmental stage distribution and mortality rates in all dsGP-treated groups (6,000, 10,000, and 14,000 ng/µL) were statistically indistinguishable from the dsGFP control group across the entire 120 h observation period (P > 0.05 at all time points) (Figure 8A and 8B). At 120 h, cumulative mortality was 8.10 ± 4.23%, 9.82 ± 0.81%, and 3.33 ± 3.33% for the 6,000, 10,000, and 14,000 ng/µL dsGP groups, respectively, comparable to the dsGFP control (Figure 8B). Pupation and adult emergence were also indistinguishable from controls (Figure 8B). Phenotypic examination revealed that dsGP-injected larvae developed normally, molted successfully, formed morphologically normal pupae, and emerged as viable adults without cuticle defects or developmental delays (Figure 8C and 8D). Thus, acute, profound suppression of PxGP expression is not lethal to P. xylostella during the critical larval-pupal transition, despite the enzyme’s established role in glycogenolysis and chitin precursor synthesis. This finding necessitates an investigation into the compensatory mechanisms that allow the insect to survive in the absence of functional GP.

PxGP knockdown does not impair development or survival.

(A) Developmental stage distribution post-injection shows similar progression across all groups. (B) Cumulative mortality (top), pupation (middle), and adult emergence (bottom) rates were not significantly different between dsGP and dsGFP control groups (P > 0.05). Data are mean ± SEM (n=3). (C) Representative larval phenotypes at 48 h and 96 h post-injection are normal in both dsGFP and high-dose dsGP groups. Scale bar, 1 mm. (D) Pupae and adults from both groups are morphologically normal. Scale bar, 1 mm.

PxGP knockdown coordinately suppresses the glycogenolysis-to-chitin pathway

To elucidate how larvae survive GP loss, we examined downstream genes linking glycogenolysis to chitin synthesis. Specifically, we measured the expression of trehalase (PxTre), which hydrolyzes circulating trehalose to glucose [8], and hexokinase (PxHex), which phosphorylates glucose to glucose-6-phosphate—a key branch point leading to UDP-GlcNAc synthesis for chitin production [23, 24].

RT-qPCR analysis revealed that suppression of PxGP triggered a coordinated transcriptional downregulation of both downstream genes (Figure 9A and 9B). At 48 h post-injection, coinciding with the peak of PxGP knockdown (87.59% suppression), PxTre expression was reduced to 30.46 ± 1.25% of the dsGFP control (69.54% suppression, P < 0.001), and PxHex to 22.82 ± 3.39% of control (77.18% suppression, P < 0.001). This suppression persisted at 72 h (PxTre: 61.27 ± 2.21% suppression, P < 0.001; PxHex: 41.96 ±4.02% suppression, P < 0.001) and by 96 h had returned to, and modestly exceeded, control levels, mirroring the recovery kinetics of PxGP itself.

PxGP knockdown triggers biphasic metabolic reprogramming, culmina ting in gluconeogenic compensation.

(A, B) Concurrent suppression of downstream genes Trehalase (PxTre) and Hexokinase (PxHex) post-knockdown. (C) Upregulation of gluconeogenic genes PxPEPCK and PxG-6-Pase at 96 h. Data in A-C are mean ± S EM (n=3); *P < 0.05, **P < 0.01, ***P < 0.001 vs. dsGFP control. (D-H) Metabolite levels post-knockdown. (D) Total protein decreased at 72 h (P < 0.05). (E) Glycogen remained stable. (F) Trehalose levels were elevated 7.4 fold relative to control at 96 h (P < 0.05). (G) Glucose-6-phosphate (G6P) increased 6.5-fold at 96 h (P < 0.0 5). (H) Free glucose showed a transient dip at 72 h (P < 0.05) before recovery. Data ar e mean ± SEM (n=3); *P < 0.05 vs. control at each time point.

This coordinated downregulation indicates that the insect responds to the loss of GP not by futilely attempting to maintain flux through a pathway that lacks its key upstream enzyme, but rather by implementing a broader transcriptional shutdown of the glycogenolysis-to-glucose-to-chitin pathway. This suggests a coherent metabolic response program, raising the question: if glycogenolysis is blocked, how does the insect obtain the glucose required for survival and chitin synthesis?

Biphasic metabolic response: transient depletion followed by gluconeogenic rescue

The absence of a lethal or developmental phenotype despite glycogenolysis pathway suppression suggested activation of alternative glucose synthesis pathway to compensate for GP loss. We therefore measured expression of key gluconeogenic enzymes and metabolite levels across the glycogeno-glucose axis.

Gene expression analysis is consistent with gluconeogenic activation

We assessed transcript levels of two rate-limiting gluconeogenic enzymes: phosphoenolpyruvate carboxykinase (PEPCK, encoded by PxPEPCK), which catalyzes the conversion of oxaloacetate to phosphoenolpyruvate; glucose-6-phosphatase (G-6-Pase, encoded by PxG-6-Pase), which catalyzes the dephosphorylation of glucose-6-phosphate to free glucose. These enzymes are reciprocally regulated relative to glycolysis/glycogenolysis and serve as molecular markers of gluconeogenic flux [25, 26].

At 96 h post-injection―a critical pre-pupal stage with high carbohydrate demand for pupal cuticle synthesis [27, 28]―both PxPEPCK and PxG-6-Pase were significantly upregulated in dsGP-treated larvae (3.95 ± 0.44-fold and 3.34 ± 0.20-fold versus dsGFP control, respectively; P < 0.001 for each) (Figure 9C). This indicates robust transcriptional activation of the gluconeogenesis pathway in response to GP loss.

Changes in protein content suggestive of catabolic substrate mobilization

To identify the carbon source for gluconeogenesis, we quantified total protein content in dsGP-treated larvae across the time course. Strikingly, a significant 29.39% reduction in protein was observed at 72 h post-injection (P < 0.05), coinciding precisely with the nadir of glucose insufficiency (Figure 9D).

This protein decline is consistent with substrate mobilization: the degradation of cellular proteins liberates amino acids, which serve as the primary carbon source for gluconeogenesis in insects [25, 29–31]. The magnitude of protein loss (∼30%) is sufficient to provide substantial amino acid flux into the gluconeogenic pathway.

Metabolite profiling reveals a biphasic adaptation

Quantification of four key metabolites (glycogen, trehalose, glucose-6-phosphat [G6P], and free glucose) along the glycogen-glucose axis at 48, 72, and 96 h post-injection uncovered a dynamic, biphasic response (Figure 9E-9H).

Glycogen Content: Contrary to the expectation of accumulation upon GP blockade, glycogen levels remained unchanged relative to controls throughout the time course (P > 0.05) (Figure 9E). This paradox suggests that insects activate alternative, GP-independent pathways for glycogen catabolism―most likely via α-amylase or glycogen debranching enzymes [32–34]―to provide emergency glucose mobilization during the period of metabolic stress, suggesting that these compensatory pathways are sufficient for short-term energy needs but are eventually superseded by gluconeogenic glucose production.

Trehalose: Trehalose levels remained stable at 48-72 h but increased substantially by 96 h (7.44-fold elevation, P < 0.05) (Figure 9F). This increase coincided with the significant upregulation of gluconeogenic genes (e.g., PEPCK and G-6-Pase), consistent with the idea that gluconeogenesis contributes to trehalose synthesis and ensures adequate carbohydrate reserves for the upcoming pupation [35]. However, it should be noted that the large fold change at 96 h primarily reflects a pronounced decline in trehalose levels in the dsGFP control group at this time point (Figure 9F), rather than a net increase in absolute trehalose content in dsGP treated larvae. Thus, rather than indicating accumulation above the initial baseline, the data suggest that GP suppressed larvae maintain their trehalose pool at a time when controls are actively depleting it—a pattern consistent with sustained gluconeogenic input.

G6P: G6P, which sits at the intersection of glycolysis, gluconeogenesis, the pentose phosphate pathway, and glycogen synthesis [36], exhibited a pattern mirroring that of trehalose (Figure 9G). G6P content declined gradually from 48 to 72 h, consistent with reduced glucose mobilization from glycogen, but then diverged at 96 h, with dsGP-treated larvae maintaining higher levels (6.49-fold elevation, P < 0.05). This relative elevation is consistent with the pattern observed for trehalose. Together, the coordinated behavior of G6P and trehalose is consistent with gluconeogenesis-derived glucose being converted into storage and transport carbohydrates.

Free Glucose: Free glucose levels showed the most subtle but perhaps most physiologically informative pattern (Figure 9H). At 48 h―coinciding with peak PxGP suppression―glucose content was statistically indistinguishable from controls (P > 0.05), indicating that initial homeostatic mechanisms (including GP-independent glycogen breakdown and dietary carbohydrate absorption) successfully buffered against acute GP loss. However, at 72 h, the nadir of the metabolic stress response, free glucose levels declined significantly, albeit modestly (18.15% reduction, P < 0.05). This transient dip reveals a critical window of metabolic insufficiency during which the insect’s compensatory mechanisms are insufficient to fully maintain glucose homeostasis. Importantly, by 96 h, glucose levels fully recovered to control values (P > 0.05), demonstrating that the gluconeogenic pathway successfully restores glucose homeostasis before critical developmental transitions occur.

Integrated interpretation

Collectively, these metabolite and gene expression data reveal a biphasic metabolic adaptation that provides a coherent explanation for why substantial GP suppression causes no mortality or developmental defects:

Phase I (emergency compensation, 48-72 h): Metabolic stress and emergency compensation during the initial response to PxGP suppression, the insect experiences genuine metabolic stress, evidenced by the glycogen levels neither accumulated nor were depleted and the transient reduction in free glucose at 72 h. The static glycogen content indicates activation of GP-independent catabolism pathways, which provide emergency glucose but are insufficient for complete homeostatic maintenance. The modest glucose dip at 72 h demonstrates that this stress is real but tolerable, likely due to several factors: (i) the magnitude is modest (only 18.15% reduction), (ii) the duration is brief (∼24 h), (iii) the timing coincides with the mid-L4 larval stage when energy demands are moderate and active feeding allows dietary carbohydrate supplementation, and (iv) trehalose stores (which remain stable during this phase) can be mobilized to buffer hemolymph glucose.

Phase II (gluconeogenic rescue, 96 h): Robust transcriptional upregulation of PEPCK and G-6-Pase (3-4-fold increases) was observed under GP suppression, which is consistent with enhanced gluconeogenic capacity. At 96 h, trehalose and G6P levels in dsGP treated larvae were maintained at higher levels than those in the dsGFP controls, which had declined markedly by this time point. This relative maintenance suggests that larvae under GP suppression sustain their carbohydrate pools at a stage when control larvae are actively depleting them. Such maintenance may help ensure adequate carbohydrate supply for the energetically demanding process of pupal cuticle synthesis, which begins shortly after 96 h.

This substrate-to-product chain for gluconeogenesis-mediated metabolic rescue— protein degradation → amino acid release → gluconeogenic conversion → glucose synthesis → trehalose storage—constitutes a complete compensatory mechanism that sustains development despite substantial GP suppression.

RNAi-mediated suppression of PxGP reduces enzyme activity in vivo

To confirm that RNAi-mediated transcript suppression of PxGP translates to reduced enzyme function in vivo, we measured glycogen phosphorylase a (GP-a) activity in crude extracts from dsGP-treated and dsGFP-treated larvae at 24, 48, 72, and 96 h post-injection using a coupled-enzyme spectrophotometric assay (Figure 10A, B). At 48 h post-injection in this cohort, PxGP mRNA levels were reduced by 54.66% (Figure 10–figure supplement 1). This is lower than the 87.59% suppression achieved in the dose–response experiment (Figure 7B), reflecting batch-to-batch variation in RNAi efficiency between independently injected cohorts.

Effects of PxGP knockdown on GP enzyme activity and the expression of glycogen metabolism-related genes.

(A) GP enzyme activity normalized to sample protein concentration. A significant decrease in activity was detected only at 48 h (P < 0.01). (B) GP enzyme activity per insect. Significant decreases were detected at 24 and 48 h (*P < 0.05, **P < 0.01). (C) Expression levels of the GBE gene at different time points after RNAi treatment. (D) Expression levels of the α-Amylase gene at different time points after RNAi treatment. All data are presented as mean ± SEM (n = 3); *P < 0.05, **P < 0.01, ***P < 0.001, dsGFP vs. dsGP treatment groups. An independent samples t-test was used in (A) and (B), and two-way ANOVA with Sidak’s test was used in (C) and (D).

When GP activity was normalized to total protein content, a significant reduction was detected only at 48 h (−10.35%, P < 0.01; Figure 10A). However, total protein concentration itself declined substantially—by 30.78% at 24 h—before gradually recovering over subsequent time points (Figure 10–figure supplement 2). This concurrent protein decline explains the modest per-protein activity reduction: as the protein pool shrinks, the specific activity calculation is inflated even if total enzyme function is diminished. Note that the total-protein data shown in Figure 9D and in Figure 10–figure supplement 2 derive from independent experimental cohorts processed at different homogenization ratios; absolute protein concentrations are therefore not directly comparable between the two panels, and the time point at which a significant reduction was detected differs (72 h in Figure 9D; 24 h in Figure 10– figure supplement 2). Both datasets nonetheless show a transient decline of approximately 30% in total protein in dsGP-treated larvae within the first 72 h.

To address this, we also calculated GP activity on a per-larva basis. Per-larva GP-a activity was significantly reduced at both 24 h (−29.13%, P < 0.01) and 48 h (−29.28%, P < 0.05; Figure 10B), confirming functional suppression of GP at the protein level. Activity partially recovered at 72 and 96 h, consistent with the transcript recovery pattern observed in our previous RNAi experiments.

The 30.78% reduction in total protein at 24 h post-injection provides independent biochemical evidence supporting protein catabolism—consistent with the mobilization of amino acids to fuel gluconeogenesis as a compensatory response to GP suppression.

PxGP knockdown selectively upregulates glycogen branching enzyme expression

To investigate whether alternative glycogen metabolic pathways are activated following GP suppression, we measured the expression of glycogen branching enzyme (GBE) and α-amylase by RT-qPCR in dsGP-treated versus dsGFP-treated larvae at 24, 48, 72, and 96 h post-injection (Figure 10C, D).

GBE, which catalyzes the formation of branch points in glycogen to increase its solubility and accessibility for degradation [37], was significantly upregulated at 24 h (+29.24%, P < 0.05), 48 h (+16.78%, P < 0.05), and 96 h (+44.46%, P < 0.001; Figure 10C). In contrast, α-amylase, a glycoside hydrolase that catalyzes hydrolysis of α-(1,4)-glycosidic bonds [38], showed no significant change at any time point (Figure 10D).

This differential expression pattern—selective GBE upregulation with unchanged α-amylase expression—indicates that the compensatory response to GP suppression involves targeted remodeling of glycogen structure rather than a generalized upregulation of all glycogen-degrading enzymes. The enhanced glycogen branching may facilitate alternative routes of glycogen mobilization even when GP-mediated phosphorolysis is impaired. However, as these observations are limited to the transcript level, further enzymatic activity assays or glycogen structure analyses would be required to determine whether this transcriptional change translates into functional alterations in glycogen mobilization.

PxGP knockdown causes transient fitness costs that are fully compensated before pupation

To determine whether the metabolic changes observed following PxGP knockdown are associated with measurable physiological consequences, we assessed multiple fitness parameters following PxGP knockdown (Figure 11A–F).

Assessment of fitness cost of PxGP gene silencing by RNAi in Plutella xylostella.

(A) Schematic diagram of fitness cost evaluation at different developmental stages of P. xylostella after dsRNA injection. (B) Food consumption per 10 larvae at different time points. (C) Average body weight per larva at different time points. (D) Pupal weight on day 3 of pupation (n = 21). (E) Adult wing area (n = 20–21). (F) Number of eggs laid by females within 3 days (n = 10–11). Data in (B) and (C) are mean ± SEM of three biological replicates. Data in (D–F) are shown as box plots (box, interquartile range; horizontal line, median; individual points, single insects), with sample sizes as given above for each panel. *P < 0.05, dsGP vs. dsGFP treatment group (independent samples t-test).

Feeding rate was not significantly different between dsGP and dsGFP groups at any time point from 24 to 120 h post-injection (Figure 11B), confirming that the metabolic changes observed are independent of food intake. Larval body weight was significantly reduced in the dsGP group at 24 h (−29.10%, P < 0.05) and 48 h (−25.38%, P < 0.05; Figure 11C), indicating a transient but measurable physiological cost associated with metabolic compensation. However, pupal weight measured on day 3 after pupation showed no significant difference between groups (Figure 11D), demonstrating that larvae fully recover from the transient weight deficit before the pupal transition. Adult wing area (Figure 11E, Figure 11–figure supplement 1 and 2) and female 3-day fecundity (Figure 11F) were also not significantly affected.

This pattern—transient larval weight loss followed by complete recovery of pupal weight, adult morphology, and reproductive performance—is consistent with our metabolic compensation model: GP suppression likely triggers protein catabolism to supply amino acids for gluconeogenesis, contributing to transient weight loss. In the presence of continuous dietary carbohydrate supply, the compensatory response appears sufficiently effective to restore metabolic homeostasis before developmentally critical transitions.

Molecular docking reveals the structural basis for selective inhibition of PxGP by GPI

To provide structural insight into the observed selectivity—in which GPI potently inhibits PxGP (IC50 = 2.96 nM) while DFB does not—we performed molecular docking and binding free energy analysis using a PxGP homology model (Figure 12, Table 1).

Binding mode analysis of GPI and DFB with PxGP.

(A) 3D binding model of PxGP with GPI. (B, C) The key intermolecular interactions for GPI (B) and DFB (C). 2D interaction profile analysis of GPI (D) and DFB (E) with PxGP.

The PxGP homodimer was modeled using SWISS-MODEL with the human liver GP–acyl urea co-crystal structure (PDB: 2ATI; [15]) as the template. Molecular docking and MM/GBSA calculations were performed using Cresset Flare V11.

GPI was predicted to bind at the allosteric site located at the dimer interface, engaging seven residues across both subunits: Asn44 and Val45 from subunit A, and Trp67, Gln71, Tyr75, Arg193, and Asp227 from subunit B (Figure 12B, D). Three strong hydrogen bonds (≤ 2.8 Å) were formed with Asp227. This cross-subunit binding mode is consistent with the experimentally determined binding site of acyl urea inhibitors in the mammalian GP crystal structure [15].

DFB contacted six residues (Val45 from chain A; Ile68, Gln72, Tyr155, Arg193, and Arg242 from chain B) and formed two hydrogen bonds with Arg193 (2.1 and 2.7 Å; Figure 12C, E). Critically, DFB’s difluorobenzoyl moiety remained entirely solvent-exposed without productive protein contacts, explaining its inability to achieve effective target engagement.

MM/GBSA analysis indicated GPI’s substantially more favorable binding (ΔG = −34.63 vs. −29.29 kcal/mol; ΔΔG = −5.34 kcal/mol; Table 1). Energy decomposition revealed van der Waals interactions as the primary driver of selectivity (ΔVDW = −11.49 kcal/mol), consistent with the superior shape complementarity of GPI within the dimer interface pocket.

These structural predictions are consistent with established structure–activity relationships of acyl urea compounds [15], in which target selectivity is governed by the side-chain substitution pattern rather than by the shared acyl urea core.

Discussion

Biochemical evidence excluding GP as a DFB target

A central challenge in mechanism-based insecticide development is distinguishing potent in vitro biochemical inhibitors from physiologically effective in vivo toxicants. This study addresses this challenge by investigating the debated molecular target of BPU insecticides. For over 40 years, the field has contended with conflicting evidence: while genetic studies implicated chitin synthase [1], direct biochemical assays consistently failed to validate it as a direct BPU target [39–41]. It has been hypothesized that BPUs might inhibit insect GP, thereby limiting glucose-derived precursors for chitin synthesis. Our systematic investigation demonstrates that while a human GPI potently inhibits PxGP, the BPU insecticide diflubenzuron does not. This discrepancy is attributable to key substitutions in its aromatic halogen substituents, which preclude effective binding to GP’s active or allosteric sites. This provides the first direct biochemical evidence excluding GP as a candidate molecular target for diflubenzuron. Consequently, BPU lethality must arise from a different mechanism, likely involving interference with chitin assembly, cuticle organization, or CHS trafficking rather than direct enzymatic inhibition, a mechanism that remains to be conclusively identified [19, 40, 41].

It is important to delimit the scope of this conclusion. Our enzymatic and structural analyses establish that diflubenzuron, specifically, does not productively engage PxGP; they do not, by themselves, exclude every member of the structurally diverse BPU class as a potential GP ligand. BPU insecticides differ substantially in their aromatic halogen substituents and, in several cases, in additional ether or alkyl side chains (Figure 1), and our MM/GBSA analysis indicates that it is precisely these side-chain architectures—rather than the shared acylurea core—that dictate whether a given acyl urea can occupy the GP allosteric pocket (Table 1). Extending the present conclusion from diflubenzuron to the BPU class as a whole would therefore require systematic testing of additional compounds. The structural framework established here (Table 1) provides a rational basis for prioritizing such an effort, by predicting which acyl urea side-chain structures are sterically and energetically compatible with productive GP binding.

Metabolic plasticity and gluconeogenic compensation underlie GP non-essentiality

The exclusion of GP as a DFB target raised a more fundamental question: despite being a potent GP inhibitor in vitro, the GPI exhibited no insecticidal activity in vivo. This paradox—wherein a nanomolar-potency inhibitor of a putatively essential metabolic enzyme causes no phenotype—demanded mechanistic investigation.

Our investigation shows that this tolerance reflects a robust, developmentally-coordinated metabolic compensation mechanism (Figure 13). Even acute RNAi-mediated suppression of PxGP expression (87.59% knockdown) induced no discernible phenotype, as insects elicited a compensatory gluconeogenic pathway. This was evidenced by the concerted suppression of downstream glycogenolytic enzymes and the significant upregulation of the key gluconeogenic enzymes PEPCK (3.95-fold) and G-6-Pase (3.34-fold) at 96 h post-knockdown. Our new experimental data substantially strengthen this evidence: GP enzyme activity measurements confirm that RNAi-mediated transcript suppression translates to functional enzyme suppression in vivo (per-larva GP-a activity reduced by ∼29% at 24–48 h). The concurrent 30.78% decline in total protein provides independent biochemical support for protein catabolism. Furthermore, the selective upregulation of glycogen branching enzyme (GBE)—but not α-amylase—reveals a targeted compensatory response within the glycogen remodeling pathway. Fitness assessment demonstrates that this compensation, while carrying a transient larval weight cost (24–48 h), is ultimately effective: pupal weight, adult wing morphology, and female fecundity are fully maintained. This response provides a metabolic “escape route,” synthesizing glucose de novo from non-carbohydrate precursors to bypass the blocked glycogenolytic pathway and maintain homeostasis, particularly during the high carbohydrate demand preceding metamorphosis. Future work utilizing transcriptomic and proteomic approaches will be essential to delineate the signaling pathways (potentially involving nutrient sensors like AMPK and transcription factors such as FOXO) that link GP suppression to this precise reprogramming of metabolic gene expression [42, 43].

Schematic of glycogen metabolism and compensatory glucose homeostasis.

Red arrows: blocked flux from glycogen to G6P; Green arrows: activated gluconeogenesis. Solid arrows represent single enzymatic steps, while dashed arrows indicate multi-step processes.

Direct metabolite quantification reflected this compensation and revealed its sophisticated temporal dynamics. The most striking finding was the lack of glycogen accumulation upon GP knockdown, a paradox that could be explained by the activation of GP-independent glycogen catabolism via alternative enzymes like α-amylase (which hydrolyzes internal α-1,4-bonds) and glycogen debranching enzymes (which hydrolyze α-1,6-branch points) [32–34]. However, the selective upregulation of GBE (rather than α-amylase) argues against the activation of alternative catabolic routes. Instead, reduced flux into glycogen synthesis—possibly through feedback inhibition of glycogen synthase—represents a more plausible explanation, and this will be a key focus of future investigations. At 96 h, trehalose and G6P levels in dsGP-treated larvae were maintained at markedly higher levels than in dsGFP controls, which had declined by this time point, coinciding with 3–4-fold upregulation of PEPCK and G-6-Pase. These changes strongly support, albeit indirectly, de novo glucose synthesis as the primary rescue mechanism. This biphasic response—initial metabolic stress followed by delayed (48-72 h) and robust compensation—creates a temporal buffer by 96 h and is consistent with the absence of overt phenotypes despite substantial GP suppression. This temporal coordination raises the possibility of a developmentally programmed response, potentially involving hormonal regulation, that anticipates energy demands at critical transitions. Critically, this pattern suggests a potential vulnerability window (∼72 h) during which combined inhibition of glycogenolysis and gluconeogenesis might overcome compensation, although this remains speculative and would require experimental validation.

We also acknowledge that the absence of a severe phenotype may reflect an additional, non mutually exclusive explanation: under our experimental conditions, with continuous dietary carbohydrate availability, GP activity may not be rate limiting for maintaining glucose homeostasis. The observed transient larval weight loss could thus reflect both active metabolic compensation and the inherently low demand on GP for glucose supply in a feeding larva. Distinguishing between active compensation and the non limiting nature of GP will require future studies under nutrient restricted conditions or during fasting intervals, where GP’s role is likely to become more critical.

Implications for insecticide target selection: metabolic redundancy and the dual-target strategy

Our findings offer direct and actionable insights for the rational design of metabolism-targeting insecticides. A key lesson is that the viability of a metabolic enzyme as an insecticidal target is determined not solely by its biochemical function or pathway position but by the presence of compensatory metabolic routes. GP appears to fail as a standalone target, at least in part because of compensation via gluconeogenesis. This contrasts sharply with other carbohydrate-metabolizing enzymes that have proven to be highly effective insecticidal targets. For example, trehalase inhibitors such as Validamycin A are potent insecticides [44, 45], suggesting that trehalase inhibition may not be subject to the same degree of metabolic compensation. This distinction informs a refined strategy: dual-target inhibition. Simultaneously blocking both glycogenolysis (via GP) and gluconeogenesis (e.g., via PEPCK or G-6-Pase) would create irreversible glucose starvation, a “metabolic pincer” attack that might overcome adaptive compensation—a strategy with precedent in cancer therapy targeting both glycolysis and oxidative phosphorylation [46, 47]. Ultimately, our work highlights the value of systems-level metabolic modeling for target prioritization. Enzymes at metabolic branch points with redundant outputs are less vulnerable than those in non-bypassable pathways (e.g., chitin synthase, trehalase). Prospective application of computational flux balance analysis and constraint-based modeling, as supported by existing literature, can systematically identify such non-compensatable nodes for effective target selection [48].

GPI concentrations and pharmacokinetic considerations

The GPI concentrations used in our larval exposure experiments (250–500 mg/L) encompass a wide dose range. Even at the highest concentration (500 mg/L), which is approximately 409,000-fold above the in vitro IC50 of 2.96 nM, no toxicity was observed. This apparent discrepancy between potent in vitro inhibition and absent in vivo toxicity could reflect several pharmacokinetic factors: limited oral bioavailability due to incomplete absorption across the midgut epithelium; metabolic inactivation by larval detoxification enzymes (cytochrome P450s, esterases, glutathione S-transferases); or sequestration by hemolymph binding proteins [49–51]. However, the metabolic phenotype we observe—elevated trehalose, reduced protein, and upregulation of gluconeogenic enzymes—provides indirect evidence that GPI does reach its target, as these changes are consistent with GP inhibition. The most parsimonious interpretation is that GPI achieves sufficient target engagement to partially suppress GP activity in vivo, but that the resulting metabolic compensation renders this suppression non-lethal. This interpretation is further supported by our RNAi data, in which direct genetic suppression of GP (bypassing all pharmacokinetic barriers) similarly fails to cause mortality.

Study limitations and future directions

We acknowledge several limitations that point to productive future investigations. First, the transient nature of RNAi-mediated gene silencing in Lepidoptera precluded assessment of long-term or stage-specific consequences of GP loss; its role in adult flight or reproduction warrants investigation via CRISPR/Cas9 knockout or advanced dsRNA delivery systems [52–54]. Second, the generalizability of this gluconeogenic compensation mechanism across insect orders (e.g., Coleoptera, Diptera) remains to be tested, as metabolic regulation can vary with ecology and life history. Third, while our multi-layered evidence—from transcriptional upregulation to phenotypic rescue— provides robust qualitative support for the activation of gluconeogenesis, future studies employing 13C-isotope tracer analysis would be valuable to quantify the precise metabolic rates, identify primary carbon precursors, and further refine potential metabolic targets for insecticide development [55]. Extending this metabolic profiling to include key enzymes in amino acid metabolism (e.g., aspartate aminotransferase, glutamate aminotransferase) and the dynamics of lipid content would offer a more comprehensive understanding of the systemic metabolic reprogramming that accompanies GP loss.

A further consideration concerns the contribution of dietary carbohydrates to the glucose homeostasis we observed. Because our assays were performed on actively feeding larvae—and food consumption did not differ between dsGP and dsGFP groups (Figure 11B)—ingested dietary carbohydrate was continuously available and may have helped sustain hemolymph glucose alongside endogenous gluconeogenesis. Although the equivalent feeding rates argue that the metabolic phenotype is not a secondary consequence of altered food intake, our data cannot fully partition the relative contributions of de novo glucose synthesis and dietary carbohydrate absorption to the homeostatic buffering we documented. Starvation-challenge experiments, in which GP-suppressed larvae are deprived of dietary carbohydrate, would help isolate the endogenous gluconeogenic contribution and provide additional insight into the sufficiency of metabolic compensation under nutrient-limited conditions.

Concluding remarks

In conclusion, our biochemical data demonstrate that DFB does not inhibit PxGP, adding enzyme-level evidence to the existing genetic framework establishing CHS involvement in BPU activity. More importantly, our findings uncover a fundamental principle of insect metabolic biology: robust compensatory pathways, specifically gluconeogenesis, can buffer against the loss of key catabolic enzymes. This metabolic plasticity renders GP non-viable as a standalone insecticidal target but illuminates a potential strategic path forward: the development of dual-target inhibitors that simultaneously block glycogenolysis and gluconeogenesis. Our structural analysis provides a molecular framework for understanding why specific acyl urea side-chain structures succeed or fail as GP inhibitors, which may inform future inhibitor design. Our findings underscore the necessity of adopting a systems-level perspective in target-based insecticide discovery. Biochemical potency in vitro is necessary but not sufficient for in vivo efficacy; the latter requires consideration of metabolic network architecture, pathway redundancy, and compensatory mechanisms. As the pest management community confronts the twin challenges of insecticide resistance and environmental sustainability, this mechanistic understanding of metabolic robustness will be essential for designing the next generation of selective, effective and durable insect control agents.

Materials and Methods

Insects and rearing conditions

A susceptible laboratory strain of P. xylostella (L.), maintained for multiple generations at the College of Plant Protection, Northwest A&F University, without prior insecticide exposure, was used for all experiments. Larvae were reared on radish (Raphanus sativus) seedlings grown in vermiculite. Adults were provisioned with a 10% (w/v) honey solution. All insect stages were maintained in a controlled environment at 25 ± 1°C, 60 ± 5% relative humidity (RH), and under a 16:8 h (light:dark) photoperiod.

Cell culture and maintenance

Spodoptera frugiperda (Sf9) cells (Gibco, CA, USA) (Cat# 11496015) were cultured in Sf-900™ III SFM medium (Gibco) supplemented with 10% fetal bovine serum (FBS; Gibco) and 1% (v/v) Antibiotic-Antimycotic solution (Gibco, Cat# 15240062). Adherent monolayer cultures were maintained in T-flasks at 27°C. For large-scale protein expression, cells were adapted to suspension culture in baffled flasks, agitated at 120 rpm at 27°C. Cells were harvested by centrifugation (4,000 rpm, 10 min, 4°C) when the density reached 5.0 × 106 cells/mL. Cell pellets were flash-frozen in liquid nitrogen and stored at −80°C until use. The supplier performs authentication and mycoplasma testing as documented in the certificate of analysis available on the manufacturer’s website. The cells were used experimentally within one year of purchase and were not re authenticated in our laboratory.

Chemicals and reagents

Diflubenzuron (DFB, 1-(4-chlorophenyl)-3-(2,6-difluorobenzoyl) urea, ≥98% purity) was purchased from Shanghai Topscience Co., Ltd. (Shanghai, China, Cat# T0109).

The Glycogen phosphorylase-IN-1 (GPI), 1-(3-(3-(2-Chloro-4,5-difluorobenzoyl)ureido)-4-methoxyphenyl)-3-methylurea (also known as compound 42 in ref [15]), was purchased from TargetMol (Shanghai, China, Cat# T37577, ≥98% purity).

Glycogen (from bovine liver, Type III), β-glycerophosphate, HEPES, imidazole, EDTA, DTT, ATP, MgCl2, protease inhibitor cocktail components (AEBSF, Pepstatin A, Leupeptin), and all other biochemical reagents were purchased from Sigma-Aldrich (Shanghai, China) or Sangon Biotech (Shanghai, China) and were of molecular biology grade (≥95% purity) unless otherwise specified.

Recombinant PxGP expression and purification

The full-length coding sequence of PxGP (2,511 bp, encoding 837 amino acids; GenBank Accession No. XM_038119125.2) was amplified from P. xylostella third-instar larval cDNA by PCR using gene-specific primers (Supplementary file 1), and cloned into the pFastBac1 vector (Invitrogen, Carlsbad, CA, USA) using homology-based cloning (ClonExpress II One Step Cloning Kit, Vazyme, Nanjing, China). The construct was engineered to express PxGP with an N-terminal 6×His tag (HHHHHH-) immediately following the initiating methionine, under the control of the polyhedrin (PH) promoter. The sequence fidelity of the construct was verified by Sanger sequencing (Tsingke Biotechnology, Beijing, China).

The sequence-verified pFastBac1-PxGP construct was transformed into DH10Bac™ competent cells (Invitrogen) to generate recombinant bacmid DNA via Tn7-mediated transposition, following the manufacturer’s Bac-to-Bac® protocol. Recombinant bacmid DNA was isolated, verified by PCR using M13 forward and reverse primers (Supplementary file 1), and transfected into Sf9 cells using Cellfectin® II Reagent (Invitrogen) to generate P0 viral stock. The viral titer was amplified through two additional passages (P1, P2) to obtain high-titer P3 viral stock.

For large-scale protein production, Sf9 cells in suspension culture (2.0 × 106 cells/mL, 500 mL culture volume in a 2-L baffled Erlenmeyer flask) were infected with P3 viral stock at a Multiplicity of Infection (MOI) of 5. Infected cultures were maintained at 27°C with agitation at 120 rpm. Cells were harvested 72 h post-infection by centrifugation (4,000 rpm, 10 min, 4°C), yielding approximately 8.25 g of cell pellet. Cell pellets were flash-frozen in liquid nitrogen and stored at −80°C until use.

For protein purification, frozen cell pellets were thawed on ice and resuspended in ice-cold Lysis Buffer (50 mM β-glycerophosphate [pH 7.5], 150 mM NaCl, 0.2 mM DTT, 0.5 mM EDTA, and 1× protease inhibitor cocktail [1 mM AEBSF, 15 µM Pepstatin A, 20 µM Leupeptin]) at a ratio of 4 mL buffer per gram of cell pellet. Cells were lysed by sonication on ice using a Scientz-IID sonicator (Ningbo Scientz Biotechnology, China) with the following parameters: 6 cycles of 30 s ON / 60 s OFF at 40% amplitude. The lysate was clarified by centrifugation (16,000 × g, 30 min, 4°C).

The clarified supernatant was loaded onto a 5-mL HisTrap™ HP column (GE Healthcare, Uppsala, Sweden) pre-equilibrated with Lysis Buffer containing 10 mM imidazole. The column was first washed with 10 column volumes (CV) of Equilibration Buffer (Lysis Buffer + 10 mM imidazole), followed by 5 CV of Wash Buffer (Lysis Buffer + 30 mM imidazole). Bound 6×His-tagged PxGP protein was eluted with a 15 CV linear gradient from 30 to 200 mM imidazole in Lysis Buffer. Fractions (2 mL each) were collected and analyzed by 13% SDS-PAGE with Coomassie Brilliant Blue R-250 staining. Fractions containing PxGP, identified as a prominent band at ∼100 kDa, were pooled and dialyzed overnight at 4°C against Storage Buffer (50 mM HEPES [pH 7.4], 150 mM NaCl, 1 mM DTT, 10% [v/v] glycerol) using a 10 kDa molecular weight cut-off dialysis membrane (Spectrum Labs, Rancho Dominguez, CA, USA), with three buffer changes. Protein concentration was determined using the Pierce™ BCA Protein Assay Kit (Thermo Fisher Scientific, Waltham, MA, USA) with bovine serum albumin (BSA) as the standard. Aliquots of purified PxGP were flash-frozen in liquid nitrogen and stored at −80°C. The typical yield was 7.5 mg of purified protein from 8.25 g of cell pellet.

In vitro activation of recombinant PxGP

Purified recombinant PxGP exists predominantly in the inactive dephosphorylated b form (PxGP-b) and requires phosphorylation at a conserved serine residue (corresponding to Ser14 in mammalian GP) for conversion to the active a form (PxGP-a) [17]. Activation was performed in vitro using immobilized Phosphorylase Kinase (PhK) (Cat# HY-P2757, MedChemExpress, Monmouth Junction, NJ, USA).

Immobilized PhK beads (100 µL slurry) were washed three times with 1 mL of Kinase Buffer (25 mM β-glycerophosphate, pH 7.5, 0.5 mM EDTA, 0.25 mM DTT, 1× protease inhibitor cocktail) and resuspended in 500 µL of dialyzed PxGP-b solution (final protein concentration: 1 mg/mL). The activation reaction was initiated by adding ATP (sodium salt) to a final concentration of 7 mM and MgCl2 to 22 mM. The suspension was incubated for 2 h at room temperature (22-25°C) with gentle end-over-end rotation.

Following incubation, the activated PxGP-a (present in the supernatant) was separated from the immobilized PhK beads using a magnetic separation stand. The supernatant was passed through a 0.22 µm syringe filter (Millipore) to remove any residual beads and was used immediately for enzymatic assays. The extent of activation was assessed by measuring specific activity (see glycogen phosphorylase activity assay). Activated PxGP-a was used fresh and not stored, as phosphorylated GP is subject to dephosphorylation by endogenous phosphatases over time.

Western blot analysis

Protein samples were separated by 13% SDS-PAGE (120 V, 90 min) and electrophoretically transferred to a nitrocellulose (NC) membrane (0.45 µm pore size, Millipore, Burlington, MA, USA) at 300 mA for 90 min at 4°C using a Bio-Rad Mini Trans-Blot® system. The membrane was blocked for 1 h at room temperature in 5% (w/v) non-fat dry milk in TBST (10 mM Tris-HCl, pH 7.4, 150 mM NaCl, 0.1% [v/v] Tween-20).

The membrane was incubated overnight at 4°C with an anti-His mouse monoclonal antibody (1:1500 dilution in blocking buffer; Cat# ABT2050, Abbkine, Wuhan, China). After three washes with TBST (10 min each), the membrane was incubated for 1 h at room temperature with an alkaline phosphatase (AP)-conjugated goat anti-mouse IgG secondary antibody (1:4000 dilution in blocking buffer; Cat# A0208, Beyotime Biotechnology, Shanghai, China). Following three additional TBST washes, immunoreactive bands were visualized using BM Purple AP Substrate (Roche, Basel, Switzerland) according to the manufacturer’s instructions. Images were captured using a ChemiDoc™ XRS+ imaging system (Bio-Rad, Hercules, CA, USA).

Glycogen phosphorylase activity assay

GP enzymatic activity was measured using a coupled-enzyme spectrophotometric assay kit (Cat# BC3345, Solarbio Life Science, Beijing, China). This assay quantifies GP activity by measuring the GP-catalyzed phosphorolysis of glycogen to glucose-1-phosphate (G-1-P). The product G-1-P is enzymatically converted to glucose-6-phosphate (G6P) by phosphoglucomutase (PGM), and G6P is subsequently oxidized by glucose-6-phosphate dehydrogenase (G6PDH), which stoichiometrically reduces NADP+ to NADPH. The rate of NADPH formation, which is directly proportional to GP activity, was monitored by measuring the increase in absorbance at 340 nm (A340) using a Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland).

The reaction mixture (200 µL total volume per well in a 96-well UV-transparent microplate) contained the following components (as supplied in the kit): 160 µL of Reaction Buffer (containing glycogen [2 mg/mL final], inorganic phosphate [10 mM], AMP [1 mM], NADP+ [0.4 mM], PGM, G6PDH, MgCl2, and appropriate buffer), 30 µL of deionized water, and 10 µL of enzyme sample (purified PxGP-a or crude larval lysate). The reaction was initiated by adding the enzyme sample, and A340 was recorded continuously at 15 s intervals for 30 min at 25°C.

Specific activity was calculated using the following formula: Specific Activity(U/mg)=(ΔA340/min)/(ε×1×[Protein])×Vt/Vs×1000. where: ΔA340 represents the change in absorbance at 340 nm, ε = molar extinction coefficient of NADPH at 340 nm (6.22 mM-1 · cm-1), l = light path length (0.6 cm for 200 µL in a 96-well microplate), [Protein] = protein concentration in the assay (mg/mL), Vt = total reaction volume, Vs = sample volume added.

One Unit (U) of GP activity is defined as the amount of enzyme that catalyzes the formation of 1 µmol of NADPH per minute under the assay conditions.

Inhibitor screening and IC50 determination

For in vitro enzyme inhibition assays, purified activated PxGP-a was diluted to 10 µg/mL in Assay Dilution Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl2). Test compounds (GPI, DFB) were prepared as 100× stock solutions in DMSO and serially diluted in Assay Dilution Buffer to achieve final assay concentrations ranging from 0.16 nM to 500 nM (for GPI) and 0.54 nM to 1700 nM (for DFB), respectively (final DMSO concentration ≤ 1% [v/v] in all wells, including vehicle controls).

The enzyme (10 µL of 10 µg/mL PxGP-a, final concentration 10 µg/mL in the assay) was pre-incubated with 10 µL of inhibitor solution (or vehicle control) for 15 min at room temperature in a 96-well microplate. The reaction was then initiated by adding 180 µL of the Reaction Buffer (as described in glycogen phosphorylase activity assay), and A340 was monitored continuously for 30 min at 25°C.

Residual enzyme activity at each inhibitor concentration was calculated as a percentage of the vehicle control (1% DMSO alone). Dose-response curves were generated by plotting percent residual activity versus log[inhibitor]. IC50 values (the concentration of inhibitor required to reduce enzyme activity by 50%), Hill slopes, and 95% confidence intervals (CI) were calculated by non-linear regression analysis using a four-parameter logistic equation in GraphPad Prism 9.0 (GraphPad Software Inc., San Diego, CA, USA): Y=Bottom+(Top−Bottom)/(1+10(LogICS 50−X)×Hill slope). where Y is percent residual activity, X is log[inhibitor], Top and Bottom are the upper and lower plateaus of the curve (constrained to 100% and 0%, respectively), and Hill slope describes the steepness of the curve.

For each compound, at least three independent experiments were performed, each with triplicate technical replicates.

Preparation of larval crude lysate for in vivo activity

To assess the inhibitory effects of GPI and DFB on native GP activity in vivo, crude enzyme extracts were prepared from P. xylostella larvae. Healthy third-instar (L3) larvae (n = 30 per biological replicate, three replicates total) were collected, briefly washed in ice-cold phosphate-buffered saline (PBS, pH 7.4) to remove surface contaminants, and gently blotted dry on filter paper. Larvae were then homogenized on ice in 3 volumes (w/v) of ice-cold Homogenization Buffer (30 mM HEPES, pH 7.2, 60 mM KCl, 1.5 mM EDTA, 1.5 mM MgCl2, 1× protease inhibitor cocktail).

The homogenate was centrifuged at 15,000 × g for 20 min at 4°C, and the supernatant (crude enzyme extract) was carefully collected, avoiding the lipid layer and pellet. Protein concentration was determined using the BCA assay, and extracts were either used immediately for activity assays or aliquoted and stored at −80°C (activity remained >90% for up to 2 weeks).

GP activity in crude lysates was measured using the same coupled-enzyme assay described in glycogen phosphorylase activity assay, with the following modifications: crude lysate (diluted to 2 mg/mL total protein) was used in place of purified enzyme, and pre-incubation with inhibitors was performed for 5 to 45 min to assess time-dependent inhibition. Final inhibitor concentrations tested were 10-5, 10-6 or 10-7 M.

Sequence alignment and bioinformatic analysis

The amino acid sequence of PxGP (XP_037975053.2) was obtained from the NCBI database. Multiple sequence alignment with orthologs from Homo sapiens (NP_002854.3), Oryctolagus cuniculus (NP_001075653), Helicoverpa armigera (XP_021190210.1), Spodoptera exigua (ACN78408.1), and Ostrinia furnacalis (AFO54708.2) was performed using ClustalX. Conserved domains and motifs were analyzed using the ScanProsite and NCBI CD-Search web servers. Protein pairwise similarity matrix was visualized as a heatmap using the “HeatMap” module in TBtools-II (v2.390) [56].

Larval bioassays

Larval bioassays were conducted using a standard leaf-dip method adapted from established protocols [57]. Stock solutions of GPI and DFB were prepared in HPLC-grade acetone at 10,000 mg/L. Serial dilutions were prepared in deionized water containing 0.1% (v/v) Triton X-100 as a wetting agent to achieve final test concentrations of 250 and 500 mg/L for GPI, and 125 and 250 mg/L for DFB. The final acetone concentration in all treatment solutions was 1% (v/v). A solvent control group (CK) consisted of 1% acetone and 0.1% Triton X-100 in water without any test compound.

Fresh cabbage (Brassica oleracea var. capitata) leaf discs (5 cm diameter) were cut using a cork borer, dipped in the respective treatment solutions for 6-8 s with gentle agitation, and then air-dried for 30 min at room temperature in a fume hood. Once dry, each leaf disc was placed in a sterile 90 mm petri dish lined with moistened filter paper.

Ten synchronized third-instar larvae of P. xylostella (24-36 h post-molt to L3) were transferred to each treated leaf disc using a soft brush. Each treatment consisted of three biological replicates (i.e., three Petri dishes, 30 larvae total per treatment). Petri dishes were sealed with Parafilm® and maintained under standard rearing conditions (25 ± 1°C, 60 ± 5% RH, 16:8 h L:D photoperiod). Leaf discs were replaced with freshly treated discs every 12 h to ensure continuous exposure.

Mortality was recorded every 24 h for 120 h (5 days). Larvae that failed to respond to gentle prodding or showed moribund behavior (inability to right themselves or move coordinately) were considered dead. Corrected mortality was calculated using Abbott’s formula [58]: Corrected Mortality (%) =[ (Treatment Mortality - Control Mortality) / (100 - Control Mortality)]×100. Pupation rates (number of pupae formed / initial number of larvae × 100) and adult emergence rates (number of emerged adults / number of pupae formed × 100) were recorded at 8 and 10 days post-treatment, respectively.

Lethal concentration values (LC30 and LC50) and 95% confidence intervals for DFB were calculated using probit analysis in POLO-Plus software (LeOra Software, Berkeley, CA, USA).

GPI exposure for gene expression analysis

For time-course gene expression analysis, synchronized third-instar larvae were exposed to 500 mg/L GPI (the highest concentration tested in bioassays) or solvent control (1% acetone + 0.1% Triton X-100) using the leaf-dip method described in larval bioassays.

Surviving larvae were collected at 24, 48, 72, and 96 h post-treatment. At each time point, three biological replicates (10 pooled larvae each) were flash-frozen in liquid nitrogen and stored at −80°C until RNA extraction. This sampling scheme ensured sufficient RNA yield while minimizing inter-individual variability.

Quantitative real-time PCR (RT-qPCR) analysis

Total RNA was extracted from frozen samples using RNAiso Plus reagent (TaKaRa, Kusatsu, Japan). RNA concentration and purity were assessed by NanoDrop 2000c spectrophotometer (Thermo Fisher Scientific). Only samples with A260/A280 ratios between 1.9 and 2.1 and A260/A230 ratios >1.8 were used for downstream analysis. RNA integrity was confirmed by agarose gel electrophoresis (sharp 28S/18S rRNA bands, ratio ≈ 2:1). cDNA was synthesized from 1 µg RNA using PrimeScript™ RT Reagent Kit with gDNA Eraser (TaKaRa).

Gene-specific primers for PxGP, trehalase (PxTre), hexokinase (PxHex), phosphoenolpyruvate carboxykinase (PxPEPCK), PxG-6-Pase, and the reference gene PxRPS13 (ribosomal protein S13) were designed using Primer-BLAST (NCBI) and are listed in Supplementary file 1. PxRPS13 was selected based on its stable expression across developmental stages and treatment conditions (geNorm M value < 0.5) [59]. All primers were validated for specificity (single peak in melt-curve), efficiency (90-110%, R² > 0.98), and absence of primer-dimers (no-template controls). RT-qPCR reactions were performed on a QuantStudio™ 5 Real-Time PCR System (Thermo Fisher Scientific) using TB Green® Premix Ex Taq™ II (Tli RNaseH Plus) (2×) (TaKaRa) in 20 µL reactions. The thermal cycling conditions were as follows: initial denaturation at 95°C for 30 s; 40 cycles of 95°C for 15 s (denaturation) and 60°C for 30 s (annealing/extension); followed by melt-curve analysis. Each sample was analyzed in technical triplicate, and each biological replicate consisted of three independent samples. Relative gene expression was calculated using the 2-ΔΔCt method [60].

RNA interference (RNAi)

A conserved 606 bp region of PxGP (nucleotides 66-671, part of the AMP-binding domain) was selected for RNAi. The fragment was amplified from larval cDNA using primers flanked by T7 promoters (Supplementary file 1). Double-stranded RNA (dsRNA) was synthesized with the T7 RiboMax™ Express RNAi System (Promega, Madison, WI, USA). The dsRNA of green fluorescent protein (dsGFP, 541 bp) was synthesized using a pGFP plasmid template as a non-specific negative control. Following synthesis, dsRNA was treated with RNase-free DNase I (30 min at 37°C), purified by isopropanol precipitation, and integrity verified by 1.5% agarose gel electrophoresis. Concentration and purity (A260/A280=1.9-2.1) was measured using a NanoDrop spectrophotometer. Purified dsRNA was resuspended in nuclease-free water at concentrations of 6,000, 10,000, or 14,000 ng/µL and stored in single-use aliquots at −80°C.

For microinjection, synchronized early third-instar larvae (6-12 h post-molt) were briefly chilled on ice for 2-3 min and injected laterally using a DMP-300 Injector (Micrology Precision Instruments, Ltd, Wuhan, China). After recovery on fresh cabbage leaves for 30 min at room temperature, larvae were maintained under standard rearing conditions. Phenotypic parameters (mortality, molting success, pupation rate, adult emergence) were monitored daily for 10 days. Larvae were collected at 24, 48, 72, and 96 h post-injection for RT-qPCR analysis to confirm target gene knockdown (n = 3 biological replicates, each containing 10 pooled larvae per time point).

Metabolite quantification

To directly assess metabolic flux alterations following PxGP knockdown, we quantified four key metabolites in the glycogen-glucose metabolic axis using commercial enzymatic assay kits: glycogen and glucose (Glycogen Assay Kit, Grace Biotechnology, Suzhou, China, Cat# G0590W96), trehalose (Trehalose Assay Kit, Grace Biotechnology, Cat# G0553W96), and glucose-6-phosphate (G6P Assay Kit, Beyotime Biotechnology, Shanghai, China, Cat# S0185). All kits utilize colorimetric or fluorometric enzymatic reactions and were performed according to the manufacturers’ protocols with minor modifications as described below.

Third-instar larvae injected with dsGP (14,000 ng/µL, the highest dose achieving maximal knockdown) or dsGFP control were collected at 48, 72, and 96 h post-injection. At each time point, three biological replicates were collected, each consisting of 10 pooled larvae. Samples were immediately flash-frozen in liquid nitrogen and stored at −80°C until analysis.

For metabolite extraction, frozen larvae were homogenized on ice in ice-cold extraction buffer specific to each metabolite assay (typically 100 µL buffer per 10 mg tissue). For glycogen extraction, samples were homogenized in deionized water and immediately boiled for 3 min to inactivate endogenous glycogenolytic enzymes. For trehalose, glucose, and G6P extraction, samples were homogenized in the respective assay buffers provided in the kits. Homogenates were centrifuged at 12,000 × g for 10 min at 4°C, and supernatants were immediately used for metabolite quantification or stored at −80°C for no more than 24 h.

Metabolite quantification was performed in 96-well microplates using Infinite 200 PRO microplate reader (Tecan, Männedorf, Switzerland) at the wavelengths specified in each kit’s protocol. Sample metabolite concentrations were calculated by comparison with the absorbance of the kit’s standard. Metabolite concentrations were normalized to tissue protein content to account for variation in tissue input. Protein concentration in the crude lysates was determined using the Protein Quantification Kit (BCA Assay) (Abbkine Scientific) with bovine serum albumin as the standard. All samples were analyzed in technical triplicate. Final metabolite levels were expressed as: Glycogen: mg/mg protein; Glucose: µmol/mg protein; Trehalose: µmol/mg protein; G6P: µmol/mg protein. Because metabolite levels are expressed per unit total protein, the changes in total protein content that accompany PxGP knockdown bear directly on their interpretation. Total protein declined substantially at the earlier time points (up to ∼30%; Figure 9D, Figure 10–figure supplement 2) before recovering, which means that for any metabolite whose absolute amount was unchanged, the protein-normalized value would be correspondingly inflated during the protein-depleted window. The protein-normalized metabolite data reported here should therefore be interpreted with this denominator effect in mind. We retained protein normalization because it remains the most appropriate correction for the variation in tissue input inherent to pooled-larvae sampling; we note, however, that absolute (per-larva) quantification — as we additionally applied to GP enzyme activity (Figure 10A, B) — provides a complementary, denominator-independent perspective and represents a useful refinement for future metabolite profiling.

Glycogen phosphorylase activity assay in RNAi-treated larvae

To assess GP enzyme activity following RNAi, early third-instar larvae were injected with 14,000 ng/μL dsGP or dsGFP. At 24, 48, 72, and 96 h post-injection, 10 larvae per replicate (three biological replicates per time point) were collected. Active glycogen phosphorylase a (GP-a) activity was measured using a coupled-enzyme spectrophotometric assay kit (catalog no. BC3345, Solarbio, Beijing, China) following the manufacturer’s protocol as described above. GP activity was normalized both to total protein content (U/mg protein) and to individual larva (U/larva).

RT-qPCR analysis of alternative glycogen metabolic enzymes

Homologs of human glycogen branching enzyme (GBE, CAG9111719.1), Pieris rapae α-amylase (XP_011548197.3), and glycogen debranching enzyme (GDE, XP_048484812.1) were identified in the P. xylostella genome database by sequence homology search. RT-qPCR primers were designed using Primer-BLAST (NCBI) (Supplementary file 1). Multiple primer pairs designed for GDE failed to yield amplification products; therefore, only GBE and α-amylase were analyzed. PxRPS13 (ribosomal protein S13) was used as the reference gene. Total RNA was extracted from larvae collected at 24, 48, 72, and 96 h post-injection (three biological replicates, 10 larvae per replicate). Relative expression levels were calculated using the 2−ΔΔCt method and analyzed by two-way ANOVA with Šidák’s multiple comparisons test [61].

Fitness cost assessment

To evaluate fitness consequences of PxGP knockdown, the following parameters were measured. Feeding rate: cabbage leaf discs of uniform size were provided every 12 h, and remaining leaf area was measured by ImageJ at 24 h intervals from 24 to 120 h post-injection. Larval weight: 10 larvae per replicate were weighed at 24 h intervals. Pupal weight: all pupae were weighed on day 3 after pupation. Adult wing area: forewings were photographed using a SteREO Discovery V20 stereomicroscope (Zeiss, Germany) and measured using ImageJ. Female fecundity: newly emerged females were paired with untreated males, and eggs deposited on leaf discs were counted for 3 consecutive days. Three biological replicates were performed. Data were analyzed by independent-samples t-test.

Homology modeling and molecular docking

The three-dimensional structure of PxGP was generated by homology modeling using SWISS-MODEL (https://swissmodel.expasy.org [62]) with the crystal structure of human liver glycogen phosphorylase a in complex with an acyl urea inhibitor (PDB: 2ATI [15]) as the template. Since PxGP functions as a homodimer, the dimeric biological assembly was constructed based on the template’s quaternary structure. The three-dimensional structures of GPI (PubChem CID: 10070301) and DFB (PubChem CID: 37123) were obtained from PubChem in SDF format. Molecular docking was performed using Cresset Flare V11 (Cresset, Litlington, UK). Binding free energies were estimated by MM/GBSA (Molecular Mechanics with Generalized Born and Surface Area solvation) single-point calculations. Two-dimensional protein-ligand interaction diagrams were generated within Flare V11.

Use of artificial intelligence tools

AI-based language editing tools were used to assist with English grammar and style refinement during manuscript preparation. All scientific content, including experimental design, data collection, analysis, and interpretation, was generated entirely by the authors.